### 2024-02-03 products/manuscript/Manuscript.qmd
Part 1 for Project

---
title: "Manuscript for MADA Project"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../manuscript/dataanalysis-references.bib

---

Project Part 1: This dataset was obtained from Kaggle.com and was created by Mark Otto and Andrew Fong at MIT. The set has 373 observations for 13 variables covering a broad spectrum associated with sleep and daily routine.

```{r}
#load in the data and check it out
library(here)
sleepdata <- read.csv(here("data","raw-data", "ss.csv"))
summary(sleepdata)
structure(sleepdata)
class(sleepdata)
```

We want to explore the influence of each variable on sleep quality and identify which ones seem to have the most impact (both positive and negative). We also wish to explore if certain variables affect male sleep quality more than females. If time allows, we'd also like to see if certain ages seem to be affected by certain variables more often than others. We can focus on every predictor in the dataset since there are only 13.

We can likely use GLMs and t-tests/ANOVA tests to see the significance that variables may have in different combinations on an individual's sleep quality. Heat maps and various plots will also help with exploratory data analysis. We aren't exceptionally experienced with statistical analyses, but these are what we can think of at the moment.

End of Project Part 1

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
#knitr::include_graphics(here("products","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

 shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
#resulttable=readRDS("../../results/summarytable.rds")
#knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
#knitr::include_graphics(here("results","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Exampleshows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
#knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper  discusses types of analyses.

These papers  are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2024-02-20 R/eda-code/eda.qmd
Emma-Part 2 for Project

---
title: "Exploratory Data Analysis: Project Part 2"
author: "Kevin Kosewick & Emma Hardin-Parker"
date: "2/20/2024"
output: html_document
---

This document contains our exploratory analysis of our chosen data set, sleepdata. We intend to perform any needed cleaning and mutation to the data as well as creating exploratory plots and tables to gain a better understanding of the data. As stated in the manuscript, we are looking to determine which variables or identifiers have significant impacts on sleep quality.

# Setup

```{r}
#Loading necessary libraries
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyverse)
```
# Getting to know the data

Load & View the data.

```{r}
sleepdata <- read.csv(here("data","raw-data", "ss.csv"))
View(sleepdata)
```

Exploring the structure and format of the data.

```{r}
summary(sleepdata)
structure(sleepdata)
```
The data set contains 373 observations for 13 variables. Some variables are not easily understandable at face value, so we went back to the data source and accessed the data dictionary. The variables are defined as follows:

Person ID: A unique identifier for each participant.
Gender: The participant's gender (Male/Female).
Age: The participant's age in years.
Occupation: The job or career of the participant.
Sleep Duration (hours): The daily sleep duration of the participant in hours.
Quality of Sleep (scale: 1-10): A subjective assessment of sleep quality on a scale from 1 to 10.
Physical Activity Level (minutes/day): The daily duration of physical activity for the participant, measured in minutes.
Stress Level (scale: 1-10): A subjective assessment of the participant's stress level on a scale from 1 to 10.
BMI Category: The participant's BMI classification (e.g., Underweight, Normal, Overweight).
Blood Pressure (systolic/diastolic): The participant's blood pressure, represented as systolic pressure over diastolic pressure.
Heart Rate (bpm): The participant's resting heart rate, measured in beats per minute.
Daily Steps: The number of steps the participant takes each day.
Sleep Disorder: The presence or absence of a sleep disorder in the participant (None, Insomnia, Sleep Apnea).

Before going further into the exploratory analysis, we need to determine whether missing values are present and if they could jeopardize any future analyses.

```{r}
print(sum(is.na(sleepdata)))
```

There are no missing values in our data set! We can now move on to creating new variables or transmuting existing variables.

# Data Transformation

We are going to create a categorical variable to go alongside the existing blood pressure variable. We are using blood pressure guidelines from the American Heart Association to create our new categorical blood pressure variable. Our goal with this is to see how blood pressure can affect sleep quality in a more digestible way. First, we will need to create new systolic and diastolic variables to ultimately create our categorical variable. We have attached the guidelines here:

[![AHA Guidelines](AHA_Guidelines.png)](https://www.heart.org/en/health-topics/high-blood-pressure)

```{r}
# Splitting the Blood.Pressure variable into systolic and diastolic variable columns
sleepdata <- sleepdata %>%
  separate(Blood.Pressure, into = c("systolic", "diastolic"), sep = "/")

# Ensuring that the new columns are considered numeric 
sleepdata$systolic <- as.numeric(sleepdata$systolic)
sleepdata$diastolic <- as.numeric(sleepdata$diastolic)

# Creating new categorical variable

sleepdata <- sleepdata %>%
  mutate(cat_bp = case_when(
    systolic < 120 & diastolic < 80 ~ "Normal",
    systolic >= 120 & systolic <= 129 & diastolic < 80 ~ "Elevated",
    (systolic >= 130 & systolic <= 139) | (diastolic >= 80 & diastolic <= 89) ~ "Stage 1 Hypertension",
    systolic >= 140 | diastolic >= 90 ~ "Stage 2 Hypertension"
  ))
```

Now that we have created this new variable and adjusted the format of the blood pressure variables, we are now going to further explore the other variables to see if they are fit, as is, for analysis. We noticed that within the BMI.Category variable, there are both "Normal" and "Normal Weight" categories. We assume this was an oversight in data entry, and intend to merge the two together to ensure all "Normal" categories are together for subsequent analyses.

```{r}
# Merging "Normal" and "Normal Weight" attributes into a singular "Normal" category

sleepdata <- sleepdata %>%
  mutate(BMI.Category = case_when(
    BMI.Category %in% c("Normal", "Normal Weight") ~ "Normal",
    BMI.Category == "Overweight" ~ "Overweight",
    BMI.Category == "Obese" ~ "Obese",
  ))
```

As of this point in our cleaning, we feel confident that the data is structured in such a way that analyses can be performed.

### 2024-02-20 products/manuscript/Manuscript.qmd
Emma-Part 2 for Project

@@
 
-Project Part 1: This dataset was obtained from Kaggle.com and was created by Mark Otto and Andrew Fong at MIT.
+Project Part 1: This dataset was obtained from Kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data) and was created by Mark Otto and Andrew Fong at MIT.
 The set has 373 observations for 13 variables covering a broad spectrum associated with sleep and daily routine.

### 2024-02-22 R/eda-code/eda.qmd
Kevin- Part 2 for Project
EDA plots and updated Manuscript document to include variables of interest.

@@
 output: html_document
+editor: 
+  markdown: 
+    wrap: sentence
 ---
@@
 ```
+
 # Getting to know the data
@@
 ```
+
 The data set contains 373 observations for 13 variables.
@@
 
-There are no missing values in our data set! We can now move on to creating new variables or transmuting existing variables.
+There are no missing values in our data set!
+We can now move on to creating new variables or transmuting existing variables.
 
@@
 As of this point in our cleaning, we feel confident that the data is structured in such a way that analyses can be performed.
+
+# Data Visualization and Exploratory Data Analysis
+
+We will now begin to create some figures to help visualize the relationships in the dataset.
+Quality of Sleep is our outcome of interest; we'll explore the effect each other variable seems to have on it.
+First, we'll start with the variables that aren't health specific.
+I asked Microsoft Copilot in Precise Mode to generate the base code for some plots that I then changed them to be specific to the dataset/have more visual appeal.
+
+```{r}
+# Boxplot for Quality of Sleep by Gender
+ggplot(sleepdata, aes(x = Gender, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Gender", x = "Gender", y = "Quality of Sleep")
+
+# Create bins for Age in order to create a clean boxplot.
+sleepdata$AgeGroup <- cut(sleepdata$Age, breaks = c(20, 30, 40, 50, 60), labels = c("20-30", "30-40", "40-50", "50-60"), include.lowest = TRUE)
+
+# Calculate mean Quality of Sleep for each Age Group to plot with means for cleaner visuals
+sleepdata_summary <- sleepdata %>%
+  group_by(AgeGroup) %>%
+  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
+
+# Bar plot for Mean Quality of Sleep by Age Group with colorblind-friendly palette
+ggplot(sleepdata_summary, aes(x = AgeGroup, y = MeanQuality, fill = AgeGroup)) +
+  geom_bar(stat = "identity") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(title = "Mean Quality of Sleep by Age Group", x = "Age Group", y = "Mean Quality of Sleep")
+
+# Calculate mean Quality of Sleep for each Occupation in order to create a bar plot.
+sleepdata_summary <- sleepdata %>%
+  group_by(Occupation) %>%
+  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
+
+# Bar plot for Mean Quality of Sleep by Occupation with manually specified colors
+ggplot(sleepdata_summary, aes(x = Occupation, y = MeanQuality, fill = Occupation)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = rainbow(length(unique(sleepdata_summary$Occupation)))) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Mean Quality of Sleep by Occupation", x = "Occupation", y = "Mean Quality of Sleep")
+
+```
+
+We can see some clear differences in self-reported Quality of Sleep between these different groups already.
+Females tend to report an average of 1 higher according to the boxplot.
+The different age groups are similar, but the oldest (50-60) reports the highest quality of sleep by far.
+Our occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
+We'll now explore Physical Activity and plot it as a predictor for Quality of Sleep.
+To make it easier to visualize Physical Activity, we'll create a new variable with categories of Low (0-30 minutes), Medium (30-60), and High (60+).
+We'll again use Microsoft Copilot for the basic plot code and modify it to our dataset/preferences.
+
+```{r}
+# Histogram for Physical Activity Level
+ggplot(sleepdata, aes(x = Physical.Activity.Level)) +
+  geom_histogram(binwidth = 10) +
+  labs(title = "Histogram of Physical Activity Level", x = "Physical Activity Level (minutes/day)", y = "Count")
+
+# Create a new variable for Physical Activity Level groups. Categorizes as "Low", "Medium", or "High"
+sleepdata$PhysicalActivityGroup <- cut(sleepdata$Physical.Activity.Level, breaks = c(0, 30, 60, Inf), labels = c("Low", "Medium", "High"))
+
+# Box plot for Quality of Sleep by Physical Activity Level Group
+ggplot(sleepdata, aes(x = PhysicalActivityGroup, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Physical Activity Level Group", x = "Physical Activity Level Group", y = "Quality of Sleep")
+```
+
+We can clearly see that Physical Activity level has a positive correlation with Quality of Sleep.
+We'll now explore Daily Steps.
+We can create a new variable with groups according to the guidelines here at [10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500).
+Nobody in the dataset goes over 10000 steps a day, however, so we won't see any data for the "Active" or "Highly Active" groups.
+
+```{r}
+# Histogram for Daily Steps
+ggplot(sleepdata, aes(x = Daily.Steps)) +
+  geom_histogram(binwidth = 500) +
+  labs(title = "Histogram of Daily Steps", x = "Daily Steps", y = "Count")
+
+# Create a new variable for Daily Steps groups based off of 10,000 steps.org categories
+sleepdata$StepsGroup <- cut(sleepdata$Daily.Steps, 
+                       breaks = c(-Inf, 5000, 7500, 10000, 12500, Inf), 
+                       labels = c("Sedentary", "Low active", "Somewhat active", "Active", "Highly active"), 
+                       include.lowest = TRUE)
+
+# Box plot for Quality of Sleep by Daily Steps Group
+ggplot(sleepdata, aes(x = StepsGroup, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Daily Steps Group", x = "Daily Steps Group", y = "Quality of Sleep")
+
+```
+
+Daily Steps don't have as clear of a trend as Physical Activity did, but there do seem to be fewer low Quality of Sleep scores in higher daily steps groups.
+Now we'll explore our Sleep Duration variable.
+According to [Medical News Today](https://www.medicalnewstoday.com/articles/sleep-cycle-stages#sleep-requirements), adults aged 18-61 are recommended 7+ hours of sleep a night.
+We'll keep this in mind when looking at our data.
+
+```{r}
+# Histogram for Sleep Duration
+ggplot(sleepdata, aes(x = Sleep.Duration)) +
+  geom_histogram(binwidth = 0.5) +
+  labs(title = "Histogram of Sleep Duration", x = "Sleep Duration (hours)", y = "Count")
+
+# Scatterplot for Quality of Sleep by Sleep Duration
+ggplot(sleepdata, aes(x = Sleep.Duration, y = Quality.of.Sleep)) +
+  geom_point() +
+  geom_smooth(method = "loess") +
+  labs(title = "Quality of Sleep by Sleep Duration", x = "Sleep Duration (hours)", y = "Quality of Sleep")
+
+```
+
+We can see a clear positive correlation between Sleep Duration and Quality of Sleep.
+This is to be expected.
+Now we'll begin plotting our final set of variables.
+We'll start with analyzing the new variable we created with categories of Blood Pressure.
+
+```{r}
+# Bar plot for cat_bp
+ggplot(sleepdata, aes(x = cat_bp)) +
+  geom_bar() +
+  labs(title = "Bar Plot of cat_bp", x = "cat_bp", y = "Count")
+
+# Box plot for Quality of Sleep by cat_bp
+ggplot(sleepdata, aes(x = cat_bp, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by cat_bp", x = "cat_bp", y = "Quality of Sleep")
+
+```
+The boxplots look a bit strange for the Elevated and Normal levels, but we can see on the histogram that it's likely due to a shortage of data for these categories.
+We can see, however, that Quality of Sleep does seem to be negatively correlated with increasing stages of hypertension.
+Now we'll explore BMI with some plots.
+
+```{r}
+# Bar plot for BMI Category
+ggplot(sleepdata, aes(x = BMI.Category)) +
+  geom_bar() +
+  labs(title = "Bar Plot of BMI Category", x = "BMI Category", y = "Count")
+
+# Box plot for Quality of Sleep by BMI Category
+ggplot(sleepdata, aes(x = BMI.Category, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by BMI Category", x = "BMI Category", y = "Quality of Sleep")
+
+```
+We can clearly see that Normal weights seem to have better sleep on average than the obese and overweight categories.
+However, there are very few in the obese category, so conclusions with that group may not be as supported as others.
+Now we'll look at Stress Level.
+
+```{r}
+# Histogram for Stress Level
+ggplot(sleepdata, aes(x = Stress.Level)) +
+  geom_histogram(binwidth = 1) +
+  labs(title = "Histogram of Stress Level", x = "Stress Level (scale: 1-10)", y = "Count")
+
+# Scatterplot for Quality of Sleep by Stress Level
+ggplot(sleepdata, aes(x = Stress.Level, y = Quality.of.Sleep)) +
+  geom_point() +
+  geom_smooth(method = "loess") +
+  labs(title = "Quality of Sleep by Stress Level", x = "Stress Level (scale: 1-10)", y = "Quality of Sleep")
+```
+Our scatterplot shows a clear negative correlation between Stress Level and Quality of Sleep.
+However, this is a subjective scale, so data should be taken with a grain of salt for now.
+Let's look at Heart Rate related to Quality of Sleep now.
+
+```{r}
+# Histogram for Heart Rate
+ggplot(sleepdata, aes(x = Heart.Rate)) +
+  geom_histogram(binwidth = 5) +
+  labs(title = "Histogram of Heart Rate", x = "Heart Rate (bpm)", y = "Count")
+
+# Scatterplot for Quality of Sleep by Heart Rate
+ggplot(sleepdata, aes(x = Heart.Rate, y = Quality.of.Sleep)) +
+  geom_point() +
+  geom_smooth(method = "loess") +
+  labs(title = "Quality of Sleep by Heart Rate", x = "Heart Rate (bpm)", y = "Quality of Sleep")
+
+```
+
+This data is a bit difficult to visualize for now but we can see a negative correlation between the two.
+Healthy Heart Rates tend to differ between each individual, however, so we'll have to wait until deeper statistical analyses to draw any educated conclusions for now.
+We've finally reached our last variable: Sleep Disorders.
+
+```{r}
+# Bar plot for Sleep Disorder
+ggplot(sleepdata, aes(x = Sleep.Disorder)) +
+  geom_bar() +
+  labs(title = "Bar Plot of Sleep Disorder", x = "Sleep Disorder", y = "Count")
+
+# Box plot for Quality of Sleep by Sleep Disorder
+ggplot(sleepdata, aes(x = Sleep.Disorder, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by Sleep Disorder", x = "Sleep Disorder", y = "Quality of Sleep")
+```
+We can see that individuals without a disorder tend to have higher sleep scores than those with Insomnia or Sleep Apnea, which is what we'd expect to see.
+The dataset also seems to reflect the population accurately, as those without a sleep disorder are more numerous than individuals with one.

### 2024-02-22 products/manuscript/Manuscript.qmd
Kevin- Part 2 for Project
EDA plots and updated Manuscript document to include variables of interest.

@@
 End of Project Part 1
+
+Project Part 2:
+
+We'll include the data cleaning here in order to generate our plots in this document. Please see the eda.qmd file for full exploratory results and the cleaning process.
+
+```{r}
+#load cleaning tools
+library(tidyverse)
+# Splitting the Blood.Pressure variable into systolic and diastolic variable columns
+sleepdata <- sleepdata %>%
+  separate(Blood.Pressure, into = c("systolic", "diastolic"), sep = "/")
+
+# Ensuring that the new columns are considered numeric 
+sleepdata$systolic <- as.numeric(sleepdata$systolic)
+sleepdata$diastolic <- as.numeric(sleepdata$diastolic)
+
+# Creating new categorical variable
+
+sleepdata <- sleepdata %>%
+  mutate(cat_bp = case_when(
+    systolic < 120 & diastolic < 80 ~ "Normal",
+    systolic >= 120 & systolic <= 129 & diastolic < 80 ~ "Elevated",
+    (systolic >= 130 & systolic <= 139) | (diastolic >= 80 & diastolic <= 89) ~ "Stage 1 Hypertension",
+    systolic >= 140 | diastolic >= 90 ~ "Stage 2 Hypertension"
+  ))
+# Merging "Normal" and "Normal Weight" attributes into a singular "Normal" category
+
+sleepdata <- sleepdata %>%
+  mutate(BMI.Category = case_when(
+    BMI.Category %in% c("Normal", "Normal Weight") ~ "Normal",
+    BMI.Category == "Overweight" ~ "Overweight",
+    BMI.Category == "Obese" ~ "Obese",
+  ))
+```
+
+Here are the most interesting/distinct plots from the EDA:
+
+Physical Activity plot:
+
+We can clearly see that Physical Activity level has a positive correlation with Quality of Sleep.
+```{r}
+#load plot package
+library(ggplot2)
+
+# Create a new variable for Physical Activity Level groups. Categorizes as "Low", "Medium", or "High"
+sleepdata$PhysicalActivityGroup <- cut(sleepdata$Physical.Activity.Level, breaks = c(0, 30, 60, Inf), labels = c("Low", "Medium", "High"))
+
+# Box plot for Quality of Sleep by Physical Activity Level Group
+ggplot(sleepdata, aes(x = PhysicalActivityGroup, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Physical Activity Level Group", x = "Physical Activity Level Group", y = "Quality of Sleep")
+```
+
+Gender, Age, and Occupation:
+
+We can see some clear differences in self-reported Quality of Sleep between these different groups already.
+Females tend to report an average of 1 higher according to the boxplot.
+The different age groups are similar, but the oldest (50-60) reports the highest quality of sleep by far.
+Our occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
+```{r}
+# Boxplot for Quality of Sleep by Gender
+ggplot(sleepdata, aes(x = Gender, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Gender", x = "Gender", y = "Quality of Sleep")
+
+# Create bins for Age in order to create a clean boxplot.
+sleepdata$AgeGroup <- cut(sleepdata$Age, breaks = c(20, 30, 40, 50, 60), labels = c("20-30", "30-40", "40-50", "50-60"), include.lowest = TRUE)
+
+# Calculate mean Quality of Sleep for each Age Group to plot with means for cleaner visuals
+sleepdata_summary <- sleepdata %>%
+  group_by(AgeGroup) %>%
+  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
+
+# Bar plot for Mean Quality of Sleep by Age Group with colorblind-friendly palette
+ggplot(sleepdata_summary, aes(x = AgeGroup, y = MeanQuality, fill = AgeGroup)) +
+  geom_bar(stat = "identity") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(title = "Mean Quality of Sleep by Age Group", x = "Age Group", y = "Mean Quality of Sleep")
+
+# Calculate mean Quality of Sleep for each Occupation in order to create a bar plot.
+sleepdata_summary <- sleepdata %>%
+  group_by(Occupation) %>%
+  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
+
+# Bar plot for Mean Quality of Sleep by Occupation with manually specified colors
+ggplot(sleepdata_summary, aes(x = Occupation, y = MeanQuality, fill = Occupation)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = rainbow(length(unique(sleepdata_summary$Occupation)))) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Mean Quality of Sleep by Occupation", x = "Occupation", y = "Mean Quality of Sleep")
+```
+
+BMI:
+We can clearly see that Normal weights seem to have better sleep on average than the obese and overweight categories.
+However, there are very few in the obese category, so conclusions with that group may not be as supported as others.
+
+```{r}
+# Box plot for Quality of Sleep by BMI Category
+ggplot(sleepdata, aes(x = BMI.Category, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by BMI Category", x = "BMI Category", y = "Quality of Sleep")
+```
+
+Blood Pressure:
+
+The boxplots look a bit strange for the Elevated and Normal levels, but we can see that it's likely due to a shortage of data for these categories.
+We can see, however, that Quality of Sleep does seem to be negatively correlated with increasing stages of hypertension.
+
+```{r}
+# Bar plot for cat_bp
+ggplot(sleepdata, aes(x = cat_bp)) +
+  geom_bar() +
+  labs(title = "Bar Plot of cat_bp", x = "cat_bp", y = "Count")
+
+# Box plot for Quality of Sleep by cat_bp
+ggplot(sleepdata, aes(x = cat_bp, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by cat_bp", x = "cat_bp", y = "Quality of Sleep")
+
+```
+
+And Stress Level:
+Our scatterplot shows a clear negative correlation between Stress Level and Quality of Sleep. However, this is a subjective scale, so data should be taken with a grain of salt for now.
+
+```{r}
+# Scatterplot for Quality of Sleep by Stress Level
+ggplot(sleepdata, aes(x = Stress.Level, y = Quality.of.Sleep)) +
+  geom_point() +
+  geom_smooth(method = "loess") +
+  labs(title = "Quality of Sleep by Stress Level", x = "Stress Level (scale: 1-10)", y = "Quality of Sleep")
+```
+
+The other plots were either self explanatory (e.g. Sleep Duration meant higher sleep quality) or had less defined trends (e.g. heart rate and daily steps).
+
+End of Project Part 2
 

### 2024-02-23 R/analysis-code/statistical-analysis.R
Emma - Removed unnecessary files

@@ -4,6 +4,8 @@
 #this script loads the processed, cleaned data, does a simple analysis
 #and saves the results to the results folder
 
+## THIS FILE WILL BE USED FOR FUTURE STATISTICAL ANALYSIS OF SLEEP DATA 
+
 #load needed packages. make sure they are installed.
 library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()

### 2024-02-23 products/manuscript/Manuscript.qmd
Emma - Part 2
Restructured the Manuscript file to flow. Inserted comments on what still needs to be inserted.

@@
 
+# Summary/Abstract
+
+  ** NEED TO INSERT **
+
+{{< pagebreak >}}
+
+# Introduction
+
+## General Background Information
+
+  ** NEED TO INSERT **
+
+## Description of data and data source
+
 Project Part 1: This dataset was obtained from Kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data) and was created by Mark Otto and Andrew Fong at MIT.
@@
 
+## Questions/Hypotheses to be addressed
+
 We want to explore the influence of each variable on sleep quality and identify which ones seem to have the most impact (both positive and negative).
@@
 
-End of Project Part 1
+{{< pagebreak >}}
 
-Project Part 2:
+# Methods
+
+  ** NEED TO INSERT **
+
+## Data import and cleaning
 
@@
 ```
+
+## Statistical analysis
+
+  ** WILL BE PERFORMED IN THE STATISTICAL ANALYSIS FILE **
+
+{{< pagebreak >}}
+
+# Results
+
+## Exploratory/Descriptive analysis
 
@@
 
-End of Project Part 2
-
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
-
-# Summary/Abstract
-_Write a summary of your project._
-
-{{< pagebreak >}}
-
-# Introduction
-
-## General Background Information
-_Provide enough background on your topic that others can understand the why and how of your analysis_
-
-## Description of data and data source
-_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
-
-## Questions/Hypotheses to be addressed
-_State the research questions you plan to answer with this analysis._
-
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
-
-{{< pagebreak >}}
-
-# Methods
-
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
-
-## Schematic of workflow
-
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-#knitr::include_graphics(here("products","antigen-recognition.png"))
-```
-
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
-
-## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
-
-## Statistical analysis
-_Explain anything related to your statistical analyses._
-
-{{< pagebreak >}}
-
-# Results
-
-## Exploratory/Descriptive analysis
-
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
- shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-#resulttable=readRDS("../../results/summarytable.rds")
-#knitr::kable(resulttable)
-```
+End of Project Part 2.
 
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-#knitr::include_graphics(here("results","height-weight-stratified.png"))
-```
-
 ## Full analysis
-
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Exampleshows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-#resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
-#knitr::kable(resulttable2)
-```
 
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
 
 ## Conclusions
-_What are the main take-home messages?_
-
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper  discusses types of analyses.
-
-These papers  are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
 

### 2024-03-12 R/analysis-code/statistical-analysis.qmd
3/12- emma stat analysis

---
title: "statistical-analysis.qmd"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2024-03-12 R/analysis-code/statistical-analysis.R

@@ -1,58 +1,80 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
+---
+#### Exploratory Data Analysis: Project Part 2
+#### Kevin Kosewick & Emma Hardin-Parker
 
-## THIS FILE WILL BE USED FOR FUTURE STATISTICAL ANALYSIS OF SLEEP DATA 
+# Loading Necessary Packages 
+library(ggplot2) 
+library(broom) 
+library(here) 
+library(tidymodels)
+library(tidyverse)
 
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
+# Writing new dataframe, sleepdata, into a separate excel file to ensure reproducibility 
 
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
+### i'm gonna need your help on this part, Kevin I tried and couldn't figure out the best way to do it ####
 
-#load data. 
-mydata <- readRDS(data_location)
 
+### For now, I'm just going to use the sleepdata frame we already have in the environment
 
 ######################################
-#Data fitting/statistical analysis
+# Fitting Models and Statistical Analysis # 
 ######################################
 
 ############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
+#### Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor
 
-lmfit1 <- lm(Height ~ Weight, mydata)  
+lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)  
 
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
+# Placing results from lmfit_quality_activty  into a data frame with the tidy function
+lmtable_quality_activity <- broom::tidy(lmfit_quality_activity)
 
-#look at fit results
-print(lmtable1)
+# Viewing the results from the first model fit 
+print(lmtable_quality_activity)
 
-# save fit results table  
-table_file1 = here("results", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
+# Saving the lmfit_quality_activity results table  
+lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
[…]

### 2024-03-14 products/manuscript/Manuscript.qmd
Added 5 new models to the statistical analysis file. Cleaned folders.
Added 5 models to the statistical analysis script. Cleaned all of the ReadMe files, organized things where they should go, and deleted residual template bits that we don't need.

@@
 
-```{r}
+```{r Package Loading, warning=FALSE, message=FALSE}
 #load in the data and check it out
 library(here)
-sleepdata <- read.csv(here("data","raw-data", "ss.csv"))
-summary(sleepdata)
-structure(sleepdata)
-class(sleepdata)
+library(tidyverse)
+library(dplyr)
+library(skimr)
 ```
@@
 
-## Data import and cleaning
+## Data import
 
-We'll include the data cleaning here in order to generate our plots in this document. Please see the eda.qmd file for full exploratory results and the cleaning process.
+We'll load in the cleaned data here. Please see the eda.qmd file for full exploratory results and the cleaning process.
 
-```{r}
-#load cleaning tools
-library(tidyverse)
-# Splitting the Blood.Pressure variable into systolic and diastolic variable columns
-sleepdata <- sleepdata %>%
-  separate(Blood.Pressure, into = c("systolic", "diastolic"), sep = "/")
+```{r Loading the Processed Data}
+#load the cleaned data
+sleepdata<- readRDS(here("data","processed-data", "sleepdataprocessed.rds"))
 
-# Ensuring that the new columns are considered numeric 
-sleepdata$systolic <- as.numeric(sleepdata$systolic)
-sleepdata$diastolic <- as.numeric(sleepdata$diastolic)
-
-# Creating new categorical variable
-
-sleepdata <- sleepdata %>%
-  mutate(cat_bp = case_when(
-    systolic < 120 & diastolic < 80 ~ "Normal",
-    systolic >= 120 & systolic <= 129 & diastolic < 80 ~ "Elevated",
-    (systolic >= 130 & systolic <= 139) | (diastolic >= 80 & diastolic <= 89) ~ "Stage 1 Hypertension",
-    systolic >= 140 | diastolic >= 90 ~ "Stage 2 Hypertension"
-  ))
-# Merging "Normal" and "Normal Weight" attributes into a singular "Normal" category
-
-sleepdata <- sleepdata %>%
-  mutate(BMI.Category = case_when(
-    BMI.Category %in% c("Normal", "Normal Weight") ~ "Normal",
-    BMI.Category == "Overweight" ~ "Overweight",
-    BMI.Category == "Obese" ~ "Obese",
-  ))
 ```
@@
 We can clearly see that Physical Activity level has a positive correlation with Quality of Sleep.
-```{r}
+```{r Physical Activity Plot, warning=FALSE, message=FALSE}
 #load plot package
 library(ggplot2)
-
-# Create a new variable for Physical Activity Level groups. Categorizes as "Low", "Medium", or "High"
-sleepdata$PhysicalActivityGroup <- cut(sleepdata$Physical.Activity.Level, breaks = c(0, 30, 60, Inf), labels = c("Low", "Medium", "High"))
 
@@
 Our occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
-```{r}
+```{r Gender, Age, and Occupation Plot}
 # Boxplot for Quality of Sleep by Gender
@@
 
-```{r}
+```{r BMI plot}
 # Box plot for Quality of Sleep by BMI Category
@@
 
-```{r}
+```{r Blood Pressure Plot}
 # Bar plot for cat_bp
@@
 
-```{r}
+```{r Stress Level Plot}
 # Scatterplot for Quality of Sleep by Stress Level
@@
 ## Basic statistical analysis
+
+**Included in the analysis-code folder  for now**
 

### 2024-03-14 products/manuscript/supplement/Supplementary-Material.qmd
Added 5 new models to the statistical analysis file. Cleaned folders.
Added 5 models to the statistical analysis script. Cleaned all of the ReadMe files, organized things where they should go, and deleted residual template bits that we don't need.

---
title: "Supplement to Example Manuscript Template for a Data Analysis Project"

subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis_template_references.bib
csl: ../../assets/vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

For now, this document is a placeholder with the original text that came with the template for this project. All text beyond here is not relevant to our project in particular and will be updated when we have supplementary materials to include.

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.

# Overview

A quick overview of what readers can find in the supplement.

# Code and file information

Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
Essentially, give a full set of instructions to re-generate everything.

{{< pagebreak >}}

# Additional Method Details

Often, the main manuscript only allows for an overview description of the methods. Use the supplement to describe all your methods, models and approaches in a lot of detail. Reference specific parts of your code as needed.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight.png"))
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

### 2024-03-14 R/analysis-code/statistical-analysis.R

@@ -9,19 +9,15 @@ library(here)
 library(tidymodels)
 library(tidyverse)
 
-# Writing new dataframe, sleepdata, into a separate excel file to ensure reproducibility 
-
-### i'm gonna need your help on this part, Kevin I tried and couldn't figure out the best way to do it ####
-
-
-### For now, I'm just going to use the sleepdata frame we already have in the environment
-
 ######################################
 # Fitting Models and Statistical Analysis # 
 ######################################
 
 ############################
 #### Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor
+
+#Load the data
+sleepdata<- readRDS(here("data","processed-data","sleepdataprocessed.RDS"))
 
 lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)  
 
@@ -36,7 +32,9 @@ lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
 saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)
 
 
-      ## include interpretation 
+      ## The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0. This has a relatively low standard error and high significance.
+## The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109. Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's.
+## This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.
 
 ############################
 #### Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
@@ -53,14 +51,15 @@ print(lmtable_quality_duration)
 lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")
 saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
 
-      ## include interpretation 
+      ## For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep. This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high. 
+## For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1. This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.
 
 ############################
[…]

### 2024-03-14 products/manuscript/Manuscript.qmd
Updated Manuscript
Added models of interest

@@
 
-  ** WILL BE PERFORMED IN THE STATISTICAL ANALYSIS FILE **
+We'll begin to look at some of the relationships between our variables now. Sleep Quality is our outcome of interest for all of these. We'll start with Phyiscal Activity level.
+
+```{r}
+############################
+#### Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor
+
+lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)
+
+# Placing results from lmfit_quality_activty  into a data frame with the tidy function
+lmtable_quality_activity <- broom::tidy(lmfit_quality_activity)
+
+# Viewing the results from the first model fit 
+print(lmtable_quality_activity)
+
+# Saving the lmfit_quality_activity results table  
+lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
+saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)
+```
+
+The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0.
+This has a relatively low standard error and high significance.The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109.
+Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's.This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.
+
+Next, we'll look at Sleep Duration as a predictor.
+
+```{r}
+############################
+#### Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
+
+lmfit_quality_duration <- lm(Quality.of.Sleep ~ Sleep.Duration, sleepdata)
+
+# Placing results from lmfit_quality_duration into a data frame with the tidy function
+lmtable_quality_duration <- broom::tidy(lmfit_quality_duration)
+
+#Viewing the results from the second model fit 
+print(lmtable_quality_duration)
+
+# Saving the lmfit_quality_duration results table  
+lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")
+saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
+
+```
+
+For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep.
+This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high.
+For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1.
+This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.
+
+Next, we'll look at Stress Level as our predictor.
+
+```{r}
+############################
+####Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.
+
+lmfit_quality_stress <- lm(Quality.of.Sleep~ Stress.Level, sleepdata)
+
+#Placing results from lmfit_quality_stress into a data frame with the tidy function
+lmtable_quality_stress <- broom::tidy(lmfit_quality_stress)
+
+#Viewing the results from the sixth model fit 
+print(lmtable_quality_stress)
+
+# Saving the lmfit_quality_stress results table  
+lmtable_quality_stress = here("results", "tables", "lmfit6table.rds")
+saveRDS(lmtable_quality_stress, file = lmtable_quality_stress)
+```
+
+According to our model, Stress has a strongly defined relationship with sleep quality.
+We can see that a value of 0 on the stress score leads to a 10 unit increase in quality of sleep.
+The Stress Level coefficient indicates that as stress increases by one unit, sleep quality decreases by about 0.6 of a unit.
+The t-statistics and p-values for the values in this model are extremely strong, indicating a well-defined relationship between stress and sleep quality.
+
+Finally, we'll look at Age as our predictor.
+
+```{r}
+############################
+####Fitting a model using Sleep Quality as the outcome and Age as a predictor.
+
+lmfit_quality_age <- lm(Quality.of.Sleep~ Age, sleepdata)
+
+#Placing results from lmfit_quality_gender_age into a data frame with the tidy function
+lmtable_quality_age <- broom::tidy(lmfit_quality_age)
+
+#Viewing the results from the seventh model fit 
+print(lmtable_quality_age)
+
+# Saving the lmfit_quality_gender_age results table  
+lmtable_quality_age = here("results", "tables", "lmfit8table.rds")
+saveRDS(lmtable_quality_age, file = lmtable_quality_age)
+
+```
+
+Our model shows a strong relationship between age and quality of sleep.
+As Age increases, sleep quality seems to improve.
+Each year increase in age is predicted to have a 0.065 increase in sleep quality score.
+The intercept here is also interesting as an age of "0" is predicted to have a sleep score of 4.57, which is relatively poor.
+The t-statistics and p-values for these are both strong.
 

### 2024-03-15 R/analysis-code/statistical-analysis.R
emma- part 3

@@ -9,6 +9,11 @@ library(here)
 library(tidymodels)
 library(tidyverse)
 
+<<<<<<< Updated upstream
+=======
+# Writing new dataframe, sleepdata, into a separate excel file to ensure reproducibility 
+
+>>>>>>> Stashed changes
 ######################################
 # Fitting Models and Statistical Analysis # 
 ######################################
@@ -32,10 +37,13 @@ lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
 saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)
 
 
+<<<<<<< Updated upstream
       ## The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0. This has a relatively low standard error and high significance.
 ## The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109. Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's.
 ## This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.
 
+=======
+>>>>>>> Stashed changes
 ############################
 #### Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
 
@@ -51,8 +59,11 @@ print(lmtable_quality_duration)
 lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")
 saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
 
+<<<<<<< Updated upstream
       ## For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep. This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high. 
 ## For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1. This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.
+=======
+>>>>>>> Stashed changes
 
 ############################
 #### Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors
@@ -69,6 +80,7 @@ print(lmtable_quality_duration_occupation)
 lmtable_quality_duration_occupation = here("results", "tables", "lmfit3table.rds")
 saveRDS(lmtable_quality_duration_occupation, file = lmtable_quality_duration_occupation)
 
+<<<<<<< Updated upstream
[…]

### 2024-04-03 R/analysis-code/statistical-analysis.R
Statistical analysis document updates
Grouped occupation variable. Fit a Random Forest model with CV and visualized/interpreted the results. Created a null model and compared the results between the two models.

@@ -3,17 +3,18 @@
 #### Kevin Kosewick & Emma Hardin-Parker
 
 # Loading Necessary Packages 
-library(ggplot2) 
+
 library(broom) 
 library(here) 
 library(tidymodels)
 library(tidyverse)
+library(corrplot)
+library(ranger)
+library(vip)
+library(ggplot2)
 
-<<<<<<< Updated upstream
-=======
 # Writing new dataframe, sleepdata, into a separate excel file to ensure reproducibility 
 
->>>>>>> Stashed changes
 ######################################
 # Fitting Models and Statistical Analysis # 
 ######################################
@@ -36,14 +37,10 @@ print(lmtable_quality_activity)
 lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
 saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)
 
-
-<<<<<<< Updated upstream
       ## The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0. This has a relatively low standard error and high significance.
 ## The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109. Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's.
 ## This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.
 
-=======
->>>>>>> Stashed changes
 ############################
 #### Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
 
@@ -59,11 +56,8 @@ print(lmtable_quality_duration)
 lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")
 saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
 
-<<<<<<< Updated upstream
       ## For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep. This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high. 
 ## For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1. This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.
-=======
->>>>>>> Stashed changes
 
 ############################
 #### Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors
[…]

### 2024-04-04 R/eda-processing-code/eda.qmd
Added 5 new models to the statistical analysis file. Cleaned folders.
Added 5 models to the statistical analysis script. Cleaned all of the ReadMe files, organized things where they should go, and deleted residual template bits that we don't need.
4/3- emma
Updated manuscript, supplementary file, working on finalizing readMe files, etc. I left many comments throughout so we can know what's still left to be done for Part 5!

--- 2024-02-22 R/eda-code/eda.qmd
@@
 The dataset also seems to reflect the population accurately, as those without a sleep disorder are more numerous than individuals with one.
+
+Finally, we'll save the completely processed data into a new file in the "processed_data" folder.
+```{r}
+#quickly change gender from a character value to a factor
+as.factor(sleepdata$Gender)
+
+#save the processed data separately from the original
+saveRDS(sleepdata, file = "sleepdataprocessed.rds")
+```
+
+Distribution of the outcome of interest, Quality of Sleep
+
+```{r}
+sleep_dis <- ggplot(sleepdata, aes(x = Quality.of.Sleep)) +
+  geom_histogram() +
+  stat_bin(bins = 15) +
+  theme_minimal()
+print(sleep_dis)
+
+summary(sleepdata$Quality.of.Sleep)
+
+```

### 2024-04-04 products/manuscript/Manuscript.qmd
4/3- emma
Updated manuscript, supplementary file, working on finalizing readMe files, etc. I left many comments throughout so we can know what's still left to be done for Part 5!

@@
 ---
-title: "Manuscript for MADA Project"
+title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
 subtitle: ""
@@
 
-# Summary/Abstract
+*ANYTHING BELOW CODE LINE 172 IS A DRAFT AND IS NOT TO BE CONSIDERED A PART OF THE FINAL MANUSCRIPT*
 
-  ** NEED TO INSERT **
+# Summary/Abstract (1)
+
+  *NEED TO INSERT*
+
+  *WILL NEED COLLAB FOR PART 5*
+
+# Introduction (2)
+
+## Background Information
+
+  *NEED TO INSERT *
+
+  *EMMA'S SOURCES ARE LINKED HERE:*
+
+  https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8345503/ --> blood pressure and sleep
+  https://www.apa.org/news/press/releases/stress/2013/sleep --> sleep duration & stress on sleep quality 
+  https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10503965/#:~:text=Regular%20physical%20activity%20can%20lead,managing%20sleep%20disorders%20like%20insomnia. --> physical activity and sleep quality
 
@@
 
-# Introduction
+# Materials and Methods (3)
 
-## General Background Information
+## Data and Processing
 
-  ** NEED TO INSERT **
+The data were obtained from kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
+The dataset was curated by Mark Otto and Andrew Fong at MIT.
+The dataset originally contained 373 observations for 13 variables covering a broad spectrum health and lifestyle variables associated with sleep quality.
+The curators of the dataset did not provide information on how these data were collected, so we solely relied on the codebook for variable definitions.
+There were no missing values or erroneous variables in the dataset that we needed to remove.
+The Blood.Pressure variable was transformed to reflect the systolic and diastolic blood pressure measurements from each subject.
+We then chose to add an additional variable, labeled 'cat_bp' to reflect the categorical blood pressure status of a subject based on the American Heart Association guidelines *INCLUDE REFERENCE HERE (can be found in eda.qmd file)* .
+Additionally we created a variable, 'StepsGroup' to determine the categorical activity level of subjects in the dataset based on these guidelines provided by 10,000 steps *[10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500).* Finally we created an additional categorical variable, 'PhysicalActivityGroup' based on 30 minute increments, to categorically represent the level of physical activity reported by a subject.
+The outcome of interest is noted as "Quality.of.Sleep" that we will reference as Sleep Quality or Quality of Sleep throughout this report.
 
-## Description of data and data source
-
-Project Part 1: This dataset was obtained from Kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data) and was created by Mark Otto and Andrew Fong at MIT.
-The set has 373 observations for 13 variables covering a broad spectrum associated with sleep and daily routine.
-
+*Unsure if we need this to be here? Or if it will even show up given the FALSE arguments we added*
 ```{r Package Loading, warning=FALSE, message=FALSE}
@@
 
+## Variables included in Analyses *Include Questions / Hypotheses Here*
+
+Given that there were only 13 original variables, and 17 variables after final data transformation; simple linear regression models were performed to determine baseline associations between variables and Sleep Quality, the outcome of interest.
+These regression models were created to determine which variables had the largest impact on Sleep Quality and drove the rest of the analyses.
+Simple linear regression models were fitted for the following variables: BMI, cat_bp, Stress Level, Gender, Age, and Occupation.
+
+The Occupation variable was then transformed to group the various occupations into the following groups: Healthcare, Education, Engineering, Business/Finance, and Science.
+Prior to performing any subsequent analyses, we removed the original versions of the variables we transformed: systolic, diastolic, daily steps, occupation, Physical.Activity.Level, Heart.Rate, and AgeGroup.
+
+## Model Development
+
+*Include LASSO to have more than one ML model?*
+
+Random Forest Models and Cross-Validation were considered to be the "best" fit for our outcome of interest, Quality of Sleep.
+The data were not split into train/test subsets as the data contains less than 400 observations and many of the values were unique in comparison to the rest of the data.
+Prior to constructing this model, a collinearity plot was constructed to determine any presence of collinearity in our data.
+*INCLUDE PLOT HERE* Stress and Sleep Quality had an absolute value of 0.9 on the correlation scale.
+Sleep Duration and Stress were also seen to be strongly correlated.
+Stress was still included in our models; however, we found it important to note the strong linear relationship between Stress and Sleep Quality.
+
+A variable importance plot (VIP) was constructed to determine which variables had the strongest impact in our Random Forest Model.
+*INCLUDE PLOT HERE* Sleep Duration, Stress Level, and Age had the highest levels of importance in our model.
+Physical Activity Level was excluded entirely, which was to be expected given our exploratory data analysis and simple linear regression models.
+
+## Defining the Models
+
+*DEFINE DECISION TREES. DEFINE BAGGING AND BOOTSTRAPPING. DEFINE RANDOM FOREST MODELS* 
+*DEFINE CROSS-VALIDATION* 
+*DEFINE IMPORTANCE OF NULL MODEL TO COMPARE RESULTS*
+
+*INCLUDE CITATIONS OF CHAPTER 9,10,11 OF HMLR LINKED IN "MANY TREE MODELS" SECTION OF CONTENT FOLDER 11*
+
+## Evaluation of Models
+
+Root Mean Squared Error (RMSE) is a common metric used to evaluate regression models.
+The RMSE is formally defined as the square root of the mean square of all error and is defined by the following formula: *Include LaTeX equation of RMSE?* .
+It is important that while RMSE is scale dependent, common practice notes that low RMSE values indicate stronger model performance.
+
+## Software Used for Analyses
+
+This analysis was conducted under R version 4.3.1 on a MacOS operating system *cite R*.
+The following R packages were used in the development of these analyses: *R PACKAGES USED: here, skimr, broom, tidyverse, ggplot2, dplyr, corrplot, ranger, vip*
+All processing and analysis code can be found in the Supplementary Material file.
+
+# Results (4)
+
+## Outcomes of Interest
+
+The 'Quality of Sleep' variable has been selected as our outcome of interest.
+It is important to note that this variable is a subjective measure of a participant's self-reported sleep quality.
+The following plot shows that most of subjects reported a sleep quality score of 8 (out of 10).
+*REFERENCE SLEEP_DIS PLOT FROM EDA.QMD FILE*
+
+## Exploratory Findings
+
+## Machine Learning Models
+
+A variable importance plot (VIP) was constructed to determine which variables had the strongest impact in our Random Forest Model.
+*INCLUDE PLOT HERE* Sleep Duration, Stress Level, and Age had the highest levels of importance in our model.
+Physical Activity Level was excluded entirely, which was to be expected given our exploratory data analysis and simple linear regression models.
+
+*Include comparison of RF Model and null model*
+
+# Discussion
+
+*PART 5*
+
+## Summary
+
+*PART 5*
+
+## Strengths and Limitations
+
+*PART 5*
+
+*Include that all parameters were self-reported and thus subjected to bias*
+
+## Conclusion
+
+*PART 5*
+
+{{< pagebreak >}}
+
+# References
+
+*INCLUDE INCREMENTALLY -- FINALIZE*
+
+*ANYTHING BELOW THIS COMMENT IS FROM PRIOR TO THE OVERHAUL AND IS STILL TO BE IGNORED / MOVED TO ANOTHER SECTION OF THE PROJECT*
+
 ## Questions/Hypotheses to be addressed
 
-We want to explore the influence of each variable on sleep quality and identify which ones seem to have the most impact (both positive and negative).
+We intend to explore the influence of each variable on sleep quality and identify which ones seem to have the most impact (both positive and negative).
 We also wish to explore if certain variables affect male sleep quality more than females.
@@
 
-# Methods
-
-  ** NEED TO INSERT **
-
 ## Data import
 
-We'll load in the cleaned data here. Please see the eda.qmd file for full exploratory results and the cleaning process.
+We'll load in the cleaned data here. Please reference the eda.qmd and Supplementary-Material.qmd files for full exploratory results and the cleaning process.
 
@@
 
-We'll begin to look at some of the relationships between our variables now. Sleep Quality is our outcome of interest for all of these. We'll start with Phyiscal Activity level.
+We'll begin to look at some of the relationships between our variables now. Sleep Quality is our outcome of interest for all of these. We'll start with Physical Activity level.
 
@@
 
+## Outcomes of Interest
+
 ## Exploratory/Descriptive analysis
@@
 
-End of Project Part 2.
-
 ## Basic statistical analysis

### 2024-04-04 products/manuscript/supplement/Supplementary-Material.qmd
4/3- emma
Updated manuscript, supplementary file, working on finalizing readMe files, etc. I left many comments throughout so we can know what's still left to be done for Part 5!

@@
 ---
-title: "Supplement to Example Manuscript Template for a Data Analysis Project"
+title: "Supplementary Material"
 
@@
 
-For now, this document is a placeholder with the original text that came with the template for this project.
-All text beyond here is not relevant to our project in particular and will be updated when we have supplementary materials to include.
+*This shows some materials that could go into a supplementary file.
+Often you want/need references here too.
+You can use the same reference bib file for this and the main text (as done here) or have separate bib files.*
 
-This shows some materials that could go into a supplementary file.
-Often you want/need references here too.
-You can use the same reference bib file for this and the main text (as done here) or have separate bib files.
-
-For illustrative purposes, I'm doing the supplement as pdf.
+*For illustrative purposes, I'm doing the supplement as pdf.
 For this to work, you need a (La)TeX system installed.
 It's easy.
-Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).
+Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).*
 
-Of course you would choose the format based on needs.
-
-I'm also using a different style for the references here.
+*I'm also using a different style for the references here.
 (vancouver vs apa in the main manuscript).
 Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header.
-We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.
+We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.*
+
+### I HAVE STRUCTURED THIS FILE. NEED TO WORK ON PULLING CERTAIN FIGURES, PLOTS, TABLES, ETC. ###
 
@@
 
-A quick overview of what readers can find in the supplement.
+This document includes additional information about the data and results.
+The full code and detailed explanation of the figures can be found at the [GitHub Repository](https://github.com/KKosewick/KOSEWICK-HARDIN-PARKER-MADA-project).
 
-# Code and file information
+# Code & File Details and Locations
 
-Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
-Essentially, give a full set of instructions to re-generate everything.
+The eda.qmd file, located in the R folder and subsequently in the eda-processing-code folder contains all data exploration, processing, and transformation.
+The statistical-analysis.R file, found in the R folder and subsequently in the analysis-code folder holds all statistical analysis and machine learning code.
 
@@
 
-# Additional Method Details
+# Exploratory Data Analysis
 
-Often, the main manuscript only allows for an overview description of the methods.
-Use the supplement to describe all your methods, models and approaches in a lot of detail.
-Reference specific parts of your code as needed.
+*Include ALL exploratory analysis figures*
+
+The data set contains 373 observations for 13 variables.
+Some variables are not easily understandable at face value, so we went back to the data source and accessed the data dictionary.
+The variables are defined as follows:
+
+Person ID: A unique identifier for each participant.
+Gender: The participant's gender (Male/Female).
+Age: The participant's age in years.
+Occupation: The job or career of the participant.
+Sleep Duration (hours): The daily sleep duration of the participant in hours.
+Quality of Sleep (scale: 1-10): A subjective assessment of sleep quality on a scale from 1 to 10.
+Physical Activity Level (minutes/day): The daily duration of physical activity for the participant, measured in minutes.
+Stress Level (scale: 1-10): A subjective assessment of the participant's stress level on a scale from 1 to 10.
+BMI Category: The participant's BMI classification (e.g., Underweight, Normal, Overweight).
+Blood Pressure (systolic/diastolic): The participant's blood pressure, represented as systolic pressure over diastolic pressure.
+Heart Rate (bpm): The participant's resting heart rate, measured in beats per minute.
+Daily Steps: The number of steps the participant takes each day.
+Sleep Disorder: The presence or absence of a sleep disorder in the participant (None, Insomnia, Sleep Apnea).
+
+Before going further into the exploratory analysis, we need to determine whether missing values are present and if they could jeopardize any future analyses.
+
+```{r}
+print(sum(is.na(sleepdata)))
+```
+
+There are no missing values in our data set!
+We can now move on to creating new variables or transmuting existing variables.
+
+# Data Transformation
+
+We are going to create a categorical variable to go alongside the existing blood pressure variable.
+We are using blood pressure guidelines from the American Heart Association to create our new categorical blood pressure variable.
+Our goal with this is to see how blood pressure can affect sleep quality in a more digestible way.
+First, we will need to create new systolic and diastolic variables to ultimately create our categorical variable.
+We have attached the guidelines here:
+
+[![AHA Guidelines](AHA_Guidelines.png)](https://www.heart.org/en/health-topics/high-blood-pressure)
+
+```{r}
+# Splitting the Blood.Pressure variable into systolic and diastolic variable columns
+sleepdata <- sleepdata %>%
+  separate(Blood.Pressure, into = c("systolic", "diastolic"), sep = "/")
+
+# Ensuring that the new columns are considered numeric 
+sleepdata$systolic <- as.numeric(sleepdata$systolic)
+sleepdata$diastolic <- as.numeric(sleepdata$diastolic)
+
+# Creating new categorical variable
+
+sleepdata <- sleepdata %>%
+  mutate(cat_bp = case_when(
+    systolic < 120 & diastolic < 80 ~ "Normal",
+    systolic >= 120 & systolic <= 129 & diastolic < 80 ~ "Elevated",
+    (systolic >= 130 & systolic <= 139) | (diastolic >= 80 & diastolic <= 89) ~ "Stage 1 Hypertension",
+    systolic >= 140 | diastolic >= 90 ~ "Stage 2 Hypertension"
+  ))
+```
+
+Now that we have created this new variable and adjusted the format of the blood pressure variables, we are now going to further explore the other variables to see if they are fit, as is, for analysis.
+We noticed that within the BMI.Category variable, there are both "Normal" and "Normal Weight" categories.
+We assume this was an oversight in data entry, and intend to merge the two together to ensure all "Normal" categories are together for subsequent analyses.
+
+```{r}
+# Merging "Normal" and "Normal Weight" attributes into a singular "Normal" category
+
+sleepdata <- sleepdata %>%
+  mutate(BMI.Category = case_when(
+    BMI.Category %in% c("Normal", "Normal Weight") ~ "Normal",
+    BMI.Category == "Overweight" ~ "Overweight",
+    BMI.Category == "Obese" ~ "Obese",
+  ))
+```
+
+As of this point in our cleaning, we feel confident that the data is structured in such a way that analyses can be performed.
+
+# Data Visualization and Exploratory Data Analysis
+
+We will now begin to create some figures to help visualize the relationships in the dataset.
+Quality of Sleep is our outcome of interest; we'll explore the effect each other variable seems to have on it.
+First, we'll start with the variables that aren't health specific.
+I asked Microsoft Copilot in Precise Mode to generate the base code for some plots that I then changed them to be specific to the dataset/have more visual appeal.
+
+```{r}
+# Boxplot for Quality of Sleep by Gender
+ggplot(sleepdata, aes(x = Gender, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Gender", x = "Gender", y = "Quality of Sleep")
+
+# Create bins for Age in order to create a clean boxplot.
+sleepdata$AgeGroup <- cut(sleepdata$Age, breaks = c(20, 30, 40, 50, 60), labels = c("20-30", "30-40", "40-50", "50-60"), include.lowest = TRUE)
+
+# Calculate mean Quality of Sleep for each Age Group to plot with means for cleaner visuals
+sleepdata_summary <- sleepdata %>%
+  group_by(AgeGroup) %>%
+  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
+
+# Bar plot for Mean Quality of Sleep by Age Group with colorblind-friendly palette
+ggplot(sleepdata_summary, aes(x = AgeGroup, y = MeanQuality, fill = AgeGroup)) +
+  geom_bar(stat = "identity") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(title = "Mean Quality of Sleep by Age Group", x = "Age Group", y = "Mean Quality of Sleep")
+
+# Calculate mean Quality of Sleep for each Occupation in order to create a bar plot.
+sleepdata_summary <- sleepdata %>%
+  group_by(Occupation) %>%
+  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
+
+# Bar plot for Mean Quality of Sleep by Occupation with manually specified colors
+ggplot(sleepdata_summary, aes(x = Occupation, y = MeanQuality, fill = Occupation)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = rainbow(length(unique(sleepdata_summary$Occupation)))) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Mean Quality of Sleep by Occupation", x = "Occupation", y = "Mean Quality of Sleep")
+
+```
+
+We can see some clear differences in self-reported Quality of Sleep between these different groups already.
+Females tend to report an average of 1 higher according to the boxplot.
+The different age groups are similar, but the oldest (50-60) reports the highest quality of sleep by far.
+Our occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
+We'll now explore Physical Activity and plot it as a predictor for Quality of Sleep.
+To make it easier to visualize Physical Activity, we'll create a new variable with categories of Low (0-30 minutes), Medium (30-60), and High (60+).
+We'll again use Microsoft Copilot for the basic plot code and modify it to our dataset/preferences.
+
+```{r}
+# Histogram for Physical Activity Level
+ggplot(sleepdata, aes(x = Physical.Activity.Level)) +
+  geom_histogram(binwidth = 10) +
+  labs(title = "Histogram of Physical Activity Level", x = "Physical Activity Level (minutes/day)", y = "Count")
+
+# Create a new variable for Physical Activity Level groups. Categorizes as "Low", "Medium", or "High"
+sleepdata$PhysicalActivityGroup <- cut(sleepdata$Physical.Activity.Level, breaks = c(0, 30, 60, Inf), labels = c("Low", "Medium", "High"))
+
+# Box plot for Quality of Sleep by Physical Activity Level Group
+ggplot(sleepdata, aes(x = PhysicalActivityGroup, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Physical Activity Level Group", x = "Physical Activity Level Group", y = "Quality of Sleep")
+```
+
+We can clearly see that Physical Activity level has a positive correlation with Quality of Sleep.
+We'll now explore Daily Steps.
+We can create a new variable with groups according to the guidelines here at [10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500).
+Nobody in the dataset goes over 10000 steps a day, however, so we won't see any data for the "Active" or "Highly Active" groups.
+
+```{r}
+# Histogram for Daily Steps
+ggplot(sleepdata, aes(x = Daily.Steps)) +
+  geom_histogram(binwidth = 500) +
+  labs(title = "Histogram of Daily Steps", x = "Daily Steps", y = "Count")
+
+# Create a new variable for Daily Steps groups based off of 10,000 steps.org categories
+sleepdata$StepsGroup <- cut(sleepdata$Daily.Steps, 
+                       breaks = c(-Inf, 5000, 7500, 10000, 12500, Inf), 
+                       labels = c("Sedentary", "Low active", "Somewhat active", "Active", "Highly active"), 
+                       include.lowest = TRUE)
+
+# Box plot for Quality of Sleep by Daily Steps Group
+ggplot(sleepdata, aes(x = StepsGroup, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Quality of Sleep by Daily Steps Group", x = "Daily Steps Group", y = "Quality of Sleep")
+
+```
+
+Daily Steps don't have as clear of a trend as Physical Activity did, but there do seem to be fewer low Quality of Sleep scores in higher daily steps groups.
+Now we'll explore our Sleep Duration variable.
+According to [Medical News Today](https://www.medicalnewstoday.com/articles/sleep-cycle-stages#sleep-requirements), adults aged 18-61 are recommended 7+ hours of sleep a night.
+We'll keep this in mind when looking at our data.
+
+```{r}
+# Histogram for Sleep Duration
+ggplot(sleepdata, aes(x = Sleep.Duration)) +
+  geom_histogram(binwidth = 0.5) +
+  labs(title = "Histogram of Sleep Duration", x = "Sleep Duration (hours)", y = "Count")
+
+# Scatterplot for Quality of Sleep by Sleep Duration
+ggplot(sleepdata, aes(x = Sleep.Duration, y = Quality.of.Sleep)) +
+  geom_point() +
+  geom_smooth(method = "loess") +
+  labs(title = "Quality of Sleep by Sleep Duration", x = "Sleep Duration (hours)", y = "Quality of Sleep")
+
+```
+
+We can see a clear positive correlation between Sleep Duration and Quality of Sleep.
+This is to be expected.
+Now we'll begin plotting our final set of variables.
+We'll start with analyzing the new variable we created with categories of Blood Pressure.
+
+```{r}
+# Bar plot for cat_bp
+ggplot(sleepdata, aes(x = cat_bp)) +
+  geom_bar() +
+  labs(title = "Bar Plot of cat_bp", x = "cat_bp", y = "Count")
+
+# Box plot for Quality of Sleep by cat_bp
+ggplot(sleepdata, aes(x = cat_bp, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by cat_bp", x = "cat_bp", y = "Quality of Sleep")
+
+```
+The boxplots look a bit strange for the Elevated and Normal levels, but we can see on the histogram that it's likely due to a shortage of data for these categories.
+We can see, however, that Quality of Sleep does seem to be negatively correlated with increasing stages of hypertension.
+Now we'll explore BMI with some plots.
+
+```{r}
+# Bar plot for BMI Category
+ggplot(sleepdata, aes(x = BMI.Category)) +
+  geom_bar() +
+  labs(title = "Bar Plot of BMI Category", x = "BMI Category", y = "Count")
+
+# Box plot for Quality of Sleep by BMI Category
+ggplot(sleepdata, aes(x = BMI.Category, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by BMI Category", x = "BMI Category", y = "Quality of Sleep")
+
+```
+We can clearly see that Normal weights seem to have better sleep on average than the obese and overweight categories.
+However, there are very few in the obese category, so conclusions with that group may not be as supported as others.
+Now we'll look at Stress Level.
+
+```{r}
+# Histogram for Stress Level
+ggplot(sleepdata, aes(x = Stress.Level)) +
+  geom_histogram(binwidth = 1) +
+  labs(title = "Histogram of Stress Level", x = "Stress Level (scale: 1-10)", y = "Count")
+
+# Scatterplot for Quality of Sleep by Stress Level
+ggplot(sleepdata, aes(x = Stress.Level, y = Quality.of.Sleep)) +
+  geom_point() +
+  geom_smooth(method = "loess") +
+  labs(title = "Quality of Sleep by Stress Level", x = "Stress Level (scale: 1-10)", y = "Quality of Sleep")
+```
+Our scatterplot shows a clear negative correlation between Stress Level and Quality of Sleep.
+However, this is a subjective scale, so data should be taken with a grain of salt for now.
+Let's look at Heart Rate related to Quality of Sleep now.
+
+```{r}
+# Histogram for Heart Rate
+ggplot(sleepdata, aes(x = Heart.Rate)) +
+  geom_histogram(binwidth = 5) +
+  labs(title = "Histogram of Heart Rate", x = "Heart Rate (bpm)", y = "Count")
+
+# Scatterplot for Quality of Sleep by Heart Rate
+ggplot(sleepdata, aes(x = Heart.Rate, y = Quality.of.Sleep)) +
+  geom_point() +
+  geom_smooth(method = "loess") +
+  labs(title = "Quality of Sleep by Heart Rate", x = "Heart Rate (bpm)", y = "Quality of Sleep")
+
+```
+
+This data is a bit difficult to visualize for now but we can see a negative correlation between the two.
+Healthy Heart Rates tend to differ between each individual, however, so we'll have to wait until deeper statistical analyses to draw any educated conclusions for now.
+We've finally reached our last variable: Sleep Disorders.
+
+```{r}
+# Bar plot for Sleep Disorder
+ggplot(sleepdata, aes(x = Sleep.Disorder)) +
+  geom_bar() +
+  labs(title = "Bar Plot of Sleep Disorder", x = "Sleep Disorder", y = "Count")
+
+# Box plot for Quality of Sleep by Sleep Disorder
+ggplot(sleepdata, aes(x = Sleep.Disorder, y = Quality.of.Sleep)) +
+  geom_boxplot() +
+  labs(title = "Box Plot of Quality of Sleep by Sleep Disorder", x = "Sleep Disorder", y = "Quality of Sleep")
+```
+We can see that individuals without a disorder tend to have higher sleep scores than those with Insomnia or Sleep Apnea, which is what we'd expect to see.
+The dataset also seems to reflect the population accurately, as those without a sleep disorder are more numerous than individuals with one.
 
@@
 
-# Additional results
+# Simple Linear Regression and Basic Statistical Analysis Figures / Results
 
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
+Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor
 
-## Example additional result
+```{r}
+#Load the data
+sleepdata<- readRDS(here("data","processed-data","sleepdataprocessed.RDS"))
 
-@tbl-resulttable1 shows an additional table summarizing a model fit.
+lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)
+
+# Placing results from lmfit_quality_activty  into a data frame with the tidy function
+lmtable_quality_activity <- broom::tidy(lmfit_quality_activity)
+
+# Viewing the results from the first model fit 
+print(lmtable_quality_activity)
+
+# Saving the lmfit_quality_activity results table  
+lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
+saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)
+```
+
+The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0. This has a relatively low standard error and high significance.
+## The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109.
+Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's.
+## This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.
+
+Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
+```{r}
+lmfit_quality_duration <- lm(Quality.of.Sleep ~ Sleep.Duration, sleepdata)
+
+# Placing results from lmfit_quality_duration into a data frame with the tidy function
+lmtable_quality_duration <- broom::tidy(lmfit_quality_duration)
+
+#Viewing the results from the second model fit 
+print(lmtable_quality_duration)
+
+# Saving the lmfit_quality_duration results table  
+lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")
+saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
+```
+
+For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep.
+This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high.
+For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1.
+This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.
+
+Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors
+
+```{r}
+lmfit_quality_duration_occupation <- lm(Quality.of.Sleep ~ Sleep.Duration + Occupation, sleepdata)
+
+# Placing results from lmfit_quality_duration_occupation into a data frame with the tidy function
+lmtable_quality_duration_occupation <- broom::tidy(lmfit_quality_duration_occupation)
+
+#Viewing the results from the third model fit 
+print(lmtable_quality_duration_occupation)
+
+# Saving the lmfit_quality_duration_occupation results table  
+lmtable_quality_duration_occupation = here("results", "tables", "lmfit3table.rds")
+saveRDS(lmtable_quality_duration_occupation, file = lmtable_quality_duration_occupation)
+```
+
+Since Occupation is a categorical variable, interpreting our intercept value is tricky.
+It's hard to get good information from it since we don't have any data from unemployed people.
+However, the coefficients for our predictors show a variety of interesting relationships.
+Sleep Duration shows a similar relationship to our bivariate model we ran earlier.
+Sales representatives have the worst sleep quality by far with a strong t-statistic and p-value.
+Scientists, doctors, and sales persons are all about the next worst (with strong t-stats and p-values).
+The occupations that seem to have the best sleep are teachers, lawyers, engineers and nurses.
+This is somewhat surprising since these are all demanding jobs that can have decently stressful work environments.
+
+Fitting a model using Sleep Quality as the outcome and BMI as a predictor.
+
+```{r}
+lmfit_quality_BMI <- lm(Quality.of.Sleep~ BMI.Category, sleepdata)
+
+#Placing results from lmfit_quality_BMI into a data frame with the tidy function
+lmtable_quality_BMI <- broom::tidy(lmfit_quality_BMI)
+
+#Viewing the results from the fourth model fit 
+print(lmtable_quality_BMI)
+
+# Saving the lmfit_quality_BMI results table  
+lmtable_quality_BMI = here("results", "tables", "lmfit4table.rds")
+saveRDS(lmtable_quality_BMI, file = lmtable_quality_BMI)
+```
+
+We can see that Obese individuals have a stronger negative relationship with Sleep Quality than Overweight individuals.
+The lower t-statistics and p-values are likely due to fewer observations for Obese individuals in the dataset.
+However, it is all still statistically significant and has reasonable standard errors.
+
+Fitting a model using Sleep Quality as the outcome and Blood Pressure category as a predictor.
+
+```{r}
+lmfit_quality_BP <- lm(Quality.of.Sleep~ cat_bp, sleepdata)
+
+#Placing results from lmfit_quality_BP into a data frame with the tidy function
+lmtable_quality_BP <- broom::tidy(lmfit_quality_BP)
+
+#Viewing the results from the fifth model fit 
+print(lmtable_quality_BP)
+
+# Saving the lmfit_quality_BP results table  
+lmtable_quality_BP = here("results", "tables", "lmfit5table.rds")
+saveRDS(lmtable_quality_BP, file = lmtable_quality_BP)
+```
+
+We have some interesting results for this one.
+Our values aren't statistically significant according to p-values, but we can see that Stage 1 Hypertension seems to have the biggest negative impact on Quality of Sleep.
+This is surprising since we would expect more severe hypertension to result in worse sleep, but according to our model, Stage 2 hypertension has less impact than Stage 2.
+Normal blood pressure seems to have a small positive impact on sleep quality according to our model.
+All of these predictors seem to have a relatively small impact as they impact the sleep quality scale by less than one whole unit.
+Given our p-values and low t-statistics, however, this model should be taken with a grain of salt.
+
+Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.
+
+```{r}
+lmfit_quality_stress <- lm(Quality.of.Sleep~ Stress.Level, sleepdata)
+
+#Placing results from lmfit_quality_stress into a data frame with the tidy function
+lmtable_quality_stress <- broom::tidy(lmfit_quality_stress)
+
+#Viewing the results from the sixth model fit 
+print(lmtable_quality_stress)
+
+# Saving the lmfit_quality_stress results table  
+lmtable_quality_stress = here("results", "tables", "lmfit6table.rds")
+saveRDS(lmtable_quality_stress, file = lmtable_quality_stress)
+```
+
+According to our model, Stress has a strongly defined relationship with sleep quality. We can see that a value of 0 on the stress score leads to a 10 unit increase in quality of sleep. 
+## The Stress Level coefficient indicates that as stress increases by one unit, sleep quality decreases by about 0.6 of a unit.
+The t-statistics and p-values for the values in this model are extremely strong, indicating a well-defined relationship between stress and sleep quality.
+
+Fitting a model using Sleep Quality as the outcome and Age and Gender as predictors.
+
+```{r}
+lmfit_quality_gender_age <- lm(Quality.of.Sleep~ Gender + Age, sleepdata)
+
+#Placing results from lmfit_quality_gender_age into a data frame with the tidy function
+lmtable_quality_gender_age <- broom::tidy(lmfit_quality_gender_age)
+
+#Viewing the results from the seventh model fit 
+print(lmtable_quality_gender_age)
+
+# Saving the lmfit_quality_gender_age results table  
+lmtable_quality_gender_age = here("results", "tables", "lmfit7table.rds")
+saveRDS(lmtable_quality_gender_age, file = lmtable_quality_gender_age)
+```
+
+According to our table, Males seem to have lower sleep quality than females by about a score of .0336.
+However, the t-statistic is very low and the p-value is very poor.
+Gender doesn't seem to be a good predictor of sleep quality.
+## When age is factored with gender, we can see that quality of sleep seems to increase slightly along with age.
+The t-statistic and p-value are both strong for this predictor; it may be worth doing a simple bivariate analysis of age.
+
+Fitting a model using Sleep Quality as the outcome and Age as a predictor.
+
+```{r}
+lmfit_quality_age <- lm(Quality.of.Sleep~ Age, sleepdata)
+
+#Placing results from lmfit_quality_age into a data frame with the tidy function
+lmtable_quality_age <- broom::tidy(lmfit_quality_age)
+
+#Viewing the results from the eighth model fit 
+print(lmtable_quality_age)
+
+# Saving the lmfit_quality_age results table  
+lmtable_quality_age = here("results", "tables", "lmfit8table.rds")
+saveRDS(lmtable_quality_age, file = lmtable_quality_age)
+```
+
+Our model shows a strong relationship between age and quality of sleep.
+As Age increases, sleep quality seems to improve.
+Each year increase in age is predicted to have a 0.065 increase in sleep quality score.
+## The intercept here is also interesting as an age of "0" is predicted to have a sleep score of 4.57, which is relatively poor. The t-statistics and p-values for these are both strong.
+
+Now we'll create a new variable that groups the different occupations to make analysis more succinct.
+
+```{r}
+# Create a new column
+sleepdata$Occupation_Group <- as.character(sleepdata$Occupation)
+
+# Map occupations to broader categories
+sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Nurse', 'Doctor')] <- 'Healthcare'
+sleepdata$Occupation_Group[sleepdata$Occupation == 'Teacher'] <- 'Education'
+sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Software Engineer', 'Engineer')] <- 'Engineering'
+sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Accountant', 'Salesperson', 'Sales Representative', 'Manager')] <- 'Business/Finance'
+sleepdata$Occupation_Group[sleepdata$Occupation == 'Scientist'] <- 'Science'
+
+# Convert the new column to a factor
+sleepdata$Occupation_Group <- as.factor(sleepdata$Occupation_Group)
+
+## This should help with analysis of this variable. We'll fit this to a model now to test it.
+
+lmfit_quality_occupation <- lm(Quality.of.Sleep~ Occupation_Group, sleepdata)
+
+#Placing results from lmfit_quality_occupation into a data frame with the tidy function
+lmtable_quality_occupation <- broom::tidy(lmfit_quality_occupation)
+
+#Viewing the results from the occupation model fit 
+print(lmtable_quality_occupation)
+
+# Saving the lmfit_quality_gender_age results table  
+lmtable_quality_occupation = here("results", "tables", "lmfit9table.rds")
+saveRDS(lmtable_quality_occupation, file = lmtable_quality_occupation)
+```
+
+*Example additional result*
+
+*@tbl-resulttable1 shows an additional table summarizing a model fit.*
 
@@
 
-# Discussion
+*# Discussion*
 

### 2024-04-04 products/manuscript/Manuscript.qmd
Cleaning folders and updating background section
Removed unnecessary files/folders and did some manuscript updating.

@@
 
+  The data was obtained from kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
+It contains data from a study on individuals of different sexes and age regarding their sleep quality and different variables that may affect this.
+We are unsure of where this data set came from and how it was collected; the publisher on Kaggle has yet to respond to a comment with this question posted on March 5th 2024.
+Regardless, it contains interesting variables that have been shown in previous studies to have an effect on sleep quality.
+
+  For instance, in Sun et al.
+(2015), a sample of Chinese individuals showed a correlation between obesity and worsened sleep quality in men but not women.
+We are interested in seeing if this trend holds up for our dataset; given that most of the observations are of men, we expect to see a correlation between lower sleep scores and obesity in our data.
+Furthermore, it has been shown that age and gender have a notable effect on sleep quality (Madrid-Valero et al.
+2016).
+Women seem to experience a deterioration in sleep quality as they age; however, while this trend is still present in men, it is notably less consistent and can vary dramatically between individuals.
+We expect to see similar trends in our dataset, but as stated previously, our observations contain significantly more men than women.
+Finally, we wish to test whether or not certain occupational industries have different effects on sleep quality.
+It has been shown that managerial positions tend to have the worst sleep quality among civilian sector workers while 24 jobs that have rotating shifts tend to have the worst sleep quality (Luckhaupt et al.
+2010).
+Our dataset contains observations of mostly white-collar/service-based jobs, so we expect to find similar trends upon analysis.
+
   *NEED TO INSERT *
@@
 
-The data were obtained from kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
-The dataset was curated by Mark Otto and Andrew Fong at MIT.
 The dataset originally contained 373 observations for 13 variables covering a broad spectrum health and lifestyle variables associated with sleep quality.
-The curators of the dataset did not provide information on how these data were collected, so we solely relied on the codebook for variable definitions.
+The curators of the dataset did not provide information on how or where this data was collected, so we solely relied on the codebook for variable definitions.
 There were no missing values or erroneous variables in the dataset that we needed to remove.
-The Blood.Pressure variable was transformed to reflect the systolic and diastolic blood pressure measurements from each subject.
+The "Blood.Pressure" variable was transformed to reflect the systolic and diastolic blood pressure measurements from each subject.
 We then chose to add an additional variable, labeled 'cat_bp' to reflect the categorical blood pressure status of a subject based on the American Heart Association guidelines *INCLUDE REFERENCE HERE (can be found in eda.qmd file)* .
-Additionally we created a variable, 'StepsGroup' to determine the categorical activity level of subjects in the dataset based on these guidelines provided by 10,000 steps *[10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500).* Finally we created an additional categorical variable, 'PhysicalActivityGroup' based on 30 minute increments, to categorically represent the level of physical activity reported by a subject.
+Additionally we created a variable, 'StepsGroup' to determine the categorical activity level of subjects in the dataset based on these guidelines provided by 10,000 steps *[10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500).* Finally, we created an additional categorical variable, 'PhysicalActivityGroup' based on 30 minute increments, to categorically represent the level of physical activity reported by a subject.
 The outcome of interest is noted as "Quality.of.Sleep" that we will reference as Sleep Quality or Quality of Sleep throughout this report.
@@
 ## Model Development
-
-*Include LASSO to have more than one ML model?*
 
@@
 It is important to note that this variable is a subjective measure of a participant's self-reported sleep quality.
-The following plot shows that most of subjects reported a sleep quality score of 8 (out of 10).
+The following plot shows that most subjects reported a sleep quality score of 8 (out of 10).
 *REFERENCE SLEEP_DIS PLOT FROM EDA.QMD FILE*
@@
 *INCLUDE INCREMENTALLY -- FINALIZE*
+
+  https://link.springer.com/article/10.1007/s11325-015-1193-z --> obesity and sleep quality (affects men but not women in China)
+
+  https://www.scielosp.org/article/gs/2017.v31n1/18-22/en/?uid=680b54f39b# --> effect of age and gender on sleep quality (women worse than men, sleep quality deteriorates with age in women but is a bit all over the place for men)
+
+  https://academic.oup.com/sleep/article/33/2/149/2454438#77919376 --> sleep quality by occupation/industry (managerial positions had the worst score)
 
@@
 
-# Results
-
-## Outcomes of Interest
-
 ## Exploratory/Descriptive analysis
@@
 The other plots were either self explanatory (e.g. Sleep Duration meant higher sleep quality) or had less defined trends (e.g. heart rate and daily steps).
-
-## Basic statistical analysis
-
-**Included in the analysis-code folder  for now**
-
-## Full analysis
-
-{{< pagebreak >}}
-
-# Discussion
-
-## Summary and Interpretation
-
-## Strengths and Limitations
-
-## Conclusions
-
-{{< pagebreak >}}
-
-# References

### 2024-04-16 products/manuscript/Manuscript.qmd
Manuscript Updates
Added figures, equations, an abstract, and updated some of the sections.

@@
 bibliography: ../manuscript/dataanalysis-references.bib
-
+editor: 
+  markdown: 
+    wrap: 72
 ---
-
-*ANYTHING BELOW CODE LINE 172 IS A DRAFT AND IS NOT TO BE CONSIDERED A PART OF THE FINAL MANUSCRIPT*
 
@@
 
-  *NEED TO INSERT*
+For our project, we obtained a dataset from Kaggle.com containing 373
+observations of 13 variables from a study on different factors that may
+affect sleep quality. Variables of interest from the study included
+physical activity level, age, sex, occupation, stress, and several more.
+In RStudio (version 4.3), we used a combination of packages to create
+models suitable for answering our primary question: which variables are
+most important for determining an individual's sleep quality and what
+effect do they have? After conducting an exploratory data analysis (EDA)
+we hypothesized that stress, sleep duration, and occupation would have
+the largest impact on sleep quality. We believed that stress and sleep
+quality would be negatively correlated, sleep duration and quality would
+be positively correlated, and that jobs with longer or more variable
+hours would negatively affect sleep quality.
 
-  *WILL NEED COLLAB FOR PART 5*
+After creating linear models for each variable's impact on sleep quality
+and fitting a Random Forest model with Cross-Validation to test variable
+importance, we determined that sleep duration, stress level, and age
+were the most important factors. Our hypothesis for sleep duration and
+stress was supported; however, we did not correctly predict that age
+would be one of the most important predictors for sleep quality. Our
+hypotheses regarding the effects of sleep duration, stress, and
+occupation were all supported as well. Overall, we determined that an
+individual's duration of sleep, stress level, and age are all key
+factors that can impact sleep quality.
+
+*Feel free to edit this Emma; I tried to include everything needed*
 
@@
 
-  The data was obtained from kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
-It contains data from a study on individuals of different sexes and age regarding their sleep quality and different variables that may affect this.
-We are unsure of where this data set came from and how it was collected; the publisher on Kaggle has yet to respond to a comment with this question posted on March 5th 2024.
-Regardless, it contains interesting variables that have been shown in previous studies to have an effect on sleep quality.
+The data was obtained from kaggle.com at this
+[link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
+It contains data from a study on individuals of different sexes and age
+regarding their sleep quality and different variables that may affect
+this. We are unsure of where this data set came from and how it was
+collected; the publisher on Kaggle has yet to respond to a comment with
+this question posted on March 5th 2024. Regardless, it contains
+interesting variables that have been shown in previous studies to have
+an effect on sleep quality.
 
-  For instance, in Sun et al.
-(2015), a sample of Chinese individuals showed a correlation between obesity and worsened sleep quality in men but not women.
-We are interested in seeing if this trend holds up for our dataset; given that most of the observations are of men, we expect to see a correlation between lower sleep scores and obesity in our data.
-Furthermore, it has been shown that age and gender have a notable effect on sleep quality (Madrid-Valero et al.
-2016).
-Women seem to experience a deterioration in sleep quality as they age; however, while this trend is still present in men, it is notably less consistent and can vary dramatically between individuals.
-We expect to see similar trends in our dataset, but as stated previously, our observations contain significantly more men than women.
-Finally, we wish to test whether or not certain occupational industries have different effects on sleep quality.
-It has been shown that managerial positions tend to have the worst sleep quality among civilian sector workers while 24 jobs that have rotating shifts tend to have the worst sleep quality (Luckhaupt et al.
-2010).
-Our dataset contains observations of mostly white-collar/service-based jobs, so we expect to find similar trends upon analysis.
+For instance, in Sun et al. (2015), a sample of Chinese individuals
+showed a correlation between obesity and worsened sleep quality in men
+but not women. We are interested in seeing if this trend holds up for
+our dataset; given that most of the observations are of men, we expect
+to see a correlation between lower sleep scores and obesity in our data.
+Furthermore, it has been shown that age and gender have a notable effect
+on sleep quality (Madrid-Valero et al. 2016). Women seem to experience a
+deterioration in sleep quality as they age; however, while this trend is
+still present in men, it is notably less consistent and can vary
+dramatically between individuals. We expect to see similar trends in our
+dataset, but as stated previously, our observations contain
+significantly more men than women. Finally, we wish to test whether or
+not certain occupational industries have different effects on sleep
+quality. It has been shown that managerial positions tend to have the
+worst sleep quality among civilian sector workers while 24 jobs that
+have rotating shifts tend to have the worst sleep quality (Luckhaupt et
+al. 2010). Our dataset contains observations of mostly
+white-collar/service-based jobs, so we expect to find similar trends
+upon analysis.
 
-  *NEED TO INSERT *
+*NEED TO INSERT*
 
@@
 
-  https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8345503/ --> blood pressure and sleep
-  https://www.apa.org/news/press/releases/stress/2013/sleep --> sleep duration & stress on sleep quality 
-  https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10503965/#:~:text=Regular%20physical%20activity%20can%20lead,managing%20sleep%20disorders%20like%20insomnia. --> physical activity and sleep quality
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8345503/ --\> blood
+pressure and sleep
+https://www.apa.org/news/press/releases/stress/2013/sleep --\> sleep
+duration & stress on sleep quality
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10503965/#:\~:text=Regular%20physical%20activity%20can%20lead,managing%20sleep%20disorders%20like%20insomnia.
+--\> physical activity and sleep quality
 
@@
 
-The dataset originally contained 373 observations for 13 variables covering a broad spectrum health and lifestyle variables associated with sleep quality.
-The curators of the dataset did not provide information on how or where this data was collected, so we solely relied on the codebook for variable definitions.
-There were no missing values or erroneous variables in the dataset that we needed to remove.
-The "Blood.Pressure" variable was transformed to reflect the systolic and diastolic blood pressure measurements from each subject.
-We then chose to add an additional variable, labeled 'cat_bp' to reflect the categorical blood pressure status of a subject based on the American Heart Association guidelines *INCLUDE REFERENCE HERE (can be found in eda.qmd file)* .
-Additionally we created a variable, 'StepsGroup' to determine the categorical activity level of subjects in the dataset based on these guidelines provided by 10,000 steps *[10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500).* Finally, we created an additional categorical variable, 'PhysicalActivityGroup' based on 30 minute increments, to categorically represent the level of physical activity reported by a subject.
-The outcome of interest is noted as "Quality.of.Sleep" that we will reference as Sleep Quality or Quality of Sleep throughout this report.
+The dataset originally contained 373 observations for 13 variables
+covering a broad spectrum health and lifestyle variables associated with
+sleep quality. The curators of the dataset did not provide information
+on how or where this data was collected, so we solely relied on the
+codebook for variable definitions. There were no missing values or
+erroneous variables in the dataset that we needed to remove. The
+"Blood.Pressure" variable was transformed to reflect the systolic and
+diastolic blood pressure measurements from each subject. We then chose
+to add an additional variable, labeled 'cat_bp' to reflect the
+categorical blood pressure status of a subject based on the [*American
+Heart Association
+guidelines*](https://www.heart.org/en/health-topics/high-blood-pressure).
+Additionally we created a variable, 'StepsGroup' to determine the
+categorical activity level of subjects in the dataset based on these
+guidelines provided by 10,000 steps [*10,000
+steps.org*](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)*.*
+Finally, we created an additional categorical variable,
+'PhysicalActivityGroup' based on 30 minute increments, to categorically
+represent the level of physical activity reported by a subject. The
+outcome of interest is noted as "Quality.of.Sleep" that we will
+reference as Sleep Quality or Quality of Sleep throughout this report.
 
-*Unsure if we need this to be here? Or if it will even show up given the FALSE arguments we added*
-```{r Package Loading, warning=FALSE, message=FALSE}
-#load in the data and check it out
-library(here)
-library(tidyverse)
-library(dplyr)
-library(skimr)
-```
+## Variables included in Analyses
 
-## Variables included in Analyses *Include Questions / Hypotheses Here*
+Given that there were only 13 original variables, and 17 variables after
+final data transformation; simple linear regression models were
+performed to determine baseline associations between variables and Sleep
+Quality, the outcome of interest. These regression models were created
+to determine which variables had the largest impact on Sleep Quality and
+drove the rest of the analyses. Simple linear regression models were
+fitted for the following variables: BMI, cat_bp, Stress Level, Gender,
+Age, and Occupation.
 
-Given that there were only 13 original variables, and 17 variables after final data transformation; simple linear regression models were performed to determine baseline associations between variables and Sleep Quality, the outcome of interest.
-These regression models were created to determine which variables had the largest impact on Sleep Quality and drove the rest of the analyses.
-Simple linear regression models were fitted for the following variables: BMI, cat_bp, Stress Level, Gender, Age, and Occupation.
+The Occupation variable was then transformed to group the various
+occupations into the following groups: Healthcare, Education,
+Engineering, Business/Finance, and Science. Prior to performing any
+subsequent analyses, we removed the original versions of the variables
+we transformed: systolic, diastolic, daily steps, occupation,
+Physical.Activity.Level, Heart.Rate, and AgeGroup.
 
-The Occupation variable was then transformed to group the various occupations into the following groups: Healthcare, Education, Engineering, Business/Finance, and Science.
-Prior to performing any subsequent analyses, we removed the original versions of the variables we transformed: systolic, diastolic, daily steps, occupation, Physical.Activity.Level, Heart.Rate, and AgeGroup.
+The objective of our analysis was to determine which of these variables
+are most important for determining an individual's sleep quality and
+what specific effect they have. After conducting an (EDA) we
+hypothesized that stress, sleep duration, and occupation would have the
+largest impact on sleep quality. We believed that stress and sleep
+quality would be negatively correlated, sleep duration and quality would
+be positively correlated, and that jobs with longer or more variable
+hours would negatively affect sleep quality.
 
@@
 
-Random Forest Models and Cross-Validation were considered to be the "best" fit for our outcome of interest, Quality of Sleep.
-The data were not split into train/test subsets as the data contains less than 400 observations and many of the values were unique in comparison to the rest of the data.
-Prior to constructing this model, a collinearity plot was constructed to determine any presence of collinearity in our data.
-*INCLUDE PLOT HERE* Stress and Sleep Quality had an absolute value of 0.9 on the correlation scale.
-Sleep Duration and Stress were also seen to be strongly correlated.
-Stress was still included in our models; however, we found it important to note the strong linear relationship between Stress and Sleep Quality.
+Random Forest Models and Cross-Validation were considered to be the
+"best" fit for our outcome of interest, Quality of Sleep. The data were
+not split into train/test subsets as the data contains less than 400
+observations and many of the values were unique in comparison to the
+rest of the data. Prior to constructing this model, a collinearity plot
+was constructed to determine any presence of collinearity in our data.
+![Collinearity between all variables.](images/collinearity.png)
 
-A variable importance plot (VIP) was constructed to determine which variables had the strongest impact in our Random Forest Model.
-*INCLUDE PLOT HERE* Sleep Duration, Stress Level, and Age had the highest levels of importance in our model.
-Physical Activity Level was excluded entirely, which was to be expected given our exploratory data analysis and simple linear regression models.
+Stress and Sleep Quality had an absolute value of 0.9 on the correlation
+scale. Sleep Duration and Stress were also seen to be strongly
+correlated. Stress was still included in our models due to its
+well-studied impact on sleep; however, we found it important to note the
+strong linear relationship between Stress and Sleep Quality.
+
+A variable importance plot (VIP) was constructed to determine which
+variables had the strongest impact in our Random Forest Model. ![The
+ranking of each variable\'s importance in determining sleep quality
+according to our Random Forest
+model.](images/variableimportance.png)Sleep Duration, Stress Level, and
+Age had the highest levels of importance in our model. Physical Activity
+Level was excluded entirely, which was to be expected given our
+exploratory data analysis and simple linear regression models.
 
@@
 
-*DEFINE DECISION TREES. DEFINE BAGGING AND BOOTSTRAPPING. DEFINE RANDOM FOREST MODELS* 
-*DEFINE CROSS-VALIDATION* 
-*DEFINE IMPORTANCE OF NULL MODEL TO COMPARE RESULTS*
+*DEFINE DECISION TREES. DEFINE BAGGING AND BOOTSTRAPPING. DEFINE RANDOM
+FOREST MODELS* *DEFINE CROSS-VALIDATION* *DEFINE IMPORTANCE OF NULL
+MODEL TO COMPARE RESULTS*
 
-*INCLUDE CITATIONS OF CHAPTER 9,10,11 OF HMLR LINKED IN "MANY TREE MODELS" SECTION OF CONTENT FOLDER 11*
+*INCLUDE CITATIONS OF CHAPTER 9,10,11 OF HMLR LINKED IN "MANY TREE
+MODELS" SECTION OF CONTENT FOLDER 11*
 
@@
 
-Root Mean Squared Error (RMSE) is a common metric used to evaluate regression models.
-The RMSE is formally defined as the square root of the mean square of all error and is defined by the following formula: *Include LaTeX equation of RMSE?* .
-It is important that while RMSE is scale dependent, common practice notes that low RMSE values indicate stronger model performance.
+Root Mean Squared Error (RMSE) is a common metric used to evaluate
+regression models. The RMSE is formally defined as the square root of
+the mean square of all error and is defined by the following formula:
+$$\sum_{i=1}^{D}(x_i-y_i)^2$$
+
+It is important to consider that while RMSE is scale dependent, common
+practice notes that low RMSE values indicate stronger model performance.
+
+R-squared (R²) is a statistical measure used in regression analyses. It
+quantifies the proportion of the variance in a dependent variable that
+can be explained by an independent variable within a regression model
+and is calculated with the following formula:
+
+$$
+[ R^2 = 1 - \frac{{\text{{Unexplained Variation}}}}{{\text{{Total Variation}}}} ]
+$$
+
+We used both RMSE and R² to evaluate the performance of our model
+compared to a "null model". Null models generate predictions for an
+outcome of interest without using any predictors. Instead, they simply
+average the values of the outcome of interest to make "predictions".
+
+![The results generated from our RF model. RMSE increases when more
+predictors are added to the model. The R-squared value also decreases as
+more predictors are added to the model.](images/RFtuningresults.png)
+
+Figure (*name/order figures after we finalize draft*) shows that as we
+include more predictors in our model, RMSE increases slightly while the
+R-squared value decreases. This indicates that some variables in our
+dataset are more important for predicting sleep quality than the others.
+Based off of our variable importance graph, we can assume that stress
+level, sleep duration, and age are the factors that carry the most
+weight when predicting sleep quality.
+
+We went on to evaluate how well our RF model predicted values by
+creating an observed vs. predicted plot.
+
+![A comparison of the actual values for sleep quality in our dataset
+(observed) compared to the values predicted by our fitted RF
+model.](images/observed_vs_predicted.png)
+
+While the predicted values don't align perfectly with the red "ideal
+value" line, they fall very close overall, indicating that our model can
+sufficiently predict sleep quality using unseen data.
 
@@
 
-This analysis was conducted under R version 4.3.1 on a MacOS operating system *cite R*.
-The following R packages were used in the development of these analyses: *R PACKAGES USED: here, skimr, broom, tidyverse, ggplot2, dplyr, corrplot, ranger, vip*
-All processing and analysis code can be found in the Supplementary Material file.
+This analysis was conducted under R version 4.3.1 on a MacOS operating
+system *cite R*. The following R packages were used in the development
+of these analyses: *R PACKAGES USED: here, skimr, broom, tidyverse,
+ggplot2, dplyr, corrplot, ranger, vip* All processing and analysis code
+can be found in the Supplementary Material file.
 
@@
 
-The 'Quality of Sleep' variable has been selected as our outcome of interest.
-It is important to note that this variable is a subjective measure of a participant's self-reported sleep quality.
-The following plot shows that most subjects reported a sleep quality score of 8 (out of 10).
-*REFERENCE SLEEP_DIS PLOT FROM EDA.QMD FILE*
+The 'Quality of Sleep' variable has been selected as our outcome of
+interest. It is important to note that this variable is a subjective
+measure of a participant's self-reported sleep quality. The following
+plot shows that most subjects reported a sleep quality score of 8 (out
+of 10).
+
+![Distribution of Sleep Quality Scores in the
+dataset.](images/unnamed-chunk-17-1.png)
 
@@
 
-A variable importance plot (VIP) was constructed to determine which variables had the strongest impact in our Random Forest Model.
-*INCLUDE PLOT HERE* Sleep Duration, Stress Level, and Age had the highest levels of importance in our model.
-Physical Activity Level was excluded entirely, which was to be expected given our exploratory data analysis and simple linear regression models.
+*To Emma: did you want this text in both places you placed it?* A
+variable importance plot (VIP) was constructed to determine which
+variables had the strongest impact in our Random Forest Model. *INCLUDE
+PLOT HERE* Sleep Duration, Stress Level, and Age had the highest levels
+of importance in our model. Physical Activity Level was excluded
+entirely, which was to be expected given our exploratory data analysis
+and simple linear regression models.
 
@@
 
-# Discussion
+# Discussion (5)
 
@@
 
-*Include that all parameters were self-reported and thus subjected to bias*
+*Include that all parameters were self-reported and thus subjected to
+bias*
 
@@
 
-  https://link.springer.com/article/10.1007/s11325-015-1193-z --> obesity and sleep quality (affects men but not women in China)
+https://link.springer.com/article/10.1007/s11325-015-1193-z --\> obesity
+and sleep quality (affects men but not women in China)
 
-  https://www.scielosp.org/article/gs/2017.v31n1/18-22/en/?uid=680b54f39b# --> effect of age and gender on sleep quality (women worse than men, sleep quality deteriorates with age in women but is a bit all over the place for men)
+https://www.scielosp.org/article/gs/2017.v31n1/18-22/en/?uid=680b54f39b#
+--\> effect of age and gender on sleep quality (women worse than men,
+sleep quality deteriorates with age in women but is a bit all over the
+place for men)
 
-  https://academic.oup.com/sleep/article/33/2/149/2454438#77919376 --> sleep quality by occupation/industry (managerial positions had the worst score)
-
-*ANYTHING BELOW THIS COMMENT IS FROM PRIOR TO THE OVERHAUL AND IS STILL TO BE IGNORED / MOVED TO ANOTHER SECTION OF THE PROJECT*
-
-## Questions/Hypotheses to be addressed
-
-We intend to explore the influence of each variable on sleep quality and identify which ones seem to have the most impact (both positive and negative).
-We also wish to explore if certain variables affect male sleep quality more than females.
-If time allows, we'd also like to see if certain ages seem to be affected by certain variables more often than others.
-We can focus on every predictor in the dataset since there are only 13.
-
-We can likely use GLMs and t-tests/ANOVA tests to see the significance that variables may have in different combinations on an individual's sleep quality.
-Heat maps and various plots will also help with exploratory data analysis.
-We aren't exceptionally experienced with statistical analyses, but these are what we can think of at the moment.
-
-{{< pagebreak >}}
-
-## Data import
-
-We'll load in the cleaned data here. Please reference the eda.qmd and Supplementary-Material.qmd files for full exploratory results and the cleaning process.
-
-```{r Loading the Processed Data}
-#load the cleaned data
-sleepdata<- readRDS(here("data","processed-data", "sleepdataprocessed.rds"))
-
-```
-
-## Statistical analysis
-
-We'll begin to look at some of the relationships between our variables now. Sleep Quality is our outcome of interest for all of these. We'll start with Physical Activity level.
-
-```{r}
-############################
-#### Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor
-
-lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)
-
-# Placing results from lmfit_quality_activty  into a data frame with the tidy function
-lmtable_quality_activity <- broom::tidy(lmfit_quality_activity)
-
-# Viewing the results from the first model fit 
-print(lmtable_quality_activity)
-
-# Saving the lmfit_quality_activity results table  
-lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
-saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)
-```
-
-The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0.
-This has a relatively low standard error and high significance.The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109.
-Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's.This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.
-
-Next, we'll look at Sleep Duration as a predictor.
-
-```{r}
-############################
-#### Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
-
-lmfit_quality_duration <- lm(Quality.of.Sleep ~ Sleep.Duration, sleepdata)
-
-# Placing results from lmfit_quality_duration into a data frame with the tidy function
-lmtable_quality_duration <- broom::tidy(lmfit_quality_duration)
-
-#Viewing the results from the second model fit 
-print(lmtable_quality_duration)
-
-# Saving the lmfit_quality_duration results table  
-lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")
-saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
-
-```
-
-For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep.
-This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high.
-For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1.
-This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.
-
-Next, we'll look at Stress Level as our predictor.
-
-```{r}
-############################
-####Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.
-
-lmfit_quality_stress <- lm(Quality.of.Sleep~ Stress.Level, sleepdata)
-
-#Placing results from lmfit_quality_stress into a data frame with the tidy function
-lmtable_quality_stress <- broom::tidy(lmfit_quality_stress)
-
-#Viewing the results from the sixth model fit 
-print(lmtable_quality_stress)
-
-# Saving the lmfit_quality_stress results table  
-lmtable_quality_stress = here("results", "tables", "lmfit6table.rds")
-saveRDS(lmtable_quality_stress, file = lmtable_quality_stress)
-```
-
-According to our model, Stress has a strongly defined relationship with sleep quality.
-We can see that a value of 0 on the stress score leads to a 10 unit increase in quality of sleep.
-The Stress Level coefficient indicates that as stress increases by one unit, sleep quality decreases by about 0.6 of a unit.
-The t-statistics and p-values for the values in this model are extremely strong, indicating a well-defined relationship between stress and sleep quality.
-
-Finally, we'll look at Age as our predictor.
-
-```{r}
-############################
-####Fitting a model using Sleep Quality as the outcome and Age as a predictor.
-
-lmfit_quality_age <- lm(Quality.of.Sleep~ Age, sleepdata)
-
-#Placing results from lmfit_quality_gender_age into a data frame with the tidy function
-lmtable_quality_age <- broom::tidy(lmfit_quality_age)
-
-#Viewing the results from the seventh model fit 
-print(lmtable_quality_age)
-
-# Saving the lmfit_quality_gender_age results table  
-lmtable_quality_age = here("results", "tables", "lmfit8table.rds")
-saveRDS(lmtable_quality_age, file = lmtable_quality_age)
-
-```
-
-Our model shows a strong relationship between age and quality of sleep.
-As Age increases, sleep quality seems to improve.
-Each year increase in age is predicted to have a 0.065 increase in sleep quality score.
-The intercept here is also interesting as an age of "0" is predicted to have a sleep score of 4.57, which is relatively poor.
-The t-statistics and p-values for these are both strong.
-
-{{< pagebreak >}}
-
-## Exploratory/Descriptive analysis
-
-Here are the most interesting/distinct plots from the EDA:
-
-Physical Activity plot:
-
-We can clearly see that Physical Activity level has a positive correlation with Quality of Sleep.
-```{r Physical Activity Plot, warning=FALSE, message=FALSE}
-#load plot package
-library(ggplot2)
-
-# Box plot for Quality of Sleep by Physical Activity Level Group
-ggplot(sleepdata, aes(x = PhysicalActivityGroup, y = Quality.of.Sleep)) +
-  geom_boxplot() +
-  labs(title = "Quality of Sleep by Physical Activity Level Group", x = "Physical Activity Level Group", y = "Quality of Sleep")
-```
-
-Gender, Age, and Occupation:
-
-We can see some clear differences in self-reported Quality of Sleep between these different groups already.
-Females tend to report an average of 1 higher according to the boxplot.
-The different age groups are similar, but the oldest (50-60) reports the highest quality of sleep by far.
-Our occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
-```{r Gender, Age, and Occupation Plot}
-# Boxplot for Quality of Sleep by Gender
-ggplot(sleepdata, aes(x = Gender, y = Quality.of.Sleep)) +
-  geom_boxplot() +
-  labs(title = "Quality of Sleep by Gender", x = "Gender", y = "Quality of Sleep")
-
-# Create bins for Age in order to create a clean boxplot.
-sleepdata$AgeGroup <- cut(sleepdata$Age, breaks = c(20, 30, 40, 50, 60), labels = c("20-30", "30-40", "40-50", "50-60"), include.lowest = TRUE)
-
-# Calculate mean Quality of Sleep for each Age Group to plot with means for cleaner visuals
-sleepdata_summary <- sleepdata %>%
-  group_by(AgeGroup) %>%
-  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
-
-# Bar plot for Mean Quality of Sleep by Age Group with colorblind-friendly palette
-ggplot(sleepdata_summary, aes(x = AgeGroup, y = MeanQuality, fill = AgeGroup)) +
-  geom_bar(stat = "identity") +
-  scale_fill_brewer(palette = "Set2") +
-  labs(title = "Mean Quality of Sleep by Age Group", x = "Age Group", y = "Mean Quality of Sleep")
-
-# Calculate mean Quality of Sleep for each Occupation in order to create a bar plot.
-sleepdata_summary <- sleepdata %>%
-  group_by(Occupation) %>%
-  summarise(MeanQuality = mean(Quality.of.Sleep, na.rm = TRUE))
-
-# Bar plot for Mean Quality of Sleep by Occupation with manually specified colors
-ggplot(sleepdata_summary, aes(x = Occupation, y = MeanQuality, fill = Occupation)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = rainbow(length(unique(sleepdata_summary$Occupation)))) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Mean Quality of Sleep by Occupation", x = "Occupation", y = "Mean Quality of Sleep")
-```
-
-BMI:
-We can clearly see that Normal weights seem to have better sleep on average than the obese and overweight categories.
-However, there are very few in the obese category, so conclusions with that group may not be as supported as others.
-
-```{r BMI plot}
-# Box plot for Quality of Sleep by BMI Category
-ggplot(sleepdata, aes(x = BMI.Category, y = Quality.of.Sleep)) +
-  geom_boxplot() +
-  labs(title = "Box Plot of Quality of Sleep by BMI Category", x = "BMI Category", y = "Quality of Sleep")
-```
-
-Blood Pressure:
-
-The boxplots look a bit strange for the Elevated and Normal levels, but we can see that it's likely due to a shortage of data for these categories.
-We can see, however, that Quality of Sleep does seem to be negatively correlated with increasing stages of hypertension.
-
-```{r Blood Pressure Plot}
-# Bar plot for cat_bp
-ggplot(sleepdata, aes(x = cat_bp)) +
-  geom_bar() +
-  labs(title = "Bar Plot of cat_bp", x = "cat_bp", y = "Count")
-
-# Box plot for Quality of Sleep by cat_bp
-ggplot(sleepdata, aes(x = cat_bp, y = Quality.of.Sleep)) +
-  geom_boxplot() +
-  labs(title = "Box Plot of Quality of Sleep by cat_bp", x = "cat_bp", y = "Quality of Sleep")
-
-```
-
-And Stress Level:
-Our scatterplot shows a clear negative correlation between Stress Level and Quality of Sleep. However, this is a subjective scale, so data should be taken with a grain of salt for now.
-
-```{r Stress Level Plot}
-# Scatterplot for Quality of Sleep by Stress Level
-ggplot(sleepdata, aes(x = Stress.Level, y = Quality.of.Sleep)) +
-  geom_point() +
-  geom_smooth(method = "loess") +
-  labs(title = "Quality of Sleep by Stress Level", x = "Stress Level (scale: 1-10)", y = "Quality of Sleep")
-```
-
-The other plots were either self explanatory (e.g. Sleep Duration meant higher sleep quality) or had less defined trends (e.g. heart rate and daily steps).
+https://academic.oup.com/sleep/article/33/2/149/2454438#77919376 --\>
+sleep quality by occupation/industry (managerial positions had the worst
+score)

### 2024-04-16 R/analysis-code/statistical-analysis.R

@@ -13,7 +13,6 @@ library(ranger)
 library(vip)
 library(ggplot2)
 
-# Writing new dataframe, sleepdata, into a separate excel file to ensure reproducibility 
 
 ######################################
 # Fitting Models and Statistical Analysis # 
@@ -23,7 +22,7 @@ library(ggplot2)
 #### Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor
 
 #Load the data
-sleepdata<- readRDS(here("data","processed-data","sleepdataprocessed.RDS"))
+sleepdata<- readRDS(here("data","processed-data","sleepdataprocessed.rds"))
 
 lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)  
 
@@ -219,7 +218,11 @@ continuous_vars <- sleepdatafinal[, c("Age", "Sleep.Duration", "Quality.of.Sleep
 correlation_matrix <- cor(continuous_vars)
 
 # Create a pairwise correlation plot (using corrplot)
-corrplot(correlation_matrix, method = "circle")
+correlation <- corrplot(correlation_matrix, method = "circle")
+
+#Save the correlation plot
+correlation = here("results", "figures", "correlation.rds")
+saveRDS(correlation, file = correlation)
 
 #It seems that stress level and quality of sleep are at an absolute value of 0.9 or higher on the correlation scale. 
 #Sleep Duration and stress level also seem closely tied. Stress level and quality of sleep are both subjective scales given by an individual; this is interesting to note.
@@ -267,6 +270,9 @@ tune_results_rf <- tune_grid(
 print(tune_results_rf)
 #Plot them too
 autoplot(tune_results_rf)
+#Save them
+tune_results_rf = here("results", "figures", "tune_results_rf.rds")
+saveRDS(tune_results_rf, file = tune_results_rf)
 
 #We've successfully fit our RF model; now, we can pick the different pieces apart and see what predictors seem the most important and which model is best.
 
@@ -284,6 +290,9 @@ fitted_model <- extract_fit_parsnip(final_fit)
 
 # Create a variable importance plot
 vip(fitted_model$fit)
+#Save it
+vip = here("results", "figures", "vip.rds")
+saveRDS(vip, file = vip)
 
 #We can see that sleep duration, stress level, and Age seem to have the most importance for our model.
[…]

### 2024-04-18 products/manuscript/Manuscript.qmd
Added background section, editing of manuscript
Emma - 4/17

@@
 
-For our project, we obtained a dataset from Kaggle.com containing 373
-observations of 13 variables from a study on different factors that may
+For our project, we obtained a dataset from [Kaggle](https://www.kaggle.com/datasets) containing 373 observations of 13 variables from a study on different factors that may
 affect sleep quality. Variables of interest from the study included
-physical activity level, age, sex, occupation, stress, and several more.
-In RStudio (version 4.3), we used a combination of packages to create
+physical activity level, age, sex, occupation, stress, gender, BMI, blood pressure, and sleep duration.
+In [RStudio](https://posit.co/download/rstudio-desktop/) (version 4.3.1), we used a combination of packages to create
 models suitable for answering our primary question: which variables are
@@
 
-*Feel free to edit this Emma; I tried to include everything needed*
-
 # Introduction (2)
@@
 [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
-It contains data from a study on individuals of different sexes and age
+The data contains observations from a study on individuals of varying sexes and age
 regarding their sleep quality and different variables that may affect
-this. We are unsure of where this data set came from and how it was
+this. We are unsure of where this data set came from and how the data were
 collected; the publisher on Kaggle has yet to respond to a comment with
@@
 showed a correlation between obesity and worsened sleep quality in men
-but not women. We are interested in seeing if this trend holds up for
-our dataset; given that most of the observations are of men, we expect
+but not women. We are interested in seeing if this trend in present in 
+our dataset; however, given that most of the observations are of men, we expect
 to see a correlation between lower sleep scores and obesity in our data.
@@
 
-*NEED TO INSERT*
+Additionally, based on the findings of Sun et.
+al, that obesity can negatively impact sleep quality, we also expect that higher levels of blood pressure will result in poorer sleep quality.
+The American Sleep Foundation states that both short and long sleep duration increase risk of hypertension in adults (Li & Shang, 2021).
+This draws interesting focus to the circular nature of sleep quality and hypertension.
+Those with hypertension also face a higher risk of insomnia, a disorder affecting the ability to fall asleep, maintain sleep, and sleep without continuous disturbance (Liu et.
+al, 2022).
+Short sleep duration is also positively associated with stress level (APA, 2013).
+Survey findings indicate that the average American sleeps 6.7 hours per night, which is well below the minimum recommended duration seven to nine hours (APA, 2013).
+Similar to the relationship between hypertension and sleep quality, as stress level increases, sleep duration decreases thereby causing poorer sleep quality.
+Poorer sleep quality then leads to higher stress, greater risk of hypertension, and shorter sleep duration.
 
-*EMMA'S SOURCES ARE LINKED HERE:*
-
-https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8345503/ --\> blood
-pressure and sleep
-https://www.apa.org/news/press/releases/stress/2013/sleep --\> sleep
-duration & stress on sleep quality
-https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10503965/#:\~:text=Regular%20physical%20activity%20can%20lead,managing%20sleep%20disorders%20like%20insomnia.
---\> physical activity and sleep quality
+Engaging in physical activity has been suggested to improve both sleep quality and sleep duration.
+Regular moderate to intense physical activity stimulates an increase of melatonin production, a hormone that regulates sleep-wake cycles (Alnawwar et.
+al, 2023).
+Additionally, moderate to intense exercise has shown to have a positive effect on those with insomnia (Alnawwar et.
+al, 2023).
+Physical activity also reduces stress and improves both sleep quality and sleep efficiency (Alnawwar et.
+al, 2023).
+However, studies also show that frequent high-intensity workouts could lead to poorer sleep quality among people with and without insomnia (Alnawwar et.
+al, 2023).
+Studies indicate that consistent moderate exercise three to four times a week (approximately 150 minutes total) is the optimal weekly regimen to improve sleep quality and reduce stress.
+We expect to see similar trends with our data, where increased stress levels negatively impact sleep quality and sleep duration, while noticing a positive association between increased physical activity levels and sleep quality.
 
@@
 erroneous variables in the dataset that we needed to remove. The
-"Blood.Pressure" variable was transformed to reflect the systolic and
+"Blood.Pressure" variable was transformed to reflect the individual systolic and
 diastolic blood pressure measurements from each subject. We then chose
 to add an additional variable, labeled 'cat_bp' to reflect the
-categorical blood pressure status of a subject based on the [*American
+categorical blood pressure status of a subject based on the [American
 Heart Association
-guidelines*](https://www.heart.org/en/health-topics/high-blood-pressure).
+guidelines](https://www.heart.org/en/health-topics/high-blood-pressure).
 Additionally we created a variable, 'StepsGroup' to determine the
 categorical activity level of subjects in the dataset based on these
-guidelines provided by 10,000 steps [*10,000
-steps.org*](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)*.*
+guidelines provided by [10,000 steps.org] (https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
 Finally, we created an additional categorical variable,
@@
 This analysis was conducted under R version 4.3.1 on a MacOS operating
-system *cite R*. The following R packages were used in the development
-of these analyses: *R PACKAGES USED: here, skimr, broom, tidyverse,
-ggplot2, dplyr, corrplot, ranger, vip* All processing and analysis code
+system. The following R packages were used in the development
+of these analyses:here, skimr, broom, tidyverse,
+ggplot2, dplyr, corrplot, ranger, and vip. All processing and analysis code
 can be found in the Supplementary Material file.
@@
 
-*To Emma: did you want this text in both places you placed it?* A
+*To Emma: did you want this text in both places you placed it?*
+*To Kevin: I placed in both sections as a placeholder, whichever you think is best!*
+
+A
 variable importance plot (VIP) was constructed to determine which
@@
 score)
+
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8345503/ --\> blood
+pressure and sleep
+
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9815923/#:~:text=Patients%20with%20hypertension%20commonly%20complain,an%20increased%20risk%20of%20insomnia. --\> hypertension and insomnia
+
+https://www.apa.org/news/press/releases/stress/2013/sleep --\> sleep
+duration & stress on sleep quality
+
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10503965/#:\~:text=Regular%20physical%20activity%20can%20lead,managing%20sleep%20disorders%20like%20insomnia.
+--\> physical activity and sleep quality
+
+https://www.heart.org/en/health-topics/high-blood-pressure
+
+https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500

### 2024-04-18 products/manuscript/Manuscript.qmd
Added Discussion section and edited manuscript
Wrote discussion section. Improved results section with more figures, descriptions of model creation, and descriptions of results. Edited writing in other parts

@@
 author: Kevin Kosewick and Emma Hardin-Parker
-date: "`r Sys.Date()`"
+date: "4-18-2024"
 format:
@@
 
-For our project, we obtained a dataset from [Kaggle](https://www.kaggle.com/datasets) containing 373 observations of 13 variables from a study on different factors that may
-affect sleep quality. Variables of interest from the study included
-physical activity level, age, sex, occupation, stress, gender, BMI, blood pressure, and sleep duration.
-In [RStudio](https://posit.co/download/rstudio-desktop/) (version 4.3.1), we used a combination of packages to create
-models suitable for answering our primary question: which variables are
-most important for determining an individual's sleep quality and what
-effect do they have? After conducting an exploratory data analysis (EDA)
-we hypothesized that stress, sleep duration, and occupation would have
-the largest impact on sleep quality. We believed that stress and sleep
+For our project, we obtained a dataset from
+[Kaggle](https://www.kaggle.com/datasets) containing 373 observations of
+13 variables from a study on different factors that may affect sleep
+quality. Variables of interest from the study included physical activity
+level, age, sex, occupation, stress, gender, BMI, blood pressure, and
+sleep duration. In [RStudio](https://posit.co/download/rstudio-desktop/)
+(version 4.3.1), we used a combination of packages to create models
+suitable for answering our primary question: which variables are most
+important for determining an individual's sleep quality and what effect
+do they have? After conducting an exploratory data analysis (EDA) we
+hypothesized that stress, sleep duration, and occupation would have the
+largest impact on sleep quality. We believed that stress and sleep
 quality would be negatively correlated, sleep duration and quality would
@@
 [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
-The data contains observations from a study on individuals of varying sexes and age
-regarding their sleep quality and different variables that may affect
-this. We are unsure of where this data set came from and how the data were
-collected; the publisher on Kaggle has yet to respond to a comment with
-this question posted on March 5th 2024. Regardless, it contains
-interesting variables that have been shown in previous studies to have
-an effect on sleep quality.
+The data contains observations from a study on individuals of varying
+sexes and age regarding their sleep quality and different variables that
+may affect this. We are unsure of where this data set came from and how
+the data were collected; the publisher on Kaggle has yet to respond to a
+comment with this question posted on March 5th 2024. Regardless, it
+contains interesting variables that have been shown in previous studies
+to have an effect on sleep quality.
 
@@
 but not women. We are interested in seeing if this trend in present in
-our dataset; however, given that most of the observations are of men, we expect
-to see a correlation between lower sleep scores and obesity in our data.
-Furthermore, it has been shown that age and gender have a notable effect
-on sleep quality (Madrid-Valero et al. 2016). Women seem to experience a
-deterioration in sleep quality as they age; however, while this trend is
-still present in men, it is notably less consistent and can vary
-dramatically between individuals. We expect to see similar trends in our
-dataset, but as stated previously, our observations contain
-significantly more men than women. Finally, we wish to test whether or
-not certain occupational industries have different effects on sleep
-quality. It has been shown that managerial positions tend to have the
-worst sleep quality among civilian sector workers while 24 jobs that
+our dataset; however, given that most of the observations are of men, we
+expect to see a correlation between lower sleep scores and obesity in
+our data. Furthermore, it has been shown that age and gender have a
+notable effect on sleep quality (Madrid-Valero et al. 2016). Women seem
+to experience a deterioration in sleep quality as they age; however,
+while this trend is still present in men, it is notably less consistent
+and can vary dramatically between individuals. We expect to see similar
+trends in our dataset, but as stated previously, our observations
+contain significantly more men than women. Finally, we wish to test
+whether or not certain occupational industries have different effects on
+sleep quality. It has been shown that managerial positions tend to have
+the worst sleep quality among civilian sector workers while 24 jobs that
 have rotating shifts tend to have the worst sleep quality (Luckhaupt et
@@
 
-Additionally, based on the findings of Sun et.
-al, that obesity can negatively impact sleep quality, we also expect that higher levels of blood pressure will result in poorer sleep quality.
-The American Sleep Foundation states that both short and long sleep duration increase risk of hypertension in adults (Li & Shang, 2021).
-This draws interesting focus to the circular nature of sleep quality and hypertension.
-Those with hypertension also face a higher risk of insomnia, a disorder affecting the ability to fall asleep, maintain sleep, and sleep without continuous disturbance (Liu et.
-al, 2022).
-Short sleep duration is also positively associated with stress level (APA, 2013).
-Survey findings indicate that the average American sleeps 6.7 hours per night, which is well below the minimum recommended duration seven to nine hours (APA, 2013).
-Similar to the relationship between hypertension and sleep quality, as stress level increases, sleep duration decreases thereby causing poorer sleep quality.
-Poorer sleep quality then leads to higher stress, greater risk of hypertension, and shorter sleep duration.
+Additionally, based on the findings of Sun et. al that obesity can
+negatively impact sleep quality, we also expect that higher levels of
+blood pressure will result in poorer sleep quality. The American Sleep
+Foundation states that both short and long sleep duration increase risk
+of hypertension in adults (Li & Shang, 2021). This draws interesting
+focus to the circular nature of sleep quality and hypertension. Those
+with hypertension also face a higher risk of insomnia, a disorder
+affecting the ability to fall asleep, maintain sleep, and sleep without
+continuous disturbance (Liu et. al, 2022). Short sleep duration is also
+positively associated with stress level (APA, 2013). Survey findings
+indicate that the average American sleeps 6.7 hours per night, which is
+well below the minimum recommended duration seven to nine hours (APA,
+2013). Similar to the relationship between hypertension and sleep
+quality, as stress level increases, sleep duration decreases thereby
+causing poorer sleep quality. Poorer sleep quality then leads to higher
+stress, greater risk of hypertension, and shorter sleep duration.
 
-Engaging in physical activity has been suggested to improve both sleep quality and sleep duration.
-Regular moderate to intense physical activity stimulates an increase of melatonin production, a hormone that regulates sleep-wake cycles (Alnawwar et.
-al, 2023).
-Additionally, moderate to intense exercise has shown to have a positive effect on those with insomnia (Alnawwar et.
-al, 2023).
-Physical activity also reduces stress and improves both sleep quality and sleep efficiency (Alnawwar et.
-al, 2023).
-However, studies also show that frequent high-intensity workouts could lead to poorer sleep quality among people with and without insomnia (Alnawwar et.
-al, 2023).
-Studies indicate that consistent moderate exercise three to four times a week (approximately 150 minutes total) is the optimal weekly regimen to improve sleep quality and reduce stress.
-We expect to see similar trends with our data, where increased stress levels negatively impact sleep quality and sleep duration, while noticing a positive association between increased physical activity levels and sleep quality.
+Engaging in physical activity has been suggested to improve both sleep
+quality and sleep duration. Regular moderate to intense physical
+activity stimulates an increase of melatonin production, a hormone that
+regulates sleep-wake cycles (Alnawwar et. al, 2023). Additionally,
+moderate to intense exercise has shown to have a positive effect on
+those with insomnia (Alnawwar et. al, 2023). Physical activity also
+reduces stress and improves both sleep quality and sleep efficiency
+(Alnawwar et. al, 2023). However, studies also show that frequent
+high-intensity workouts could lead to poorer sleep quality among people
+with and without insomnia (Alnawwar et. al, 2023). Studies indicate that
+consistent moderate exercise three to four times a week (approximately
+150 minutes total) is the optimal weekly regimen to improve sleep
+quality and reduce stress. We expect to see similar trends with our
+data, where increased stress levels negatively impact sleep quality and
+sleep duration, while noticing a positive association between increased
+physical activity levels and sleep quality.
 
@@
 The dataset originally contained 373 observations for 13 variables
-covering a broad spectrum health and lifestyle variables associated with
-sleep quality. The curators of the dataset did not provide information
-on how or where this data was collected, so we solely relied on the
-codebook for variable definitions. There were no missing values or
-erroneous variables in the dataset that we needed to remove. The
-"Blood.Pressure" variable was transformed to reflect the individual systolic and
-diastolic blood pressure measurements from each subject. We then chose
-to add an additional variable, labeled 'cat_bp' to reflect the
-categorical blood pressure status of a subject based on the [American
-Heart Association
+covering a broad spectrum of health and lifestyle variables associated
+with sleep quality. The curators of the dataset did not provide
+information on how or where this data was collected, so we solely relied
+on the codebook for variable definitions. There were no missing values
+or erroneous variables in the dataset that we needed to remove. The
+"Blood.Pressure" variable was transformed to reflect the individual
+systolic and diastolic blood pressure measurements from each subject. We
+then chose to add an additional variable, labeled 'cat_bp' to reflect
+the categorical blood pressure status of a subject based on the
+[American Heart Association
 guidelines](https://www.heart.org/en/health-topics/high-blood-pressure).
-Additionally we created a variable, 'StepsGroup' to determine the
+Additionally we created a variable named 'StepsGroup' to determine the
 categorical activity level of subjects in the dataset based on these
-guidelines provided by [10,000 steps.org] (https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
-Finally, we created an additional categorical variable,
-'PhysicalActivityGroup' based on 30 minute increments, to categorically
-represent the level of physical activity reported by a subject. The
-outcome of interest is noted as "Quality.of.Sleep" that we will
-reference as Sleep Quality or Quality of Sleep throughout this report.
+guidelines provided by \[10,000 steps.org\]
+(https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:\~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
+Finally, we created an additional categorical variable called
+'PhysicalActivityGroup' to categorically represent the level of physical
+activity reported by a subject. The levels are differentiated by 30
+minute intervals of weekly physical activity. The outcome of interest is
+noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
+Quality of Sleep throughout this report.
 
@@
 
-Given that there were only 13 original variables, and 17 variables after
-final data transformation; simple linear regression models were
-performed to determine baseline associations between variables and Sleep
-Quality, the outcome of interest. These regression models were created
-to determine which variables had the largest impact on Sleep Quality and
-drove the rest of the analyses. Simple linear regression models were
-fitted for the following variables: BMI, cat_bp, Stress Level, Gender,
-Age, and Occupation.
+Given that there were only 17 variables after final data transformation,
+simple linear regression models were performed to determine baseline
+associations between variables and Sleep Quality (the outcome of
+interest). These regression models were created to determine which
+variables had the largest impact on Sleep Quality and drove the rest of
+the analyses. Simple linear regression models were fitted for the
+following variables: BMI, cat_bp, Stress Level, Gender, Age, and
+Occupation.
 
@@
 are most important for determining an individual's sleep quality and
-what specific effect they have. After conducting an (EDA) we
-hypothesized that stress, sleep duration, and occupation would have the
-largest impact on sleep quality. We believed that stress and sleep
-quality would be negatively correlated, sleep duration and quality would
-be positively correlated, and that jobs with longer or more variable
-hours would negatively affect sleep quality.
+what specific effect they have. After conducting an EDA we hypothesized
+that stress, sleep duration, and occupation would have the largest
+impact on sleep quality. We believed that stress and sleep quality would
+be negatively correlated, sleep duration and quality would be positively
+correlated, and that jobs with longer or more variable hours would
+negatively affect sleep quality.
 
@@
 
-Random Forest Models and Cross-Validation were considered to be the
-"best" fit for our outcome of interest, Quality of Sleep. The data were
-not split into train/test subsets as the data contains less than 400
-observations and many of the values were unique in comparison to the
-rest of the data. Prior to constructing this model, a collinearity plot
-was constructed to determine any presence of collinearity in our data.
-![Collinearity between all variables.](images/collinearity.png)
+Random Forest Models and 5-fold Cross-Validation were considered to be
+the "best" fit for our outcome of interest, Quality of Sleep. The data
+were not split into train/test subsets as the data contains less than
+400 observations and many of the values were unique in comparison to the
+rest of the data. Prior to constructing this model, a colinearity plot
+was constructed to determine any presence of colinearity in our data.
+
+![Figure 1: A colinearity plot demonstrating the colinearity between our
+variables.](images/collinearity.png){width="586"}
 
@@
 
-A variable importance plot (VIP) was constructed to determine which
-variables had the strongest impact in our Random Forest Model. ![The
-ranking of each variable\'s importance in determining sleep quality
-according to our Random Forest
-model.](images/variableimportance.png)Sleep Duration, Stress Level, and
-Age had the highest levels of importance in our model. Physical Activity
-Level was excluded entirely, which was to be expected given our
-exploratory data analysis and simple linear regression models.
-
 ## Defining the Models
 
-*DEFINE DECISION TREES. DEFINE BAGGING AND BOOTSTRAPPING. DEFINE RANDOM
-FOREST MODELS* *DEFINE CROSS-VALIDATION* *DEFINE IMPORTANCE OF NULL
-MODEL TO COMPARE RESULTS*
-
-*INCLUDE CITATIONS OF CHAPTER 9,10,11 OF HMLR LINKED IN "MANY TREE
-MODELS" SECTION OF CONTENT FOLDER 11*
+We fit GLMs for each predictor in order to determine their effect on
+sleep quality. Each model only used a single predictor as the only aim
+was to quantify and understand the impact of each for our outcome of
+interest. For our Random Forest model, we used mtry (with a range of
+1-7) and min_n (with a range of 1-21) as tuning parameters to create 300
+different decision trees. Mtry controls how many predictors a decision
+tree can consider at any given point in time. It adds randomness to the
+decision tree creation process and ensures that all of the trees do not
+look the same. Min_n specifies the minimum number of samples that should
+be present in the leaf node after splitting a node. In our case, this
+means that if any terminal node has more than 21 observations and is not
+a pure node, we can split it further into subnodes. This was all done
+using 5-fold Cross-Validation repeated 5 times in order to train the
+data to predict on an unseen dataset.
 
@@
 
-We used both RMSE and R² to evaluate the performance of our model
-compared to a "null model". Null models generate predictions for an
-outcome of interest without using any predictors. Instead, they simply
-average the values of the outcome of interest to make "predictions".
-
-![The results generated from our RF model. RMSE increases when more
-predictors are added to the model. The R-squared value also decreases as
-more predictors are added to the model.](images/RFtuningresults.png)
-
-Figure (*name/order figures after we finalize draft*) shows that as we
-include more predictors in our model, RMSE increases slightly while the
-R-squared value decreases. This indicates that some variables in our
-dataset are more important for predicting sleep quality than the others.
-Based off of our variable importance graph, we can assume that stress
-level, sleep duration, and age are the factors that carry the most
-weight when predicting sleep quality.
-
-We went on to evaluate how well our RF model predicted values by
-creating an observed vs. predicted plot.
-
-![A comparison of the actual values for sleep quality in our dataset
-(observed) compared to the values predicted by our fitted RF
-model.](images/observed_vs_predicted.png)
-
-While the predicted values don't align perfectly with the red "ideal
-value" line, they fall very close overall, indicating that our model can
-sufficiently predict sleep quality using unseen data.
+Null models provide a baseline for the evaluation of a model's
+performance. They generate predictions for an outcome of interest
+without using any predictor variables. Instead, they simply average the
+values of the outcome of interest to make these "predictions". We used
+both RMSE and R² values to evaluate the performance of our model
+compared to a "null model".
 
@@
 This analysis was conducted under R version 4.3.1 on a MacOS operating
-system. The following R packages were used in the development
-of these analyses:here, skimr, broom, tidyverse,
-ggplot2, dplyr, corrplot, ranger, and vip. All processing and analysis code
-can be found in the Supplementary Material file.
+system. The following R packages were used in the development of these
+analyses: here, skimr, broom, tidyverse, ggplot2, dplyr, corrplot,
+ranger, and vip. All processing and analysis code can be found in the
+Supplementary Material file.
 
@@
 interest. It is important to note that this variable is a subjective
-measure of a participant's self-reported sleep quality. The following
-plot shows that most subjects reported a sleep quality score of 8 (out
-of 10).
+measure of a participant's self-reported sleep quality. Figure 2 shows
+that most subjects reported a sleep quality score of 8 (out of 10).
 
-![Distribution of Sleep Quality Scores in the
+![Figure 2: Distribution of Sleep Quality Scores in the
 dataset.](images/unnamed-chunk-17-1.png)
 
-## Exploratory Findings
+Our Random Forest Model built with 5-fold Cross-Validation explored
+different combinations of predictors to determine which parameters gave
+the most accurate predictions of sleep quality.
+
+![Figure 3: The results generated from our RF model. RMSE increases when
+more predictors are added to the model. The R-squared value also
+decreases as more predictors are added to the
+model.](images/RFtuningresults.png)
+
+Figure 3 shows that as we include more predictors in our model, RMSE
+increases slightly while the R-squared value decreases. This indicates
+that some variables in our dataset are more important for predicting
+sleep quality than the others. Based off of our variable importance
+graph, we can assume that stress level, sleep duration, and age are the
+factors that carry the most weight when predicting sleep quality.
+
+We went on to evaluate how well our RF model predicted values by
+creating an observed vs. predicted plot as seen in figure 4.
+
+![Figure 4: A comparison of the actual values for sleep quality in our
+dataset (observed) compared to the values predicted by our fitted RF
+model.](images/observed_vs_predicted.png)
+
+While the predicted values don't align perfectly with the red "ideal
+value" line, they fall very close overall, indicating that our model can
+sufficiently predict sleep quality using unseen data.
+
+![Figure 5: The observed vs. predicted values for our "Null
+model".](images/null_observedvspredicted.png)
+
+We can see in figure 5 that the null model only predicts values that are
+the average of our observed values. As expected, these predictions are
+far from our actual observed values.
+
+We believed that stress and sleep quality would be negatively
+correlated, sleep duration and quality would be positively correlated,
+and that jobs with longer or more variable hours would negatively affect
+sleep quality. We found that all of our assumptions were supported; the
+results of our generalized linear models (GLMs) are included in the
+supplementary material folder as there are too many of them to include
+in the main manuscript.
 
@@
 
-*To Emma: did you want this text in both places you placed it?*
-*To Kevin: I placed in both sections as a placeholder, whichever you think is best!*
+A variable importance plot (VIP) (figure 6) was constructed using the
+"best" set of parameters (as guided by our RF model) to determine which
+variables had the strongest impact on sleep quality.
 
-A
-variable importance plot (VIP) was constructed to determine which
-variables had the strongest impact in our Random Forest Model. *INCLUDE
-PLOT HERE* Sleep Duration, Stress Level, and Age had the highest levels
-of importance in our model. Physical Activity Level was excluded
-entirely, which was to be expected given our exploratory data analysis
-and simple linear regression models.
+![The ranking of each variable's importance in determining sleep quality
+according to our Random Forest model.](images/variableimportance.png)
 
-*Include comparison of RF Model and null model*
+As seen in figure 6, Sleep Duration, Stress Level, and Age had the
+highest levels of importance in our model. Physical Activity Level was
+excluded entirely, which was to be expected given our exploratory data
+analysis and simple linear regression models.
+
+![Table 1: A comparison of the RMSE and R-squared values generated by
+the null and RF models. The null model has no R-squared value due to its
+method of generating predictions.](images/metric_table.png){width="510"}
+
+As seen in table 1, our null model was found to have an RMSE of 1.193;
+compared to our best parameter RF model's RMSE of 0.059, this is
+relatively high. The R-squared value doesn't exist for the null model
+since there is technically no variation between the observed and
+predicted values.
 
@@
 
-*PART 5*
-
 ## Summary
 
-*PART 5*
+The differences between the null model and RF model's metrics show that
+our RF model is better suited for predicting sleep quality on unseen
+data than a model using no predictors. The R-squared value of 0.997 is
+extremely high; this shows that a large amount of variation is explained
+using our RF model. The variable importance plot we created answers our
+questions about which variables are most important for predicting sleep
+quality. As we can see, sleep duration, stress level, and age were the
+top three predictors for sleep quality in our data. This reflects the
+relationships we saw in our colinearity plot as well as the clear
+correlations we found between the variables in our EDA. We did not
+expect to see age so high on the list; this could be due to us grouping
+ages together in our EDA and therefore misleading us in our preliminary
+data exploration. We did not use the groups when creating our models as
+we learned that grouping continuous variables is a poor practice as we
+progressed in our MADA course.
+
+Through our linear models (included in the supplementary materials
+folder), we determined that our hypotheses regarding each variable's
+effect on quality of sleep were supported. The variables determined to
+be the most important by our RF model had the following effects on sleep
+quality: sleep duration was positively correlated, age was positively
+correlated, and stress level was negatively correlated.
 
@@
 
-*PART 5*
+The main limitations of our models stemmed from the data itself. As
+stated in our introduction section, we have been unable to identify
+exactly where the data were collected and how it was gathered. The
+values for stress level and sleep duration were also based on a
+subjective rating given by each participant in the study; this could
+introduce individual bias to the data and prevent an objective
+evaluation of the effect of these variables. An inherent limitation in
+RF models is difficulty of interpretation; we attempted to minimize this
+by creating easy to read tables and graphs for the results. Future
+analyses would benefit from more detailed meta data and a better
+understanding of the experimental design that generated our
+observations.
 
-*Include that all parameters were self-reported and thus subjected to
-bias*
+The strengths of our analysis lie in the easily interpretation of our
+GLMs and the predictive power of an RF model. We were able to clearly
+display and quantify the relationships between our predictors and sleep
+quality, answering our question about each variable's effect on our
+outcome of interest. We were also able to create a powerful RF model
+that explained nearly all of the variation in our data as well as a
+ranking of each variable's impact on sleep quality. Fitting our RF model
+with CV allowed us to train the model to create predictions for "unseen"
+data as well.
 
@@
 
-*PART 5*
+Overall, while our study limitations prevent us from generalizing the
+findings of this analysis broadly, we were able to clearly answer our
+questions about the data. Our findings supported our hypotheses about
+the effects of each predictor on sleep quality. However, our hypothesis
+about which variables would be the most important was not supported by
+the analysis. Instead, we determined that stress level, sleep duration,
+and age were the most important predictors for our outcome of interest.
+Ultimately, the results indicate that the sleep quality of this study's
+participants is impacted disproportionately by an assortment of
+variables dealing with physical, mental, and environmental factors.
 
@@
 
-https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9815923/#:~:text=Patients%20with%20hypertension%20commonly%20complain,an%20increased%20risk%20of%20insomnia. --\> hypertension and insomnia
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9815923/#:\~:text=Patients%20with%20hypertension%20commonly%20complain,an%20increased%20risk%20of%20insomnia.
+--\> hypertension and insomnia
 
@@
 
-https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500
+https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:\~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500

### 2024-04-18 R/analysis-code/statistical-analysis.R

@@ -270,9 +270,6 @@ tune_results_rf <- tune_grid(
 print(tune_results_rf)
 #Plot them too
 autoplot(tune_results_rf)
-#Save them
-tune_results_rf = here("results", "figures", "tune_results_rf.rds")
-saveRDS(tune_results_rf, file = tune_results_rf)
 
 #We've successfully fit our RF model; now, we can pick the different pieces apart and see what predictors seem the most important and which model is best.
 
@@ -290,9 +287,6 @@ fitted_model <- extract_fit_parsnip(final_fit)
 
 # Create a variable importance plot
 vip(fitted_model$fit)
-#Save it
-vip = here("results", "figures", "vip.rds")
-saveRDS(vip, file = vip)
 
 #We can see that sleep duration, stress level, and Age seem to have the most importance for our model.
 #PhysicalActivityLevel doesn't seem to have much of an impact on Quality of Sleep, as it was excluded from the variable importance plot altogether. These results are pretty in line with what we discovered
@@ -306,16 +300,30 @@ predictions <- predict(final_fit, new_data = sleepdatafinal)
 sleepdatafinal$Predicted <- predictions$.pred
 
 # Create an observed vs. predicted plot
-observed_vs_predicted <-ggplot(sleepdatafinal, aes(x = Quality.of.Sleep, y = Predicted)) +
+ggplot(sleepdatafinal, aes(x = Quality.of.Sleep, y = Predicted)) +
   geom_point() +
   geom_abline(color = "red") +
   labs(x = "Observed", y = "Predicted", title = "Observed vs. Predicted Plot") +
   theme_minimal()
 
-#Save it
-observed_vs_predicted = here("results", "figures", "observed_vs_predicted.rds")
-saveRDS(observed_vs_predicted, file = observed_vs_predicted)
+#Now we'll calculate the RMSE and R squared values of our "best" set of parameters
+# Calculate residuals
+residuals <- sleepdatafinal$Quality.of.Sleep - predictions$.pred
+
+ # Calculate RMSE
+ rf_rmse <- sqrt(mean(residuals^2))
+
+ # Calculate R squared
+ sst <- sum((sleepdatafinal$Quality.of.Sleep - mean(sleepdatafinal$Quality.of.Sleep))^2)
+ssr <- sum(residuals^2)
+rf_r_squared <- 1 - (ssr / sst)
+
+#print results
+print(rf_rmse)
+print(rf_r_squared)
+
[…]

### 2024-04-19 products/manuscript/Manuscript.qmd
Linear Models Table
Added a table and description for the linear model results to the manuscript

@@
 the analyses. Simple linear regression models were fitted for the
-following variables: BMI, cat_bp, Stress Level, Gender, Age, and
-Occupation.
+following variables: BMI, cat_bp, Stress Level, Physical Activity Level,
+Sleep Duration, Gender, Age, and Occupation. The other variables were
+determined to have too weak of a relationship after our EDA and were
+therefore left out of the analyses.
 
@@
 
+![Table 1: The results of our GLMs. The "term" column identifies the
+category/variable being analyzed. The "model" column identifies the
+predictors used to fit each model. The "statistic" column is a relative
+measure of the strength of the relationship between each predictor and
+sleep quality. The "estimate" column shows the predicted change in sleep
+quality score as a result of each associated
+predictor.](images/table.png)
+
 We believed that stress and sleep quality would be negatively
@@
 and that jobs with longer or more variable hours would negatively affect
-sleep quality. We found that all of our assumptions were supported; the
-results of our generalized linear models (GLMs) are included in the
-supplementary material folder as there are too many of them to include
-in the main manuscript.
+sleep quality. As seen in table 1, we found that all of our assumptions
+were supported. The "intercept" values under the term column represents
+what the expected sleep quality score would be if a continuous predictor
+was at a value of 0. For categorical variables, it represents the
+reference category (e.g. the intercept for the Gender model represents
+the expected sleep quality score for females).
 
@@
 
-![The ranking of each variable's importance in determining sleep quality
-according to our Random Forest model.](images/variableimportance.png)
+![Figure 6: The ranking of each variable's importance in determining
+sleep quality according to our Random Forest
+model.](images/variableimportance.png)
 
@@
 
-![Table 1: A comparison of the RMSE and R-squared values generated by
+![Table 2: A comparison of the RMSE and R-squared values generated by
 the null and RF models. The null model has no R-squared value due to its
@@
 
-As seen in table 1, our null model was found to have an RMSE of 1.193;
+As seen in table 2, our null model was found to have an RMSE of 1.193;
 compared to our best parameter RF model's RMSE of 0.059, this is
@@
 
-Through our linear models (included in the supplementary materials
-folder), we determined that our hypotheses regarding each variable's
-effect on quality of sleep were supported. The variables determined to
-be the most important by our RF model had the following effects on sleep
-quality: sleep duration was positively correlated, age was positively
-correlated, and stress level was negatively correlated.
+Through our linear models, we determined that our hypotheses regarding
+each variable's effect on quality of sleep were supported. The variables
+determined to be the most important by our RF model had the following
+effects on sleep quality: sleep duration was positively correlated, age
+was positively correlated, and stress level was negatively correlated.
 

### 2024-04-19 R/analysis-code/statistical-analysis.R

@@ -201,6 +201,52 @@ saveRDS(lmtable_quality_occupation, file = lmtable_quality_occupation)
 # Engineers seem to have the best sleep quality score; they're predicted to have a score of about 8.3 (with strong p-values and t-statistics). Scientists seem to have the worst score; they're predicted to have 2 points lower sleep quality than our intercept. The t-statistic and p-value are both strong for this group.
 # Lawyers fall in the middle, having a decent increase in sleep quality compared to the intercept and solid p-values and t-statistics to support this.
 
+#Now we'll create a table that combines all of these results to put in the manuscript.
+# List of all models
+models <- list(
+  PhysicalActivity = lmfit_quality_activity,
+  Occupation = lmfit_quality_occupation,
+  Age = lmfit_quality_age,
+  StressLevel = lmfit_quality_stress,
+  Hypertension = lmfit_quality_BP,
+  SleepDuration = lmfit_quality_duration,
+  BMI = lmfit_quality_BMI,
+  Gender_and_Age = lmfit_quality_gender_age
+)
+
+# Empty data frame to store results
+all_results <- data.frame()
+
+# Loop through each model, tidy the results, and bind to all_results
+for(i in 1:length(models)) {
+  model_name <- names(models)[i]
+  model_results <- broom::tidy(models[[i]])
+  model_results$model <- model_name  # add a column to identify the model
+  all_results <- rbind(all_results, model_results)
+}
+
+# Now all_results contains the combined results of all models
+print(all_results)
+
+# Save the combined results table
+all_results_path = here("results", "tables", "all_results.rds")
+saveRDS(all_results, file = all_results_path)
+
+# Load the package
+library(flextable)
+
+# Transform p-values to scientific notation
+all_results$p.value <- format(all_results$p.value, scientific = TRUE, digits = 2)
+
+# Create a flextable
+ft <- flextable::flextable(all_results)
+
+# Print the table
+print(ft)
+
+# Save the flextable as an image
+flextable::save_as_image(ft, path = "table.png")
+
 ##############################
 #### Now we'll remove unnecessary variables (e.g. ones we've feature engineered into more useful ones) to make our analyses easier.
 
@@ -379,8 +425,8 @@ library(gt)
 # Create a data frame
 results <- data.frame(
   Model = c("Random Forest", "Null Model"),
[…]

### 2024-04-21 products/manuscript/Manuscript.qmd
4/21- Finalized references

@@
 
-*INCLUDE INCREMENTALLY -- FINALIZE*
-
-https://link.springer.com/article/10.1007/s11325-015-1193-z --\> obesity
-and sleep quality (affects men but not women in China)
-
-https://www.scielosp.org/article/gs/2017.v31n1/18-22/en/?uid=680b54f39b#
---\> effect of age and gender on sleep quality (women worse than men,
-sleep quality deteriorates with age in women but is a bit all over the
-place for men)
-
-https://academic.oup.com/sleep/article/33/2/149/2454438#77919376 --\>
-sleep quality by occupation/industry (managerial positions had the worst
-score)
-
-https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8345503/ --\> blood
-pressure and sleep
-
-https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9815923/#:\~:text=Patients%20with%20hypertension%20commonly%20complain,an%20increased%20risk%20of%20insomnia.
---\> hypertension and insomnia
-
-https://www.apa.org/news/press/releases/stress/2013/sleep --\> sleep
-duration & stress on sleep quality
-
-https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10503965/#:\~:text=Regular%20physical%20activity%20can%20lead,managing%20sleep%20disorders%20like%20insomnia.
---\> physical activity and sleep quality
-
-https://www.heart.org/en/health-topics/high-blood-pressure
-
-https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:\~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500
+*INCLUDED IN .BIB FILE* 
+*IF ANYONE HAS TIPS ON HOW TO GET THE REFERENCES IN THE MANUSCRIPT ITSELF PLEASE HELP!*

### 2024-04-21 products/manuscript/Manuscript.qmd
I figured out the references!

@@
     highlight-style: github
-bibliography: ../manuscript/dataanalysis-references.bib
+bibliography: ../manuscript/references_sleepdata.bib
 editor: 
@@
 # References
-
-*INCLUDED IN .BIB FILE* 
-*IF ANYONE HAS TIPS ON HOW TO GET THE REFERENCES IN THE MANUSCRIPT ITSELF PLEASE HELP!*

### 2024-04-26 Review_by Malika/project-review-template.qmd
Project Review

---
title: Project Review Template 
author: MALIKA DHAKHWA
date: 4/25/2024
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: An Analysis of Health & Lifestyle Factors Associated with Sleep Quality

Name of project author(s): Kevin Kosewick and Emma Hardin-Parker

Name of project reviewer: Malika Dhakhwa

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments
You have effectively set the objective and contextualized the project within existing literature, showing how your work aligns with prior studies. I recommend adding a brief overview on the broader importance of sleep quality, such as its impact on health and daily performance, to explain the relevance of your study.

The end of section 2.1 (Background Information) mentions that your study expect to see trends where increased stress levels negatively impact sleep quality and sleep duration. It creates some confusion whether sleep duration is also an outcome variable of the study. I would recommend making necessary adjustments for the clarity of your research focus.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments
Your hypothesis that stress, sleep duration, and occupation would have the largest impact on sleep quality is clearly stated. It is well formulated using the preliminary data analysis.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
You have outlined the data and its variables in the Abstract and Data Processing sections. The data and units of measurement of the variables are clearly explained in the exploratory data analysis (EDA) file. For greater clarity, it would be beneficial to mention the units of measurement in the manuscript as well.

I was unable to find a codebook. The provided data link leads to the general Kaggle datasets page rather than directly linking to your specific dataset.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments
You have done a great job in performing and explaining the data cleaning, processing and exploring.

While you have described the methods used for feature engineering, detailing the levels of the new factor variables 'cat_bp', 'StepsGroup', and 'PhysicalActivityGroup' would further clarify your methods.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments
The use of Random Forest (RF) analysis appears well-suited to the dataset. You have also employed the appropriate performance measures, variable selection techniques, data pre-processing steps, and model evaluation strategies.

However, there is limited explanation regarding the decision to apply GLM to each predictor separately and not to use it in a model including all predictors. Additionally, I would recommend detailing the rationale for comparing only the null model with the RF model.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
You have made a good use of the figures and tables in the manuscript to help visualize the data and understand the results. Overall, the presentation of results is commendable, with clear and accessible tables and figures.

The readability of tables could be further improved by limiting the number of decimal places displayed. This would make the tables less cluttered.

For categorical variables, I would recommend to include the reference categories in parentheses (e.g., BMI.CategoryObese(Ref.: Normal)) directly in the results tables. This addition would clarify the comparisons being made, creating an ease in interpreting the results.

I noticed a discrepancy in the explanation of Figure 3, where the RMSE is described as increasing and the R-squared as decreasing with the addition of more predictors.

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
You have thoroughly explained the findings of your study in the discussion section, clearly acknowledging both its strengths and limitations. To enhance the robustness of your discussion, I would recommend addressing the fact that the data was not split into training and testing subsets, which is a common practice for assessing model generalizability.

### Summary assessment
* strong, complete and clear discussion

## Further comments
In your statistical-analysis.R file, at line 199, I believe there is a misunderstanding in the interpretation of the results. It is mentioned that the intercept indicates that being unemployed results in a sleep quality score of about 7. However, it seems that this interpretation actually pertains to the reference category of Business/Finance instead of the unemployed category.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments
You have done a great job in organizing your project, with files generally placed in well-labeled folders with the exception for figures. They are located in the images folder within the product folder, which differs from what is described in the main ReadMe file.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments
You have done a great job with documentation of the project. Especially in the eda and analysis code files, each phase and the decisions behind each step are well commented. Anyone reviewing your project can easily follow and comprehend the entire analysis process.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments
Your project stands out in reproducibility. The documentation clearly explains how to replicate the analysis without the need of any manual intervention. I was able to re-run the entire analysis without any issues.

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments
Your study addresses its hypotheses using the Random Forest model. However, incorporating comparisons with other advanced machine learning tools, such as LASSO, could provide a more comprehensive analysis. Additionally, enhancing the validation process by splitting the data into training and testing subsets would strengthen the robustness and generalizability of your findings.

### Summary assessment
* decent level of thoroughness

## Further comments
- Consider organizing all library import statements at the top of your statistical-analysis.R script. 
- Refine your references and in-text citation of the link 10,000 steps.org

### 2024-04-26 Review_by_Kelly_Cao/Project_Review.qmd
Project Review-KellyCao

--- 2024-04-26 Review_by Malika/project-review-template.qmd
@@
 ---
-title: Project Review Template 
-author: MALIKA DHAKHWA
-date: 4/25/2024
+title: Project Review 
+author: Kelly Cao
+date: 04/26/2024
 format: 
@@
 
-Name of project reviewer: Malika Dhakhwa
-
-# Specific project content evaluation
-Evaluate the different parts of the project by filling in the sections below.
+Name of project reviewer: Kelly Cao
 
@@
 ### Feedback and Comments
-You have effectively set the objective and contextualized the project within existing literature, showing how your work aligns with prior studies.
-I recommend adding a brief overview on the broader importance of sleep quality, such as its impact on health and daily performance, to explain the relevance of your study.
 
-The end of section 2.1 (Background Information) mentions that your study expect to see trends where increased stress levels negatively impact sleep quality and sleep duration.
-It creates some confusion whether sleep duration is also an outcome variable of the study.
-I would recommend making necessary adjustments for the clarity of your research focus.
+The background was well described with plenty of notable references.
+It's comprehensive and pairs well with the study you guys implemented.
+Your background clearly note the important predictors for sleep quality, while also pairing it with many references to different studies.
+
+Although, I did have some confusion with how you mentioned conflicting references at the end.
+Perhaps, you can bring some more clarity for the direction you wish to support or if that's another question you wish to answer.
 
@@
 ### Feedback and Comments
-Your hypothesis that stress, sleep duration, and occupation would have the largest impact on sleep quality is clearly stated. It is well formulated using the preliminary data analysis.
+
+While reading you background, I understand you guys were hoping to study the relationship between the many predictors listed and quality of sleep. This was easily determined.
 
@@
 ### Feedback and Comments
-You have outlined the data and its variables in the Abstract and Data Processing sections.
-The data and units of measurement of the variables are clearly explained in the exploratory data analysis (EDA) file.
-For greater clarity, it would be beneficial to mention the units of measurement in the manuscript as well.
 
-I was unable to find a codebook. The provided data link leads to the general Kaggle datasets page rather than directly linking to your specific dataset.
+Although it is unfortunate you guys were unable to determine the source of the data set, you guys clearly noted where in Kaggle you discovered it and why you still proceeded to use this set.
+I note that you guys mentioned a codebook in your manuscript, but I was unable to find it in your repository or in the Kaggle link provided.
 
@@
 ### Feedback and Comments
-You have done a great job in performing and explaining the data cleaning, processing and exploring.
 
-While you have described the methods used for feature engineering, detailing the levels of the new factor variables 'cat_bp', 'StepsGroup', and 'PhysicalActivityGroup' would further clarify your methods.
+It appears that the raw data set was fairly clean from the get-go.
+You guys clearly documented any changes you made to it and the new variables added.
+I would suggest cleaning up some of the formatting issue in your data section, as it made it somewhat difficult to read.
 
@@
 ### Feedback and Comments
-The use of Random Forest (RF) analysis appears well-suited to the dataset.
-You have also employed the appropriate performance measures, variable selection techniques, data pre-processing steps, and model evaluation strategies.
 
-However, there is limited explanation regarding the decision to apply GLM to each predictor separately and not to use it in a model including all predictors.
-Additionally, I would recommend detailing the rationale for comparing only the null model with the RF model.
+You guys used the appropriate models for the study.
+The dataset is well-suited for Random Forest (RF) analysis, especially considering its classification nature, which aligns perfectly with the strengths of RF in handling such problems.
+Additionally, the use of the simple linear regression model was a great step in determining major predictors to focus on.
 
@@
 ### Feedback and Comments
-You have made a good use of the figures and tables in the manuscript to help visualize the data and understand the results.
-Overall, the presentation of results is commendable, with clear and accessible tables and figures.
 
-The readability of tables could be further improved by limiting the number of decimal places displayed. This would make the tables less cluttered.
+The figures and tables are super clear and easy to read. I felt that you chose the best images to represent the data and the analysis you provided.
 
-For categorical variables, I would recommend to include the reference categories in parentheses (e.g., BMI.CategoryObese(Ref.: Normal)) directly in the results tables.
-This addition would clarify the comparisons being made, creating an ease in interpreting the results.
-
-I noticed a discrepancy in the explanation of Figure 3, where the RMSE is described as increasing and the R-squared as decreasing with the addition of more predictors.
+I particularly enjoyed the colinearity plot and the observed vs prediction plots, as I felt they provided the most information.
 
@@
 ### Feedback and Comments
-You have thoroughly explained the findings of your study in the discussion section, clearly acknowledging both its strengths and limitations.
-To enhance the robustness of your discussion, I would recommend addressing the fact that the data was not split into training and testing subsets, which is a common practice for assessing model generalizability.
+
+The strengths and limitations were clearly mentioned. The limitation being the dataset, and the strength being the easy and clear interpretation of the data set.
 
@@
 ## Further comments
-In your statistical-analysis.R file, at line 199, I believe there is a misunderstanding in the interpretation of the results.
-It is mentioned that the intercept indicates that being unemployed results in a sleep quality score of about 7.
-However, it seems that this interpretation actually pertains to the reference category of Business/Finance instead of the unemployed category.
-
-# Overall project content evaluation
-Evaluate overall features of the project  by filling in the sections below.
+It appears that the project is practically finished. You guys just need to clean up your manuscript and include the references.
 
@@
 ### Feedback and Comments
-You have done a great job in organizing your project, with files generally placed in well-labeled folders with the exception for figures.
-They are located in the images folder within the product folder, which differs from what is described in the main ReadMe file.
+
+The repository was clean and neat. The README files were clear in the description. The files were reasonably named. It seems all of the junk files were properly removed.
 
@@
 ### Feedback and Comments
-You have done a great job with documentation of the project.
-Especially in the eda and analysis code files, each phase and the decisions behind each step are well commented.
-Anyone reviewing your project can easily follow and comprehend the entire analysis process.
+
+It appears everything is properly documented.
+Your code was easy to follow.
+As a personal preference, I would enjoy having the statistical analysis be performed in a quarto document, as it would allow me to render and view it in a more digestable format.
 
@@
 ### Feedback and Comments
-Your project stands out in reproducibility.
-The documentation clearly explains how to replicate the analysis without the need of any manual intervention.
-I was able to re-run the entire analysis without any issues.
+
+All of the major components of your project was easily reproducible.
+I started off the review by quickly running through each of your codes and major documents to see if it would run and render.
+However, since you mentioned your Supplementary-Material.qmd in a README.qmd, I went to see it and ran into a problem of rendering it and got this error:
+
+```
+processing file: Supplementary-Material.qmd
+  |....                                               |   8% [unnamed-chunk-2] 
+Quitting from lines 73-74 [unnamed-chunk-2] (Supplementary-Material.qmd)
+Error:
+! object 'sleepdata' not found
+Backtrace:
+ 1. base::print(sum(is.na(sleepdata)))
+
+Execution halted
+```
+
+If you are showing supplementary exploratory analysis, I would recommend using the `read_chunk()` function, as it will allow you to pull specific lines from a .R script with some specific formatting.
+If that wasn't the intended goal, I would also suggest you saving the data-frame/data-set you intend to work across different files, as it would make things easier for other people to access.
+The error seems to indicate that the object you are referencing is not found in the environment we are working with.
+
+I'm not sure if this was an issue on my end, but when attempting to source your analysis .R script, this error would pop up:
+
+```
++ library(broo .... [TRUNCATED] 
+Error in -library(broom) : invalid argument to unary operator
+```
+The other lines after that ran perfectly fine.
+
+Other than that, all of the remaining codes and documents ran easily and matched the .html output file.
 
@@
 ### Feedback and Comments
-Your study addresses its hypotheses using the Random Forest model.
-However, incorporating comparisons with other advanced machine learning tools, such as LASSO, could provide a more comprehensive analysis.
-Additionally, enhancing the validation process by splitting the data into training and testing subsets would strengthen the robustness and generalizability of your findings.
+
+The study was performed very well, but I imagine it can be strengthened further with additional analysis, using different models for comparison.
+I would also like to see further assessment on your model with cross validation studies to further prove the robustness of your model.
+
+Despite that, I felt the hypotheses were addressed properly.
 
@@
 ## Further comments
-- Consider organizing all library import statements at the top of your statistical-analysis.R script. 
-- Refine your references and in-text citation of the link 10,000 steps.org
+Minor issues noted above:
+
+ - Formatting issue in data section of manuscript
+ - References in manuscript
+ - Supplementary markdown file 
+ - Missing codebook
+
+Overall, I felt the study was thorough and nicely documented. Everything was easy to follow, and there were little to no issue with going through your repository.

### 2024-04-29 R/analysis-code/statistical-analysis.qmd
Final Part Revisions
Made statistical analysis file a .qmd and have cleaned some issues pointed out in the project reviews.

@@
 ---
-title: "statistical-analysis.qmd"
+title: "statistical-analysis"
 format: html
@@
 
-## Quarto
+#Exploratory Data Analysis: Project Part 2
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+Kevin Kosewick & Emma Hardin-Parker
 
-## Running Code
+```{r, message=FALSE, warning=FALSE}
+# Loading Necessary Packages
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+library(broom) 
+library(here) 
+library(tidymodels)
+library(tidyverse)
+library(corrplot)
+library(ranger)
+library(vip)
+library(ggplot2)
+library(flextable)
+library(gt)
+```
+
+# Fitting Models and Statistical Analysis
+
+#### Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor
 
 ```{r}
-1 + 1
+#Load the data
+
+sleepdata<- readRDS(here("data","processed-data","sleepdataprocessed.rds"))
+
+lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)
+
+# Placing results from lmfit_quality_activty  into a data frame with the tidy function
+
+lmtable_quality_activity <- broom::tidy(lmfit_quality_activity)
+
+# Viewing the results from the first model fit
+
+print(lmtable_quality_activity)
+
+# Saving the lmfit_quality_activity results table
+
+lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")
+
+saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)
+
 ```
 
-You can add options to executable code like this
+The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0.
+This has a relatively low standard error and high significance.
+The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109.
+Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's.
+This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.
+
+####Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
+```{r}
+lmfit_quality_duration <- lm(Quality.of.Sleep ~ Sleep.Duration, sleepdata)
+
+# Placing results from lmfit_quality_duration into a data frame with the tidy function
+
+lmtable_quality_duration <- broom::tidy(lmfit_quality_duration)
+
+#Viewing the results from the second model fit
+
+print(lmtable_quality_duration)
+
+# Saving the lmfit_quality_duration results table
+
+lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")
+
+saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
+```
+
+For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep.
+This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high.
+
+For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1.
+This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.
+
+#### Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors
+```{r}
+lmfit_quality_duration_occupation <- lm(Quality.of.Sleep ~ Sleep.Duration + Occupation, sleepdata)
+
+# Placing results from lmfit_quality_duration_occupation into a data frame with the tidy function
+
+lmtable_quality_duration_occupation <- broom::tidy(lmfit_quality_duration_occupation)
+
+#Viewing the results from the third model fit
+
+print(lmtable_quality_duration_occupation)
+
+# Saving the lmfit_quality_duration_occupation results table
+
+lmtable_quality_duration_occupation = here("results", "tables", "lmfit3table.rds")
+
+saveRDS(lmtable_quality_duration_occupation, file = lmtable_quality_duration_occupation)
+```
+
+Since Occupation is a categorical variable, interpreting our intercept value is tricky.
+It's hard to get good information from it since we don't have any data from unemployed people.
+However, the coefficients for our predictors show a variety of interesting relationships.
+
+Sleep Duration shows a similar relationship to our bivariate model we ran earlier.
+Sales representatives have the worst sleep quality by far with a strong t-statistic and p-value.
+Scientists, doctors, and sales persons are all about the next worst (with strong t-stats and p-values).
+
+The occupations that seem to have the best sleep are teachers, lawyers, engineers and nurses.
+This is somewhat surprising since these are all demanding jobs that can have decently stressful work environments.
+
+####Fitting a model using Sleep Quality as the outcome and BMI as a predictor.
 
 ```{r}
-#| echo: false
-2 * 2
+lmfit_quality_BMI <- lm(Quality.of.Sleep~ BMI.Category, sleepdata)
+
+#Placing results from lmfit_quality_BMI into a data frame with the tidy function
+
+lmtable_quality_BMI <- broom::tidy(lmfit_quality_BMI)
+
+#Viewing the results from the fourth model fit
+
+print(lmtable_quality_BMI)
+
+# Saving the lmfit_quality_BMI results table
+
+lmtable_quality_BMI = here("results", "tables", "lmfit4table.rds")
+
+saveRDS(lmtable_quality_BMI, file = lmtable_quality_BMI)
+
 ```
 
-The `echo: false` option disables the printing of code (only output is displayed).
+We can see that Obese individuals have a stronger negative relationship with Sleep Quality than Overweight individuals.
+The lower t-statistics and p-values are likely due to fewer observations for Obese individuals in the dataset.
+However, it is all still statistically significant and has reasonable standard errors.
+
+####Fitting a model using Sleep Quality as the outcome and Blood Pressure category as a predictor.
+```{r}
+lmfit_quality_BP <- lm(Quality.of.Sleep~ cat_bp, sleepdata)
+
+#Placing results from lmfit_quality_BP into a data frame with the tidy function
+
+lmtable_quality_BP <- broom::tidy(lmfit_quality_BP)
+
+#Viewing the results from the fifth model fit
+
+print(lmtable_quality_BP)
+
+# Saving the lmfit_quality_BP results table
+
+lmtable_quality_BP = here("results", "tables", "lmfit5table.rds")
+
+saveRDS(lmtable_quality_BP, file = lmtable_quality_BP)
+```
+
+We have some interesting results for this one.
+Our values aren't statistically significant according to p-values, but we can see that Stage 1 Hypertension seems to have the biggest negative impact on Quality of Sleep.
+This is surprising since we would expect more severe hypertension to result in worse sleep, but according to our model, Stage 2 hypertension has less impact than Stage 2.
+
+Normal blood pressure seems to have a small positive impact on sleep quality according to our model.
+All of these predictors seem to have a relatively small impact as they impact the sleep quality scale by less than one whole unit.
+Given our p-values and low t-statistics, however, this model should be taken with a grain of salt.
+
+####Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.
+```{r}
+lmfit_quality_stress <- lm(Quality.of.Sleep~ Stress.Level, sleepdata)
+
+#Placing results from lmfit_quality_stress into a data frame with the tidy function
+
+lmtable_quality_stress <- broom::tidy(lmfit_quality_stress)
+
+#Viewing the results from the sixth model fit
+
+print(lmtable_quality_stress)
+
+# Saving the lmfit_quality_stress results table
+
+lmtable_quality_stress = here("results", "tables", "lmfit6table.rds")
+
+saveRDS(lmtable_quality_stress, file = lmtable_quality_stress)
+```
+
+According to our model, Stress has a strongly defined relationship with sleep quality. We can see that a value of 0 on the stress score leads to a 10 unit increase in quality of sleep.
+
+The Stress Level coefficient indicates that as stress increases by one unit, sleep quality decreases by about 0.6 of a unit.
+The t-statistics and p-values for the values in this model are extremely strong, indicating a well-defined relationship between stress and sleep quality.
+
+####Fitting a model using Sleep Quality as the outcome and Age and Gender as predictors.
+```{r}
+lmfit_quality_gender_age <- lm(Quality.of.Sleep~ Gender + Age, sleepdata)
+
+#Placing results from lmfit_quality_gender_age into a data frame with the tidy function
+
+lmtable_quality_gender_age <- broom::tidy(lmfit_quality_gender_age)
+
+#Viewing the results from the seventh model fit
+
+print(lmtable_quality_gender_age)
+
+# Saving the lmfit_quality_gender_age results table
+
+lmtable_quality_gender_age = here("results", "tables", "lmfit7table.rds")
+
+saveRDS(lmtable_quality_gender_age, file = lmtable_quality_gender_age)
+```
+
+According to our table, Males seem to have lower sleep quality than females by about a score of .0336.
+However, the t-statistic is very low and the p-value is very poor.
+Gender doesn't seem to be a good predictor of sleep quality.
+
+When age is factored with gender, we can see that quality of sleep seems to increase slightly along with age.
+The t-statistic and p-value are both strong for this predictor; it may be worth doing a simple bivariate analysis of age.
+
+####Fitting a model using Sleep Quality as the outcome and Age as a predictor.
+```{r}
+lmfit_quality_age <- lm(Quality.of.Sleep~ Age, sleepdata)
+
+#Placing results from lmfit_quality_age into a data frame with the tidy function
+
+lmtable_quality_age <- broom::tidy(lmfit_quality_age)
+
+#Viewing the results from the eighth model fit
+
+print(lmtable_quality_age)
+
+# Saving the lmfit_quality_age results table
+
+lmtable_quality_age = here("results", "tables", "lmfit8table.rds")
+
+saveRDS(lmtable_quality_age, file = lmtable_quality_age)
+
+```
+
+Our model shows a strong relationship between age and quality of sleep.
+As Age increases, sleep quality seems to improve.
+Each year increase in age is predicted to have a 0.065 increase in sleep quality score.
+
+The intercept here is also interesting as an age of "0" is predicted to have a sleep score of 4.57, which is relatively poor. The t-statistics and p-values for these are both strong.
+
+####Fitting a model using Sleep Quality as the outcome and a newly grouped Occupatiin variable as the predictor.
+
+Now we'll create a new variable that groups the different occupations to make analysis more succinct.
+```{r}
+# Create a new column
+
+sleepdata$Occupation_Group <- as.character(sleepdata$Occupation)
+
+# Map occupations to broader categories
+
+sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Nurse', 'Doctor')] <- 'Healthcare'
+
+sleepdata$Occupation_Group[sleepdata$Occupation == 'Teacher'] <- 'Education'
+
+sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Software Engineer', 'Engineer')] <- 'Engineering'
+
+sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Accountant', 'Salesperson', 'Sales Representative', 'Manager')] <- 'Business/Finance'
+
+sleepdata$Occupation_Group[sleepdata$Occupation == 'Scientist'] <- 'Science'
+
+# Convert the new column to a factor
+
+sleepdata$Occupation_Group <- as.factor(sleepdata$Occupation_Group)
+
+## This should help with analysis of this variable. We'll fit this to a model now to test it.
+
+lmfit_quality_occupation <- lm(Quality.of.Sleep~ Occupation_Group, sleepdata)
+
+#Placing results from lmfit_quality_occupation into a data frame with the tidy function
+
+lmtable_quality_occupation <- broom::tidy(lmfit_quality_occupation)
+
+#Viewing the results from the occupation model fit
+
+print(lmtable_quality_occupation)
+
+# Saving the lmfit_quality_gender_age results table
+
+lmtable_quality_occupation = here("results", "tables", "lmfit9table.rds")
+
+saveRDS(lmtable_quality_occupation, file = lmtable_quality_occupation)
+
+```
+
+Looking at our newly grouped occupations, we can see that our intercept has a high t-statistic and significance; this indicates that our reference category of Business/Finance jobs have a sleep quality score of about 7, which is pretty good.
+
+Teaching and healthcare professions don't seem to have a significant impact on an individual's sleep quality score. They have low t-statistics and large p-values.
+
+Engineers seem to have the best sleep quality score; they're predicted to have a score of about 8.3 (with strong p-values and t-statistics).
+Scientists seem to have the worst score; they're predicted to have 2 points lower sleep quality than our intercept.
+The t-statistic and p-value are both strong for this group.
+
+Lawyers fall in the middle, having a decent increase in sleep quality compared to the intercept and solid p-values and t-statistics to support this.
+
+Now we'll create a table that combines all of these results to put in the manuscript.
+
+```{r}
+# Create a list of all models
+
+models <- list(
+
+PhysicalActivity = lmfit_quality_activity,
+
+Occupation = lmfit_quality_occupation,
+
+Age = lmfit_quality_age,
+
+StressLevel = lmfit_quality_stress,
+
+Hypertension = lmfit_quality_BP,
+
+SleepDuration = lmfit_quality_duration,
+
+BMI = lmfit_quality_BMI,
+
+Gender_and_Age = lmfit_quality_gender_age
+
+)
+
+# Create an empty data frame to store results
+
+all_results <- data.frame()
+
+# Loop through each model, tidy the results, and bind to all_results
+
+for(i in 1:length(models)) {
+
+model_name <- names(models)[i]
+
+model_results <- broom::tidy(models[[i]])
+
+model_results$model <- model_name # add a column to identify the model
+
+all_results <- rbind(all_results, model_results)
+
+}
+
+# Now all_results contains the combined results of all models
+
+print(all_results)
+
+# Save the combined results table
+
+all_results_path = here("results", "tables", "all_results.rds")
+
+saveRDS(all_results, file = all_results_path)
+
+# Transform p-values to scientific notation to make table cleaner to read
+
+all_results$p.value <- format(all_results$p.value, scientific = TRUE, digits = 2)
+
+# Create a flextable
+
+ft <- flextable::flextable(all_results)
+
+# Print the table
+
+print(ft)
+
+# Save the flextable as an image
+
+flextable::save_as_image(ft, path = here("results", "tables", "table.png"))
+
+# Now we'll remove unnecessary variables (e.g. ones we've feature engineered into more useful ones) to make our analyses easier.
+
+sleepdatafinal <- subset(sleepdata, select = -c(Occupation, systolic, diastolic, Daily.Steps, Physical.Activity.Level, Heart.Rate, AgeGroup))
+```
+
+# Creating a collinearity plot
+
+We'll make sure there isn't collinearity between our continuous variables before we fit more complex models.
+```{r}
+# Select the variables
+
+continuous_vars <- sleepdatafinal[, c("Age", "Sleep.Duration", "Quality.of.Sleep", "Stress.Level")]
+
+# Compute correlation matrix
+
+correlation_matrix <- cor(continuous_vars)
+
+# Create a pairwise correlation plot (using corrplot)
+
+correlation <- corrplot(correlation_matrix, method = "circle")
+
+#Save the correlation plot
+
+correlation = here("results", "figures", "correlation.rds")
+
+saveRDS(correlation, file = correlation)
+```
+
+It seems that stress level and quality of sleep are at an absolute value of 0.9 or higher on the correlation scale.
+Sleep Duration and stress level also seem closely tied.
+Stress level and quality of sleep are both subjective scales given by an individual; this is interesting to note.
+
+We're definitely still interested in stress as a predictor as previous work has clearly identified it as a key factor in sleep quality.
+We'll keep it in our analysis, but it seems good to note that it has a very linear relationship with sleep quality.
+
+# Fitting a Random Forest model using cross validation
+
+First, we'll create a random seed to aid in reproducibility.
+```{r}
+#Set the seed
+rngseed=1234
+
+set.seed(rngseed)
+```
+
+We'll now train a Random Forest model using CV with no train/test split (as the data has less than 400 observations and many of the values are rare/unique for some columns).
+We'll follow the tidymodels framework.
+
+```{r}
+# Random Forest model specification; include the mtry and min_n parameters
+
+model_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 300) %>%
+
+set_engine("ranger", importance = 'impurity', seed = rngseed) %>%
+
+set_mode("regression")
+
+# Specify the recipe
+
+data_recipe <- recipe(Quality.of.Sleep ~ ., data = sleepdatafinal)
+
+# Create Random Forest workflow
+
+data_workflow <- workflow() %>%
+
+add_model(model_spec) %>%
+
+add_recipe(data_recipe)
+
+# Create resamples using 5-fold cross-validation, 5 times repeated
+
+resamples <- vfold_cv(sleepdatafinal, v = 5, repeats = 5)
+
+# Define the grid of parameters
+
+mtry_param <- mtry(range = c(1, 7))
+
+min_n_param <- min_n(range = c(1, 21))
+
+rf_grid <- grid_regular(mtry_param, min_n_param, levels = 7)
+
+# Tune the Random Forest model
+
+tune_results_rf <- tune_grid(
+
+data_workflow,
+
+resamples = resamples,
+
+grid = rf_grid
+
+)
+
+# Print RF results
+
+print(tune_results_rf)
+
+#Plot them too
+
+autoplot(tune_results_rf)
+```
+
+We've successfully fit our RF model; now, we can pick the different pieces apart and see what predictors seem the most important and which model is best.
+
+####Best RMSE model
+
+```{r}
+best_params <- select_best(tune_results_rf, metric = "rmse")
+
+#Variable importance info
+
+# Finalize the workflow with the best parameters
+
+final_wf <- finalize_workflow(data_workflow, best_params)
+
+# Fit the finalized workflow to the entire data
+
+final_fit <- fit(final_wf, data = sleepdatafinal)
+
+# Extract the fitted model
+
+fitted_model <- extract_fit_parsnip(final_fit)
+
+# Create a variable importance plot
+
+vip(fitted_model$fit)
+
+```
+
+We can see that sleep duration, stress level, and Age seem to have the most importance for our model.
+
+PhysicalActivityLevel doesn't seem to have much of an impact on Quality of Sleep, as it was excluded from the variable importance plot altogether.
+These results are pretty in line with what we discovered with our EDA and our simple linear models, but now we have a ranking given by this plot.
+
+Now, we'll see how well this model predicts on new data.
+
+```{r}
+# Make predictions
+
+predictions <- predict(final_fit, new_data = sleepdatafinal)
+
+# Add the predictions to dataframe
+
+sleepdatafinal$Predicted <- predictions$.pred
+
+# Create an observed vs. predicted plot
+
+ggplot(sleepdatafinal, aes(x = Quality.of.Sleep, y = Predicted)) +
+
+geom_point() +
+
+geom_abline(color = "red") +
+
+labs(x = "Observed", y = "Predicted", title = "Observed vs. Predicted Plot") +
+
+theme_minimal()
+
+#Now we'll calculate the RMSE and R squared values of our "best" set of parameters
+
+# Calculate residuals
+
+residuals <- sleepdatafinal$Quality.of.Sleep - predictions$.pred
+
+# Calculate RMSE
+
+rf_rmse <- sqrt(mean(residuals^2))
+
+# Calculate R squared
+
+sst <- sum((sleepdatafinal$Quality.of.Sleep - mean(sleepdatafinal$Quality.of.Sleep))^2)
+
+ssr <- sum(residuals^2)
+
+rf_r_squared <- 1 - (ssr / sst)
+
+#print results
+
+print(rf_rmse)
+
+print(rf_r_squared)
+```
+
+This seems like a very solid fit; not too close to where overfitting is obvious, but not too far to where the predictions aren't useful either.
+The RMSE is extremely low (.059) and the R squared is extremely high (0.997).
+This is a good indicator that our selected model is powerful.
+
+#### Random Forest Model Evaluation
+
+Now, we'll make a "null" model and compare its results to that of our RF in order to make sure that the RF is useful.
+```{r}
+# Specify the null model
+
+null_mod <- null_model() %>%
+
+set_engine("parsnip") %>%
+
+set_mode("regression")
+
+# Fit the null model to the data
+
+null_fit <- fit(null_mod, Quality.of.Sleep ~ 1, data = sleepdatafinal)
+
+# Make predictions
+
+null_predictions <- predict(null_fit, new_data = sleepdatafinal)
+
+# Bind the predictions with the actual values
+
+data_with_predictions <- bind_cols(sleepdatafinal, null_predictions)
+
+# Compute RMSE
+
+null_rmse <- rmse(data_with_predictions, truth = Quality.of.Sleep, estimate = .pred)
+
+# Print the RMSE
+
+print(null_rmse)
+
+```
+
+This RMSE is a good bit higher than the 0.2-0.32 range of all of our Random Forest models. This is a good sign that the RF model performs well on the data.
+
+```{r}
+# Compute R-squared
+
+r_squared <- rsq(data_with_predictions, truth = Quality.of.Sleep, estimate = .pred)
+
+# Print the R-squared
+
+print(r_squared)
+
+# Create a predicted v observed plot
+
+ggplot(data_with_predictions, aes(x = Quality.of.Sleep, y = .pred)) +
+
+geom_point() +
+
+geom_abline(color = "red") +
+
+labs(x = "Observed", y = "Predicted", title = "Observed vs Predicted Null Values") +
+
+theme_minimal()
+
+```
+
+Since there is no variation between the model predictions and the observed values in a null model, the R squared is not possible to generate.
+The observed vs predicted plot is also as expected; a straight line of predictions that all match the average sleep quality score.
+
+Now we'll make a table for the manuscript comparing the metrics we evaluated.
+
+```{r, message=FALSE}
+#extract the metric values for the null model from the objects they're stored in
+
+null_rmse_value <- null_rmse$.estimate
+
+null_r_squared_value <- r_squared$.estimate
+
+# Create a data frame
+
+results <- data.frame(
+
+Model = c("Random Forest", "Null Model"),
+
+`R-squared` = c(rf_r_squared, null_r_squared_value),
+
+RMSE = c(rf_rmse, null_rmse_value)
+
+)
+
+# Create the table
+
+gt_table <- gt(results)
+
+# Print the table
+
+print(gt_table)
+```

### 2024-04-29 products/manuscript/Manuscript.qmd
Final Part Revisions
Made statistical analysis file a .qmd and have cleaned some issues pointed out in the project reviews.

@@
 author: Kevin Kosewick and Emma Hardin-Parker
-date: "4-18-2024"
+date: "4-29-2024"
 format:
@@
 For our project, we obtained a dataset from
-[Kaggle](https://www.kaggle.com/datasets) containing 373 observations of
+Kaggle.com containing 373 observations of
 13 variables from a study on different factors that may affect sleep
@@
 150 minutes total) is the optimal weekly regimen to improve sleep
-quality and reduce stress. We expect to see similar trends with our
-data, where increased stress levels negatively impact sleep quality and
-sleep duration, while noticing a positive association between increased
-physical activity levels and sleep quality.
+quality and reduce stress. We expect to see similar trends with our data
+where increased stress levels negatively impact sleep quality. We expect to see
+a positive association between increasing physical activity levels and
+sleep quality until the point of "too much" activity.
 
@@
 information on how or where this data was collected, so we solely relied
-on the codebook for variable definitions. There were no missing values
-or erroneous variables in the dataset that we needed to remove. The
-"Blood.Pressure" variable was transformed to reflect the individual
+on the codebook for variable definitions. We have added this codebook to
+the supplementary materials file for ease of access but you can also
+find it on the Kaggle page for this dataset. There were no missing
+values or erroneous variables in the dataset that we needed to remove.
+The "Blood.Pressure" variable was transformed to reflect the individual
 systolic and diastolic blood pressure measurements from each subject. We
@@
 categorical activity level of subjects in the dataset based on these
-guidelines provided by \[10,000 steps.org\]
-(https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:\~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
+guidelines provided by [10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
 Finally, we created an additional categorical variable called
@@
 RF models is difficulty of interpretation; we attempted to minimize this
-by creating easy to read tables and graphs for the results. Future
+by creating easy to read tables and graphs for the results.
+We chose not to split our data into a train/test split due to the small number of observations and presence of rare values in our data.
+We believe that fitting using CV minimized the downsides of this.
+Future
 analyses would benefit from more detailed meta data and a better

### 2024-04-29 products/manuscript/supplement/Supplementary-Material.qmd
Final Part Revisions
Made statistical analysis file a .qmd and have cleaned some issues pointed out in the project reviews.

@@
 
-*This shows some materials that could go into a supplementary file.
-Often you want/need references here too.
-You can use the same reference bib file for this and the main text (as done here) or have separate bib files.*
-
-*For illustrative purposes, I'm doing the supplement as pdf.
-For this to work, you need a (La)TeX system installed.
-It's easy.
-Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).*
-
-*I'm also using a different style for the references here.
-(vancouver vs apa in the main manuscript).
-Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header.
-We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.*
-
-### I HAVE STRUCTURED THIS FILE. NEED TO WORK ON PULLING CERTAIN FIGURES, PLOTS, TABLES, ETC. ###
-
 # Overview
@@
 *Include ALL exploratory analysis figures*
+
+The following information is the codebook for this dataset:
 
@@
 ```
-
-*Example additional result*
-
-*@tbl-resulttable1 shows an additional table summarizing a model fit.*
-
-```{r}
-#| label: tbl-resulttable1
-#| tbl-cap: "Another fit table."
-#| echo: FALSE
-resulttable1 = readRDS(here("results","resulttable1.rds"))
-knitr::kable(resulttable1)
-```
-
-@fig-result2 shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result2
-#| fig-cap: "Height and weight."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight.png"))
-```
-
-{{< pagebreak >}}
-
-*# Discussion*
-
-Any additional discussion regarding the supplementary material/findings.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-{{< pagebreak >}}
-
-# References

### 2024-04-30 products/manuscript/Manuscript.qmd
emma- checking refs

@@
     highlight-style: github
-bibliography: ../manuscript/references_sleepdata.bib
+bibliography: ../../assets/references_sleepdata.bib
+csl: ../../assets/apa.csl
 editor: 

### 2024-05-01 R/analysis-code/statistical-analysis.qmd
References Check

@@
 library(gt)
+library(glmnet)
 ```
@@
 ####Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
+
 ```{r}
@@
 #### Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors
+
 ```{r}
@@
 ####Fitting a model using Sleep Quality as the outcome and Blood Pressure category as a predictor.
+
 ```{r}
@@
 ####Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.
+
 ```{r}
@@
 ####Fitting a model using Sleep Quality as the outcome and Age and Gender as predictors.
+
 ```{r}
@@
 ####Fitting a model using Sleep Quality as the outcome and Age as a predictor.
+
 ```{r}
@@
 Now we'll create a new variable that groups the different occupations to make analysis more succinct.
+
 ```{r}
@@
 We'll make sure there isn't collinearity between our continuous variables before we fit more complex models.
+
 ```{r}
@@
 First, we'll create a random seed to aid in reproducibility.
+
 ```{r}
@@
 Now, we'll make a "null" model and compare its results to that of our RF in order to make sure that the RF is useful.
+
 ```{r}
@@
 
-null_fit <- fit(null_mod, Quality.of.Sleep ~ 1, data = sleepdatafinal)
+null_fit <- fit(null_mod, Quality.of.Sleep ~ ., data = sleepdatafinal)
 
@@
 
-Now we'll make a table for the manuscript comparing the metrics we evaluated.
+# Fitting a LASSO Model using cross validation
 
-```{r, message=FALSE}
-#extract the metric values for the null model from the objects they're stored in
+First, we'll create a random seed to aid in reproducibility.
 
-null_rmse_value <- null_rmse$.estimate
+```{r}
+#Set the seed
+rngseed= 3456
 
-null_r_squared_value <- r_squared$.estimate
+set.seed(rngseed)
+```
 
-# Create a data frame
+We'll now train a LASSO model using CV with no train/test split (as the data has less than 400 observations and many of the values are rare/unique for some columns).
+We'll follow the tidymodels framework.
 
-results <- data.frame(
+```{r}
+# Recipe preparation
+data_recipe <- recipe(Quality.of.Sleep ~ ., data = sleepdatafinal) %>%
+  step_dummy(all_nominal(), -all_outcomes())  # One-hot encode categorical variables
 
-Model = c("Random Forest", "Null Model"),
+# Prepare the recipe
+prepared_recipe <- prep(data_recipe)
 
-`R-squared` = c(rf_r_squared, null_r_squared_value),
+# Apply the prepared recipe to transform your data
+processed_data <- bake(prepared_recipe, new_data = sleepdatafinal)
 
-RMSE = c(rf_rmse, null_rmse_value)
+# Define the LASSO model
+lasso_model <- linear_reg(penalty = tune(), mixture = 1) %>%
+  set_engine("glmnet")
 
+# Set up a grid of values to tune over
+lasso_grid <- tibble(penalty = 10^seq(-5, 2, length.out = 50))
+
+# Create a workflow
+lasso_wf <- workflow() %>%
+  add_recipe(data_recipe) %>%
+  add_model(lasso_model)
+
+# Create resamples using 5-fold cross-validation, 5 times repeated
+
+resamples2 <- vfold_cv(sleepdatafinal, v = 5, repeats = 5)
+
+# Tune the LASSO model using cross-validation
+lasso_tune <- tune_grid(
+  lasso_wf,
+  resamples = resamples2,  
+  grid = lasso_grid,
+  metrics = metric_set(rmse), 
+  control = control_grid(save_pred = TRUE)
 )
 
-# Create the table
+# Plot results
+autoplot(lasso_tune)
 
-gt_table <- gt(results)
+# Select the best hyperparameters
+best_param <- select_best(lasso_tune, metric = "rmse")
 
-# Print the table
+# Finalize the workflow with the best hyperparameters
+final_wf2 <- finalize_workflow(lasso_wf, best_param)
 
-print(gt_table)
+# Fit the final LASSO model using the entire dataset
+final_fit2 <- fit(final_wf2, data = sleepdatafinal)
+
+# Extract the fitted LASSO model
+fitted_model2 <- extract_fit_parsnip(final_fit2)
+
+# Create the variable importance plot
+vip(fitted_model2$fit)
+
 ```
+*Kevin: Please check this*
+
+#### Predicting on new data
+
+```{r}
+# Make predictions
+
+predictions2 <- predict(final_fit2, new_data = sleepdatafinal)
+
+# Add the predictions to dataframe
+
+sleepdatafinal$Predicted <- predictions2$.pred
+
+# Create an observed vs. predicted plot
+
+ggplot(sleepdatafinal, aes(x = Quality.of.Sleep, y = Predicted)) +
+
+geom_point() +
+
+geom_abline(color = "red") +
+
+labs(x = "Observed", y = "Predicted", title = "Observed vs. Predicted Plot") +
+
+theme_minimal()
+
+#Now we'll calculate the RMSE and R squared values of our "best" set of parameters
+
+# Calculate residuals
+
+residuals2 <- sleepdatafinal$Quality.of.Sleep - predictions2$.pred
+
+# Calculate RMSE
+
+lasso_rmse <- sqrt(mean(residuals2^2))
+
+# Calculate R squared
+
+sst2 <- sum((sleepdatafinal$Quality.of.Sleep - mean(sleepdatafinal$Quality.of.Sleep))^2)
+
+ssr2 <- sum(residuals2^2)
+
+lasso_r2 <- 1 - (ssr2 / sst2)
+
+#print results
+
+print(lasso_rmse)
+
+print(lasso_r2)
+```
+The Observed vs.
+Predicted plot looks pretty good!
+Our RMSE value was once again quite low, at 0.0557 and our R-squared value was high at 0.997.
+Both of these metrics, combined with the Observed vs.
+Predicted plot indicate that the LASSO model is performing well on new data.

### 2024-05-01 products/manuscript/Manuscript.qmd
References Check

--- 2024-04-29 products/manuscript/Manuscript.qmd
@@
     highlight-style: github
-bibliography: ../manuscript/references_sleepdata.bib
+bibliography: ../../assets/dataanalysis-references.bib
+csl: ../../assets/apa.csl 
 editor: 

### 2024-05-01 products/manuscript/Manuscript.qmd
Ref Check pt. 2000

--- 2024-04-29 products/manuscript/Manuscript.qmd
@@
     highlight-style: github
-bibliography: ../manuscript/references_sleepdata.bib
+bibliography: dataanalysis-references.bib
+csl: apa.csl 
 editor: 
@@
 categorical activity level of subjects in the dataset based on these
-guidelines provided by [10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
+guidelines provided by [10,000 steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
 Finally, we created an additional categorical variable called

### 2024-05-01 products/manuscript/Manuscript.qmd
Ref Troubleshoot

@@
     highlight-style: github
-bibliography: dataanalysis-references.bib
-csl: apa.csl 
+bibliography: ../../assets/sleepref2.bib
+csl: ../../assets/apa.csl 
 editor: 

### 2024-05-02 R/analysis-code/statistical-analysis.qmd
Supplementary file and table modifications
Formatted the supplementary file and cleaned the linear model table in the manuscript.

@@
 # Loop through each model, tidy the results, and bind to all_results
-
 for(i in 1:length(models)) {
-
   model_name <- names(models)[i]
-
   model_results <- broom::tidy(models[[i]])
 
+  # Round the estimate and std.error to 2 decimal places
+  model_results$estimate <- round(model_results$estimate, 2)
+  model_results$std.error <- round(model_results$std.error, 2)
+  model_results$statistic <- round(model_results$statistic, 3)  # Round to 3 decimal places
+
   model_results$model <- model_name # add a column to identify the model
-
   all_results <- rbind(all_results, model_results)
-
 }
 
-# Now all_results contains the combined results of all models
-
-print(all_results)
-
-# Save the combined results table
-
-all_results_path = here("results", "tables", "all_results.rds")
-
-saveRDS(all_results, file = all_results_path)
-
 # Transform p-values to scientific notation to make table cleaner to read
-
 all_results$p.value <- format(all_results$p.value, scientific = TRUE, digits = 2)
 
+# Create a named vector of new term names for categorical variables
+new_term_names <- c(
+  "PhysicalActivity" = "PhysicalActivity.CategoryHigh(Reference: Low)",
+  "Occupation_GroupEducation" = "Occupation_Education(Reference: BusinessFinance)",
+  "Occupation_GroupEngineering" = "Occupation_Engineering(Reference: BusinessFinance)",
+  "Occupation_GroupHealthcare" = "Occupation_Healthcare(Reference: BusinessFinance)",
+  "Occupation_GroupLawyer" = "Occupation_Lawyer(Reference: BusinessFinance)",
+  "Occupation_GroupScience" = "Occupation_Science(Reference: BusinessFinance)",
+  "cat_bpStage 1 Hypertension" = "Stage_1_Hypertension(Reference: Elevated)",
+  "cat_bpStage 2 Hypertension" = "Stage_2_Hypertension(Reference: Elevated)",
+  "cat_bpNormal" = "Normal(Reference: Elevated)",
+  "BMI.CategoryObese" = "BMI.CategoryObese(Reference: Normal)",
+  "BMI.CategoryOverweight" = "BMI.CategoryOverweight(Reference: Normal)",
+  "GenderMale" = "GenderMale(Reference: Female)"
+  # Add more terms as needed
+)
+
+# Update the term names in the results
+all_results$term <- ifelse(all_results$term %in% names(new_term_names), 
+                           new_term_names[all_results$term], 
+                           all_results$term)
+
 # Create a flextable
-
 ft <- flextable::flextable(all_results)
@@
 # Print the table
-
 print(ft)
@@
 # Save the flextable as an image
-
 flextable::save_as_image(ft, path = here("results", "tables", "table.png"))
@@
 ```
+
 *Kevin: Please check this*
@@
 ```
+
 The Observed vs.

### 2024-05-02 R/eda-processing-code/eda.qmd
Cleaning folders and updating background section
Removed unnecessary files/folders and did some manuscript updating.
Supplementary file and table modifications
Formatted the supplementary file and cleaned the linear model table in the manuscript.

@@
 We intend to perform any needed cleaning and mutation to the data as well as creating exploratory plots and tables to gain a better understanding of the data.
-As stated in the manuscript, we are looking to determine which variables or identifiers have significant impacts on sleep quality.
+We are looking to determine which variables or identifiers have significant impacts on sleep quality.
 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE}
 #Loading necessary libraries
@@
 #save the processed data separately from the original
-saveRDS(sleepdata, file = "sleepdataprocessed.rds")
+saveRDS(sleepdata, file = here("data", "processed-data", "sleepdataprocessed.rds"))
 ```

### 2024-05-02 products/manuscript/Manuscript.qmd
Supplementary file and table modifications
Formatted the supplementary file and cleaned the linear model table in the manuscript.

@@
 
-For our project, we obtained a dataset from
-Kaggle.com containing 373 observations of
-13 variables from a study on different factors that may affect sleep
-quality. Variables of interest from the study included physical activity
-level, age, sex, occupation, stress, gender, BMI, blood pressure, and
-sleep duration. In [RStudio](https://posit.co/download/rstudio-desktop/)
-(version 4.3.1), we used a combination of packages to create models
-suitable for answering our primary question: which variables are most
-important for determining an individual's sleep quality and what effect
-do they have? After conducting an exploratory data analysis (EDA) we
-hypothesized that stress, sleep duration, and occupation would have the
-largest impact on sleep quality. We believed that stress and sleep
-quality would be negatively correlated, sleep duration and quality would
-be positively correlated, and that jobs with longer or more variable
-hours would negatively affect sleep quality.
+For our project, we obtained a dataset from Kaggle.com containing 373
+observations of 13 variables from a study on different factors that may
+affect sleep quality. Variables of interest from the study included
+physical activity level, age, sex, occupation, stress, gender, BMI,
+blood pressure, and sleep duration. In
+[RStudio](https://posit.co/download/rstudio-desktop/) (version 4.3.1),
+we used a combination of packages to create models suitable for
+answering our primary question: which variables are most important for
+determining an individual's sleep quality and what effect do they have?
+After conducting an exploratory data analysis (EDA) we hypothesized that
+stress, sleep duration, and occupation would have the largest impact on
+sleep quality. We believed that stress and sleep quality would be
+negatively correlated, sleep duration and quality would be positively
+correlated, and that jobs with longer or more variable hours would
+negatively affect sleep quality.
 
@@
 quality and reduce stress. We expect to see similar trends with our data
-where increased stress levels negatively impact sleep quality. We expect to see
-a positive association between increasing physical activity levels and
-sleep quality until the point of "too much" activity.
+where increased stress levels negatively impact sleep quality. We expect
+to see a positive association between increasing physical activity
+levels and sleep quality until the point of "too much" activity.
 
@@
 categorical activity level of subjects in the dataset based on these
-guidelines provided by [10,000 steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
+guidelines provided by [10,000
+steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
 Finally, we created an additional categorical variable called
@@
 sleep quality. The "estimate" column shows the predicted change in sleep
-quality score as a result of each associated
-predictor.](images/table.png)
+quality score as a result of each associated predictor. Each reference
+category that is used for categorical variables is included next to the
+term in parentheses (e.g. Reference: Female).](images/table-01.png)
 
@@
 was at a value of 0. For categorical variables, it represents the
-reference category (e.g. the intercept for the Gender model represents
-the expected sleep quality score for females).
+reference category that the others are being compared to.
 
@@
 RF models is difficulty of interpretation; we attempted to minimize this
-by creating easy to read tables and graphs for the results.
-We chose not to split our data into a train/test split due to the small number of observations and presence of rare values in our data.
-We believe that fitting using CV minimized the downsides of this.
-Future
-analyses would benefit from more detailed meta data and a better
-understanding of the experimental design that generated our
-observations.
+by creating easy to read tables and graphs for the results. We chose not
+to split our data into a train/test split due to the small number of
+observations and presence of rare values in our data. We believe that
+fitting using CV minimized the downsides of this. Future analyses would
+benefit from more detailed meta data and a better understanding of the
+experimental design that generated our observations.
 

### 2024-05-02 products/manuscript/supplement/Supplementary-Material.qmd
Supplementary file and table modifications
Formatted the supplementary file and cleaned the linear model table in the manuscript.

--- 2024-03-14 products/manuscript/supplement/Supplementary-Material.qmd
@@
 ---
-title: "Supplement to Example Manuscript Template for a Data Analysis Project"
-
+title: "Supplementary Material"
 subtitle: ""
@@
 format:
-  pdf:
+  docx:
     toc: false
@@
     highlight-style: github
-bibliography: ../../assets/dataanalysis_template_references.bib
-csl: ../../assets/vancouver.csl
+editor: 
+  markdown: 
+    wrap: sentence
 ---
 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # load a few R packages
@@
 
-For now, this document is a placeholder with the original text that came with the template for this project.
-All text beyond here is not relevant to our project in particular and will be updated when we have supplementary materials to include.
-
-This shows some materials that could go into a supplementary file.
-Often you want/need references here too.
-You can use the same reference bib file for this and the main text (as done here) or have separate bib files.
-
-For illustrative purposes, I'm doing the supplement as pdf.
-For this to work, you need a (La)TeX system installed.
-It's easy.
-Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).
-
-Of course you would choose the format based on needs.
-
-I'm also using a different style for the references here.
-(vancouver vs apa in the main manuscript).
-Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header.
-We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.
-
 # Overview
 
-A quick overview of what readers can find in the supplement.
+This document includes additional information about the data and all of our interesting EDA plots.
+The full code and process of creating each plot can be found in our "eda-processing-code" folder at the [GitHub Repository](https://github.com/KKosewick/KOSEWICK-HARDIN-PARKER-MADA-project).
 
-# Code and file information
+# Code & File Details and Locations
 
-Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
-Essentially, give a full set of instructions to re-generate everything.
+The eda.qmd file, located in the R folder and subsequently in the eda-processing-code folder contains all data exploration, processing, and transformation.
+The statistical-analysis.R file, found in the R folder and subsequently in the analysis-code folder holds all statistical analysis and machine learning code.
+
+# Exploratory Data Analysis
+
+The data was obtained from kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
+This page contains a detailed codebook and all available information about the data set.
+The following information is the codebook for this dataset:
+
+The data set contains 373 observations for 13 variables.
+Some variables are not easily understandable at face value, so we went back to the data source and accessed the data dictionary.
+The variables are defined as follows:
+
+Person ID: A unique identifier for each participant.
+Gender: The participant's gender (Male/Female).
+Age: The participant's age in years.
+Occupation: The job or career of the participant.
+Sleep Duration (hours): The daily sleep duration of the participant in hours.
+Quality of Sleep (scale: 1-10): A subjective assessment of sleep quality on a scale from 1 to 10.
+Physical Activity Level (minutes/day): The daily duration of physical activity for the participant, measured in minutes.
+Stress Level (scale: 1-10): A subjective assessment of the participant's stress level on a scale from 1 to 10.
+BMI Category: The participant's BMI classification (e.g., Underweight, Normal, Overweight).
+Blood Pressure (systolic/diastolic): The participant's blood pressure, represented as systolic pressure over diastolic pressure.
+Heart Rate (bpm): The participant's resting heart rate, measured in beats per minute.
+Daily Steps: The number of steps the participant takes each day.
+Sleep Disorder: The presence or absence of a sleep disorder in the participant (None, Insomnia, Sleep Apnea).
+
+# Data Transformation
+
+We created a categorical hypertension variable from our current continuous blood pressure variable.
+We used blood pressure guidelines from the American Heart Association to do this.
+Our goal with this was to see how blood pressure can affect sleep quality in a more digestible way.
+We have attached the guidelines here: [AHA Guidelines](https://www.heart.org/en/health-topics/high-blood-pressure)
+
+![](images/AHA_Guidelines.png){fig-align="center"}
+
+To make it easier to visualize Physical Activity, we created a new variable with categories of Low (0-30 minutes), Medium (30-60), and High (60+).
+We also created a new variable with groups according to the guidelines here at [10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500).
+Nobody in the dataset goes over 10000 steps a day, however, so we won't see any data for the "Active" or "Highly Active" groups.
+Lastly, we created a new variable that groups the different occupations to make analysis more succinct.
 
@@
 
-# Additional Method Details
+# Data Visualization and Exploratory Data Analysis
 
-Often, the main manuscript only allows for an overview description of the methods.
-Use the supplement to describe all your methods, models and approaches in a lot of detail.
-Reference specific parts of your code as needed.
+We have included the plots generated by our EDA here alongside some interpretations.
+These are all of the preliminary findings that guided our feature engineering and variable selection steps for our final models.
+
+![Figure 1: A boxplot comparing the average quality of sleep between males and females in our dataset.](images/unnamed-chunk-7-1.png)
+
+![Figure 2: A bar plot showing the mean quality of sleep by age group. We eventually decided to ungroup the ages for our final model fitting processes.](images/unnamed-chunk-7-2.png)
+
+![Figure 3: A bar plot showing the mean quality of sleep across the occupations in our dataset.
+We eventually decided to group these into broader categories for our final model fitting.](images/unnamed-chunk-7-3.png)
+
+We can see some clear differences in self-reported Quality of Sleep between these different groups.
+Females tend to report an average of 1 higher according to the boxplot.
+The different age groups are similar, but the oldest (50-60) reports the highest quality of sleep by far.
+Our ungrouped occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
 
@@
 
-# Additional results
+![Figure 4: A boxplot of the feature engineered categorical variable for physical activity levels.](images/unnamed-chunk-8-2.png)
 
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
-
-## Example additional result
-
-@tbl-resulttable1 shows an additional table summarizing a model fit.
-
-```{r}
-#| label: tbl-resulttable1
-#| tbl-cap: "Another fit table."
-#| echo: FALSE
-resulttable1 = readRDS(here("results","resulttable1.rds"))
-knitr::kable(resulttable1)
-```
-
-@fig-result2 shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result2
-#| fig-cap: "Height and weight."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight.png"))
-```
+We then explored Physical Activity Level as a predictor of sleep quality using our engineered categories. We see a positive correlation between increasing physical activity and sleep quality here.
 
@@
 
-# Discussion
+![Figure 5: A boxplot of daily steps as a predictor of quality of sleep. We used the different categories as previously described.](images/unnamed-chunk-9-2.png)
 
-Any additional discussion regarding the supplementary material/findings.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
+We then explored Daily Steps with a box plot of our new engineered categorical variable.
+Daily Steps don't have as clear of a trend as Physical Activity did, but there do seem to be fewer low Quality of Sleep scores in higher daily steps groups.
 
@@
 
-# References
+![Figure 6: A scatterplot for sleep duration as a predictor of quality of sleep.](images/unnamed-chunk-10-2.png)
+
+We then explored our Sleep Duration variable.
+According to [Medical News Today](https://www.medicalnewstoday.com/articles/sleep-cycle-stages#sleep-requirements), adults aged 18-61 are recommended 7+ hours of sleep a night.
+We kept this in mind when looking at our data. We can see a clear and strong positive correlation between Sleep Duration and Quality of Sleep.
+
+{{< pagebreak >}}
+
+![Figure 7: A bar plot of our categorical blood pressure variable.](images/unnamed-chunk-11-1.png)
+
+![Figure 8: A box plot of hypertension category as a predictor of quality of sleep.](images/unnamed-chunk-11-2.png)
+
+We then looked at the new variable we created with categories of Blood Pressure. We can see that elevated and normal blood pressure have the least observations.
+Most individuals fall under the category of "Stage 1 Hypertension" as defined by the AHA.
+We can't see a clear relationship between our outcome and predictor based off of these plots.
+
+{{< pagebreak >}}
+
+![Figure 9: A bar plot showing the counts of our different BMI categories in the data.](images/unnamed-chunk-12-1.png)
+
+![Figure 10: A box plot of BMI category as a predictor of sleep quality.](images/unnamed-chunk-12-2.png)
+
+We then explored BMI with some plots.
+We can clearly see that Normal weights seem to have better sleep on average than the obese and overweight categories.
+However, there are very few observations in the obese category, so conclusions with that group may not be as easy to support.
+
+{{< pagebreak >}}
+
+![Figure 11: A scatterplot of Stress Level as a predictor for sleep quality.](images/unnamed-chunk-13-2.png)
+
+We then looked at Stress Level as a predictor for quality of sleep. Our scatterplot shows a clear negative correlation between Stress Level and Quality of Sleep.
+
+{{< pagebreak >}}
+
+![Figure 12: A scatterplot of Heart Rate as a predictor for quality of sleep.](images/unnamed-chunk-14-2.png)
+
+We then looked at Heart Rate's relationship to Quality of Sleep. We can see a bit of a skewed relationship between the two variables.
+Healthy Heart Rates tend to differ between each individual, making it difficult to draw conclusions about this variable.
+
+{{< pagebreak >}}
+
+![Figure 13: A bar plot showing the count of sleep disorders in our data.](images/unnamed-chunk-15-1.png)
+
+![Figure 14: A box plot of Sleep Disorder as a predictor of quality of sleep.](images/unnamed-chunk-15-2.png)
+
+The last variable we investigated was Sleep Disorders.
+We can see that individuals without a disorder tend to have higher sleep scores than those with Insomnia or Sleep Apnea, which is what we'd expect to see.
+There are far more observations of individuals without any disorders than of either Sleep Apnea or Insomnia.

### 2024-05-02 products/manuscript/Manuscript.qmd
check

@@
     highlight-style: github
-bibliography: ../../assets/sleepref2.bib
-csl: ../../assets/apa.csl 
+bibliography: sleepref2.bib
+csl: vancouver.csl
 editor: 

### 2024-05-02 products/manuscript/Manuscript.qmd
check2

@@
     highlight-style: github
-bibliography: sleepref2.bib
-csl: vancouver.csl
+bibliography: ../../assets/Sleep-MADA.bib
+csl: ../../assets/apa.csl
 editor: 

### 2024-05-02 products/manuscript/Manuscript.qmd
check3

--- 2024-05-02 products/manuscript/Manuscript.qmd
@@
     highlight-style: github
-bibliography: sleepref2.bib
-csl: vancouver.csl
+bibliography: ../../assets/dataanalysis-references.bib
+csl: ../../assets/apa.csl
 editor: 

### 2024-05-02 products/manuscript/Manuscript.qmd
check4

--- 2024-05-02 products/manuscript/Manuscript.qmd
@@
     highlight-style: github
-bibliography: sleepref2.bib
-csl: vancouver.csl
+    pandoc-citeproc: true
+    bibliography: ../../assets/Sleep-MADA.bib
+    csl: ../../assets/apa.csl
 editor: 

### 2024-05-02 products/manuscript/Manuscript.qmd
check5

--- 2024-05-02 products/manuscript/Manuscript.qmd
@@
 csl: ../../assets/apa.csl
-editor: 
-  markdown: 
-    wrap: 72
 ---

### 2024-05-02 products/manuscript/Manuscript.qmd
check6

@@
     highlight-style: github
-bibliography: ../../assets/Sleep-MADA.bib
+bibliography: sleepref2.bib
 csl: ../../assets/apa.csl

### 2024-05-02 products/manuscript/Manuscript.qmd
pushing for zane to look

@@
     highlight-style: github
-bibliography: sleepref2.bib
+bibliography: ../../assets/sleepref2.bib
 csl: ../../assets/apa.csl
@@
 
+For our LASSO Regression model, used a 5-fold Cross-Validation repeated 5 times 
+to train our data. We used a tuning parameters and penalties within the tuning
+grid. Using a regularization method for this analysis was something we were 
+concerned about due to the high level of correlation between predictors;
+however, we wanted to include it to ensure a more robust analysis of our data.
+
 ## Evaluation of Models
@@
 
+![Figure 6: LASSO Tuning Results](images/lasso_tuneresults.png)
+
+Figure 6 suggests that regularization may not be the best fit for our data. 
+Though the RMSE values are small, we should be careful in analyzing this model
+due to the high level of correlation and collinearity between variables.
+
+![Figure 7: The observed vs. predicted values for our 
+"LASSO Model.](images/lasso_obsvpred.png)
+
+We can see in Figure 7 that the LASSO Model did decently well predicting on 
+the new data. The LASSO model can sufficiently predict sleep quality using
+unseen data.
+
 ![Table 1: The results of our GLMs. The "term" column identifies the
@@
 
-![Figure 6: The ranking of each variable's importance in determining
+![Figure 8: The ranking of each variable's importance in determining
 sleep quality according to our Random Forest
@@
 
-As seen in figure 6, Sleep Duration, Stress Level, and Age had the
+As seen in Figure 8, Sleep Duration, Stress Level, and Age had the
 highest levels of importance in our model. Physical Activity Level was
@@
 we learned that grouping continuous variables is a poor practice as we
-progressed in our MADA course.
+progressed in our MADA course. Our RF model performed better than the LASSO 
+model, which was to be expected. We struggled to create a variable importance
+plot for the LASSO model, which could be due to sensitive parameters and 
+tuning penalties.
 
@@
 experimental design that generated our observations.
+
+Additionally, while we conducted a LASSO regression model and received favorable 
+results, regularization methods are difficult to conduct and interpret when 
+predictors are highly correlated. The penalty paramaters appeared to be too 
+sensitive and were affected by the collinearity between predictors.
 

### 2024-05-03 products/manuscript/Manuscript.qmd
5/3- final touches

@@
 answering our primary question: which variables are most important for
-determining an individual's sleep quality and what effect do they have?
+determining an individual's sleep quality and what effect do they have [@noautho
+r_rstudio_nodate]?
 After conducting an exploratory data analysis (EDA) we hypothesized that
@@
 The data was obtained from kaggle.com at this
-[link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
+[link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data)
+[@noauthor_find_nodate].
 The data contains observations from a study on individuals of varying
@@
 showed a correlation between obesity and worsened sleep quality in men
-but not women. We are interested in seeing if this trend in present in
+but not women [@sun_poor_2016]. We are interested in seeing if this 
+trend in present in
 our dataset; however, given that most of the observations are of men, we
@@
 our data. Furthermore, it has been shown that age and gender have a
-notable effect on sleep quality (Madrid-Valero et al. 2016). Women seem
-to experience a deterioration in sleep quality as they age; however,
-while this trend is still present in men, it is notably less consistent
+notable effect on sleep quality (Madrid-Valero et al. 2016)[@madrid-valero_age_
+2017]. Women seem to experience a deterioration in sleep quality as they age; 
+however,while this trend is still present in men, it is notably less consistent
 and can vary dramatically between individuals. We expect to see similar
@@
 the worst sleep quality among civilian sector workers while 24 jobs that
-have rotating shifts tend to have the worst sleep quality (Luckhaupt et
-al. 2010). Our dataset contains observations of mostly
+have rotating shifts tend to have the worst sleep quality [@luckhaupt_prevalence
+_2010]. Our dataset contains observations of mostly
 white-collar/service-based jobs, so we expect to find similar trends
@@
 negatively impact sleep quality, we also expect that higher levels of
-blood pressure will result in poorer sleep quality. The American Sleep
-Foundation states that both short and long sleep duration increase risk
-of hypertension in adults (Li & Shang, 2021). This draws interesting
-focus to the circular nature of sleep quality and hypertension. Those
-with hypertension also face a higher risk of insomnia, a disorder
+blood pressure will result in poorer sleep quality [@sun_poor_2016]. 
+The American SleepFoundation states that both short and long sleep duration 
+increase risk of hypertension in adults [@liu_association_2022]. This draws 
+interesting focus to the circular nature of sleep quality and hypertension. 
+Those with hypertension also face a higher risk of insomnia, a disorder
 affecting the ability to fall asleep, maintain sleep, and sleep without
-continuous disturbance (Liu et. al, 2022). Short sleep duration is also
-positively associated with stress level (APA, 2013). Survey findings
+continuous disturbance [@liu_association_2022]. Short sleep duration is also
+positively associated with stress level [@noauthor_stress_nodate]. Survey findings
 indicate that the average American sleeps 6.7 hours per night, which is
@@
 activity stimulates an increase of melatonin production, a hormone that
-regulates sleep-wake cycles (Alnawwar et. al, 2023). Additionally,
+regulates sleep-wake cycles [@alnawwar_effect_2023]. Additionally,
 moderate to intense exercise has shown to have a positive effect on
-those with insomnia (Alnawwar et. al, 2023). Physical activity also
+those with insomnia [@alnawwar_effect_2023]. Physical activity also
 reduces stress and improves both sleep quality and sleep efficiency
-(Alnawwar et. al, 2023). However, studies also show that frequent
+[@alnawwar_effect_2023]. However, studies also show that frequent
 high-intensity workouts could lead to poorer sleep quality among people
-with and without insomnia (Alnawwar et. al, 2023). Studies indicate that
+with and without insomnia [@alnawwar_effect_2023]. Studies indicate that
 consistent moderate exercise three to four times a week (approximately
@@
 [American Heart Association
-guidelines](https://www.heart.org/en/health-topics/high-blood-pressure).
+guidelines](https://www.heart.org/en/health-topics/high-blood-pressure) [@noauthor_understanding_nodate].
 Additionally we created a variable named 'StepsGroup' to determine the
@@
 guidelines provided by [10,000
-steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
+steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps
+/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,
+day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_
+nodate].
 Finally, we created an additional categorical variable called
@@
 top three predictors for sleep quality in our data. This reflects the
-relationships we saw in our colinearity plot as well as the clear
+relationships we saw in our collinearity plot as well as the clear
 correlations we found between the variables in our EDA. We did not

### 2024-05-03 products/manuscript/Manuscript.qmd
5/3- check

@@
 our data. Furthermore, it has been shown that age and gender have a
-notable effect on sleep quality (Madrid-Valero et al. 2016)[@madrid-valero_age_
-2017]. Women seem to experience a deterioration in sleep quality as they age; 
+notable effect on sleep quality[@madrid-valero_age_2017]. 
+Women seem to experience a deterioration in sleep quality as they age; 
 however,while this trend is still present in men, it is notably less consistent
@@
 the worst sleep quality among civilian sector workers while 24 jobs that
-have rotating shifts tend to have the worst sleep quality [@luckhaupt_prevalence
-_2010]. Our dataset contains observations of mostly
-white-collar/service-based jobs, so we expect to find similar trends
-upon analysis.
+have rotating shifts tend to have the worst sleep quality [@luckhaupt_prevalence_2010]. 
+Our dataset contains observations of mostly white-collar/service-based jobs, so 
+we expect to find similar trends upon analysis.
 
@@
 blood pressure will result in poorer sleep quality [@sun_poor_2016]. 
-The American SleepFoundation states that both short and long sleep duration 
+The American Sleep Foundation states that both short and long sleep duration 
 increase risk of hypertension in adults [@liu_association_2022]. This draws 
@@
 indicate that the average American sleeps 6.7 hours per night, which is
-well below the minimum recommended duration seven to nine hours (APA,
-2013). Similar to the relationship between hypertension and sleep
-quality, as stress level increases, sleep duration decreases thereby
-causing poorer sleep quality. Poorer sleep quality then leads to higher
-stress, greater risk of hypertension, and shorter sleep duration.
+well below the minimum recommended duration seven to nine hours [@noauthor_stress_nodate]. 
+Similar to the relationship between hypertension and sleep quality, as stress 
+level increases, sleep duration decreases thereby causing poorer sleep quality. 
+Poorer sleep quality then leads to higher stress, greater risk of hypertension, 
+and shorter sleep duration.
 
@@
 /#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,
-day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_
-nodate].
+day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_nodate].
 Finally, we created an additional categorical variable called

### 2024-05-03 products/manuscript/Manuscript.qmd
The End
Finally done!

@@
 
-For our project, we obtained a dataset from Kaggle.com containing 373
-observations of 13 variables from a study on different factors that may
-affect sleep quality. Variables of interest from the study included
-physical activity level, age, sex, occupation, stress, gender, BMI,
-blood pressure, and sleep duration. In
-[RStudio](https://posit.co/download/rstudio-desktop/) (version 4.3.1),
-we used a combination of packages to create models suitable for
-answering our primary question: which variables are most important for
-determining an individual's sleep quality and what effect do they have [@noautho
-r_rstudio_nodate]?
-After conducting an exploratory data analysis (EDA) we hypothesized that
-stress, sleep duration, and occupation would have the largest impact on
-sleep quality. We believed that stress and sleep quality would be
-negatively correlated, sleep duration and quality would be positively
-correlated, and that jobs with longer or more variable hours would
-negatively affect sleep quality.
+For our project, we obtained a dataset from Kaggle.com containing 373 observations of 13 variables from a study on different factors that may affect sleep quality.
+Variables of interest from the study included physical activity level, age, sex, occupation, stress, gender, BMI, blood pressure, and sleep duration.
+In [RStudio](https://posit.co/download/rstudio-desktop/) (version 4.3.1), we used a combination of packages to create models suitable for answering our primary question: which variables are most important for determining an individual's sleep quality and what effect do they have?
+After conducting an exploratory data analysis (EDA) we hypothesized that stress, sleep duration, and occupation would have the largest impact on sleep quality.
+We believed that stress and sleep quality would be negatively correlated, sleep duration and quality would be positively correlated, and that jobs with longer or more variable hours would negatively affect sleep quality.
 
-After creating linear models for each variable's impact on sleep quality
-and fitting a Random Forest model with Cross-Validation to test variable
-importance, we determined that sleep duration, stress level, and age
-were the most important factors. Our hypothesis for sleep duration and
-stress was supported; however, we did not correctly predict that age
-would be one of the most important predictors for sleep quality. Our
-hypotheses regarding the effects of sleep duration, stress, and
-occupation were all supported as well. Overall, we determined that an
-individual's duration of sleep, stress level, and age are all key
-factors that can impact sleep quality.
+After creating linear models for each variable's impact on sleep quality and fitting a Random Forest model with Cross-Validation to test variable importance, we determined that sleep duration, stress level, and age were the most important factors.
+Our hypothesis for sleep duration and stress was supported; however, we did not correctly predict that age would be one of the most important predictors for sleep quality.
+Our hypotheses regarding the effects of sleep duration, stress, and occupation were all supported as well.
+Overall, we determined that an individual's duration of sleep, stress level, and age are all key factors that can impact sleep quality.
 
@@
 
-The data was obtained from kaggle.com at this
-[link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data)
-[@noauthor_find_nodate].
-The data contains observations from a study on individuals of varying
-sexes and age regarding their sleep quality and different variables that
-may affect this. We are unsure of where this data set came from and how
-the data were collected; the publisher on Kaggle has yet to respond to a
-comment with this question posted on March 5th 2024. Regardless, it
-contains interesting variables that have been shown in previous studies
-to have an effect on sleep quality.
+The data was obtained from kaggle.com at this [link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data) [@noauthor_find_nodate].
+The data contains observations from a study on individuals of varying sexes and age regarding their sleep quality and different variables that may affect this.
+We are unsure of where this data set came from and how the data were collected; the publisher on Kaggle has yet to respond to a comment with this question posted on March 5th 2024.
+Regardless, it contains interesting variables that have been shown in previous studies to have an effect on sleep quality.
 
-For instance, in Sun et al. (2015), a sample of Chinese individuals
-showed a correlation between obesity and worsened sleep quality in men
-but not women [@sun_poor_2016]. We are interested in seeing if this 
-trend in present in
-our dataset; however, given that most of the observations are of men, we
-expect to see a correlation between lower sleep scores and obesity in
-our data. Furthermore, it has been shown that age and gender have a
-notable effect on sleep quality[@madrid-valero_age_2017]. 
-Women seem to experience a deterioration in sleep quality as they age; 
-however,while this trend is still present in men, it is notably less consistent
-and can vary dramatically between individuals. We expect to see similar
-trends in our dataset, but as stated previously, our observations
-contain significantly more men than women. Finally, we wish to test
-whether or not certain occupational industries have different effects on
-sleep quality. It has been shown that managerial positions tend to have
-the worst sleep quality among civilian sector workers while 24 jobs that
-have rotating shifts tend to have the worst sleep quality [@luckhaupt_prevalence_2010]. 
-Our dataset contains observations of mostly white-collar/service-based jobs, so 
-we expect to find similar trends upon analysis.
+For instance, in Sun et al.
+(2015), a sample of Chinese individuals showed a correlation between obesity and worsened sleep quality in men but not women [@sun_poor_2016].
+We are interested in seeing if this trend in present in our dataset; however, given that most of the observations are of men, we expect to see a correlation between lower sleep scores and obesity in our data.
+Furthermore, it has been shown that age and gender have a notable effect on sleep quality[@madrid-valero_age_2017].
+Women seem to experience a deterioration in sleep quality as they age; however,while this trend is still present in men, it is notably less consistent and can vary dramatically between individuals.
+We expect to see similar trends in our dataset, but as stated previously, our observations contain significantly more men than women.
+Finally, we wish to test whether or not certain occupational industries have different effects on sleep quality.
+It has been shown that managerial positions tend to have the worst sleep quality among civilian sector workers while 24 jobs that have rotating shifts tend to have the worst sleep quality [@luckhaupt_prevalence_2010].
+Our dataset contains observations of mostly white-collar/service-based jobs, so we expect to find similar trends upon analysis.
 
-Additionally, based on the findings of Sun et. al that obesity can
-negatively impact sleep quality, we also expect that higher levels of
-blood pressure will result in poorer sleep quality [@sun_poor_2016]. 
-The American Sleep Foundation states that both short and long sleep duration 
-increase risk of hypertension in adults [@liu_association_2022]. This draws 
-interesting focus to the circular nature of sleep quality and hypertension. 
-Those with hypertension also face a higher risk of insomnia, a disorder
-affecting the ability to fall asleep, maintain sleep, and sleep without
-continuous disturbance [@liu_association_2022]. Short sleep duration is also
-positively associated with stress level [@noauthor_stress_nodate]. Survey findings
-indicate that the average American sleeps 6.7 hours per night, which is
-well below the minimum recommended duration seven to nine hours [@noauthor_stress_nodate]. 
-Similar to the relationship between hypertension and sleep quality, as stress 
-level increases, sleep duration decreases thereby causing poorer sleep quality. 
-Poorer sleep quality then leads to higher stress, greater risk of hypertension, 
-and shorter sleep duration.
+Additionally, based on the findings of Sun et.
+al that obesity can negatively impact sleep quality, we also expect that higher levels of blood pressure will result in poorer sleep quality [@sun_poor_2016].
+The American Sleep Foundation states that both short and long sleep duration increase risk of hypertension in adults [@liu_association_2022].
+This draws interesting focus to the circular nature of sleep quality and hypertension.
+Those with hypertension also face a higher risk of insomnia, a disorder affecting the ability to fall asleep, maintain sleep, and sleep without continuous disturbance [@liu_association_2022].
+Short sleep duration is also positively associated with stress level [@noauthor_stress_nodate].
+Survey findings indicate that the average American sleeps 6.7 hours per night, which is well below the minimum recommended duration seven to nine hours [@noauthor_stress_nodate].
+Similar to the relationship between hypertension and sleep quality, as stress level increases, sleep duration decreases thereby causing poorer sleep quality.
+Poorer sleep quality then leads to higher stress, greater risk of hypertension, and shorter sleep duration.
 
-Engaging in physical activity has been suggested to improve both sleep
-quality and sleep duration. Regular moderate to intense physical
-activity stimulates an increase of melatonin production, a hormone that
-regulates sleep-wake cycles [@alnawwar_effect_2023]. Additionally,
-moderate to intense exercise has shown to have a positive effect on
-those with insomnia [@alnawwar_effect_2023]. Physical activity also
-reduces stress and improves both sleep quality and sleep efficiency
-[@alnawwar_effect_2023]. However, studies also show that frequent
-high-intensity workouts could lead to poorer sleep quality among people
-with and without insomnia [@alnawwar_effect_2023]. Studies indicate that
-consistent moderate exercise three to four times a week (approximately
-150 minutes total) is the optimal weekly regimen to improve sleep
-quality and reduce stress. We expect to see similar trends with our data
-where increased stress levels negatively impact sleep quality. We expect
-to see a positive association between increasing physical activity
-levels and sleep quality until the point of "too much" activity.
+Engaging in physical activity has been suggested to improve both sleep quality and sleep duration.
+Regular moderate to intense physical activity stimulates an increase of melatonin production, a hormone that regulates sleep-wake cycles [@alnawwar_effect_2023].
+Additionally, moderate to intense exercise has shown to have a positive effect on those with insomnia [@alnawwar_effect_2023].
+Physical activity also reduces stress and improves both sleep quality and sleep efficiency [@alnawwar_effect_2023].
+However, studies also show that frequent high-intensity workouts could lead to poorer sleep quality among people with and without insomnia [@alnawwar_effect_2023].
+Studies indicate that consistent moderate exercise three to four times a week (approximately 150 minutes total) is the optimal weekly regimen to improve sleep quality and reduce stress.
+We expect to see similar trends with our data where increased stress levels negatively impact sleep quality.
+We expect to see a positive association between increasing physical activity levels and sleep quality until the point of "too much" activity.
 
@@
 
-The dataset originally contained 373 observations for 13 variables
-covering a broad spectrum of health and lifestyle variables associated
-with sleep quality. The curators of the dataset did not provide
-information on how or where this data was collected, so we solely relied
-on the codebook for variable definitions. We have added this codebook to
-the supplementary materials file for ease of access but you can also
-find it on the Kaggle page for this dataset. There were no missing
-values or erroneous variables in the dataset that we needed to remove.
-The "Blood.Pressure" variable was transformed to reflect the individual
-systolic and diastolic blood pressure measurements from each subject. We
-then chose to add an additional variable, labeled 'cat_bp' to reflect
-the categorical blood pressure status of a subject based on the
-[American Heart Association
-guidelines](https://www.heart.org/en/health-topics/high-blood-pressure) [@noauthor_understanding_nodate].
-Additionally we created a variable named 'StepsGroup' to determine the
-categorical activity level of subjects in the dataset based on these
-guidelines provided by [10,000
-steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps
-/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,
-day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_nodate].
-Finally, we created an additional categorical variable called
-'PhysicalActivityGroup' to categorically represent the level of physical
-activity reported by a subject. The levels are differentiated by 30
-minute intervals of weekly physical activity. The outcome of interest is
-noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
-Quality of Sleep throughout this report.
+The dataset originally contained 373 observations for 13 variables covering a broad spectrum of health and lifestyle variables associated with sleep quality.
+The curators of the dataset did not provide information on how or where this data was collected, so we solely relied on the codebook for variable definitions.
+We have added this codebook to the supplementary materials file for ease of access but you can also find it on the Kaggle page for this dataset.
+There were no missing values or erroneous variables in the dataset that we needed to remove.
+The "Blood.Pressure" variable was transformed to reflect the individual systolic and diastolic blood pressure measurements from each subject.
+We then chose to add an additional variable, labeled 'cat_bp' to reflect the categorical blood pressure status of a subject based on the [American Heart Association guidelines](https://www.heart.org/en/health-topics/high-blood-pressure) [@noauthor_understanding_nodate].
+Additionally we created a variable named 'StepsGroup' to determine the categorical activity level of subjects in the dataset based on these guidelines provided by [10,000 steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps%20/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,%20day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_nodate].
+Finally, we created an additional categorical variable called 'PhysicalActivityGroup' to categorically represent the level of physical activity reported by a subject.
+The levels are differentiated by 30 minute intervals of weekly physical activity.
+The outcome of interest is noted as "Quality.of.Sleep" that we will reference as Sleep Quality or Quality of Sleep throughout this report.
 
@@
 
-Given that there were only 17 variables after final data transformation,
-simple linear regression models were performed to determine baseline
-associations between variables and Sleep Quality (the outcome of
-interest). These regression models were created to determine which
-variables had the largest impact on Sleep Quality and drove the rest of
-the analyses. Simple linear regression models were fitted for the
-following variables: BMI, cat_bp, Stress Level, Physical Activity Level,
-Sleep Duration, Gender, Age, and Occupation. The other variables were
-determined to have too weak of a relationship after our EDA and were
-therefore left out of the analyses.
+Given that there were only 17 variables after final data transformation, simple linear regression models were performed to determine baseline associations between variables and Sleep Quality (the outcome of interest).
+These regression models were created to determine which variables had the largest impact on Sleep Quality and drove the rest of the analyses.
+Simple linear regression models were fitted for the following variables: BMI, cat_bp, Stress Level, Physical Activity Level, Sleep Duration, Gender, Age, and Occupation.
+The other variables were determined to have too weak of a relationship after our EDA and were therefore left out of the analyses.
 
-The Occupation variable was then transformed to group the various
-occupations into the following groups: Healthcare, Education,
-Engineering, Business/Finance, and Science. Prior to performing any
-subsequent analyses, we removed the original versions of the variables
-we transformed: systolic, diastolic, daily steps, occupation,
-Physical.Activity.Level, Heart.Rate, and AgeGroup.
+The Occupation variable was then transformed to group the various occupations into the following groups: Healthcare, Education, Engineering, Business/Finance, and Science.
+Prior to performing any subsequent analyses, we removed the original versions of the variables we transformed: systolic, diastolic, daily steps, occupation, Physical.Activity.Level, Heart.Rate, and AgeGroup.
 
-The objective of our analysis was to determine which of these variables
-are most important for determining an individual's sleep quality and
-what specific effect they have. After conducting an EDA we hypothesized
-that stress, sleep duration, and occupation would have the largest
-impact on sleep quality. We believed that stress and sleep quality would
-be negatively correlated, sleep duration and quality would be positively
-correlated, and that jobs with longer or more variable hours would
-negatively affect sleep quality.
+The objective of our analysis was to determine which of these variables are most important for determining an individual's sleep quality and what specific effect they have.
+After conducting an EDA we hypothesized that stress, sleep duration, and occupation would have the largest impact on sleep quality.
+We believed that stress and sleep quality would be negatively correlated, sleep duration and quality would be positively correlated, and that jobs with longer or more variable hours would negatively affect sleep quality.
 
@@
 
-Random Forest Models and 5-fold Cross-Validation were considered to be
-the "best" fit for our outcome of interest, Quality of Sleep. The data
-were not split into train/test subsets as the data contains less than
-400 observations and many of the values were unique in comparison to the
-rest of the data. Prior to constructing this model, a colinearity plot
-was constructed to determine any presence of colinearity in our data.
+Random Forest Models and 5-fold Cross-Validation were considered to be the "best" fit for our outcome of interest, Quality of Sleep.
+The data were not split into train/test subsets as the data contains less than 400 observations and many of the values were unique in comparison to the rest of the data.
+Prior to constructing this model, a colinearity plot was constructed to determine any presence of colinearity in our data.
 
-![Figure 1: A colinearity plot demonstrating the colinearity between our
-variables.](images/collinearity.png){width="586"}
+![Figure 1: A colinearity plot demonstrating the colinearity between our variables.](images/collinearity.png){width="586"}
 
-Stress and Sleep Quality had an absolute value of 0.9 on the correlation
-scale. Sleep Duration and Stress were also seen to be strongly
-correlated. Stress was still included in our models due to its
-well-studied impact on sleep; however, we found it important to note the
-strong linear relationship between Stress and Sleep Quality.
+Stress and Sleep Quality had an absolute value of 0.9 on the correlation scale.
+Sleep Duration and Stress were also seen to be strongly correlated.
+Stress was still included in our models due to its well-studied impact on sleep; however, we found it important to note the strong linear relationship between Stress and Sleep Quality.
 
@@
 
-We fit GLMs for each predictor in order to determine their effect on
-sleep quality. Each model only used a single predictor as the only aim
-was to quantify and understand the impact of each for our outcome of
-interest. For our Random Forest model, we used mtry (with a range of
-1-7) and min_n (with a range of 1-21) as tuning parameters to create 300
-different decision trees. Mtry controls how many predictors a decision
-tree can consider at any given point in time. It adds randomness to the
-decision tree creation process and ensures that all of the trees do not
-look the same. Min_n specifies the minimum number of samples that should
-be present in the leaf node after splitting a node. In our case, this
-means that if any terminal node has more than 21 observations and is not
-a pure node, we can split it further into subnodes. This was all done
-using 5-fold Cross-Validation repeated 5 times in order to train the
-data to predict on an unseen dataset.
+We fit GLMs for each predictor in order to determine their effect on sleep quality.
+Each model only used a single predictor as the only aim was to quantify and understand the impact of each for our outcome of interest.
+For our Random Forest model, we used mtry (with a range of 1-7) and min_n (with a range of 1-21) as tuning parameters to create 300 different decision trees.
+Mtry controls how many predictors a decision tree can consider at any given point in time.
+It adds randomness to the decision tree creation process and ensures that all of the trees do not look the same.
+Min_n specifies the minimum number of samples that should be present in the leaf node after splitting a node.
+In our case, this means that if any terminal node has more than 21 observations and is not a pure node, we can split it further into subnodes.
+This was all done using 5-fold Cross-Validation repeated 5 times in order to train the data to predict on an unseen dataset.
 
-For our LASSO Regression model, used a 5-fold Cross-Validation repeated 5 times 
-to train our data. We used a tuning parameters and penalties within the tuning
-grid. Using a regularization method for this analysis was something we were 
-concerned about due to the high level of correlation between predictors;
-however, we wanted to include it to ensure a more robust analysis of our data.
+For our LASSO Regression model, used a 5-fold Cross-Validation repeated 5 times to train our data.
+We used a tuning parameters and penalties within the tuning grid.
+Using a regularization method for this analysis was something we were concerned about due to the high level of correlation between predictors; however, we wanted to include it to ensure a more robust analysis of our data.
 
@@
 
-Root Mean Squared Error (RMSE) is a common metric used to evaluate
-regression models. The RMSE is formally defined as the square root of
-the mean square of all error and is defined by the following formula:
-$$\sum_{i=1}^{D}(x_i-y_i)^2$$
+Root Mean Squared Error (RMSE) is a common metric used to evaluate regression models.
+The RMSE is formally defined as the square root of the mean square of all error and is defined by the following formula: $$\sum_{i=1}^{D}(x_i-y_i)^2$$
 
-It is important to consider that while RMSE is scale dependent, common
-practice notes that low RMSE values indicate stronger model performance.
+It is important to consider that while RMSE is scale dependent, common practice notes that low RMSE values indicate stronger model performance.
 
-R-squared (R²) is a statistical measure used in regression analyses. It
-quantifies the proportion of the variance in a dependent variable that
-can be explained by an independent variable within a regression model
-and is calculated with the following formula:
+R-squared (R²) is a statistical measure used in regression analyses.
+It quantifies the proportion of the variance in a dependent variable that can be explained by an independent variable within a regression model and is calculated with the following formula:
 
@@
 
-Null models provide a baseline for the evaluation of a model's
-performance. They generate predictions for an outcome of interest
-without using any predictor variables. Instead, they simply average the
-values of the outcome of interest to make these "predictions". We used
-both RMSE and R² values to evaluate the performance of our model
-compared to a "null model".
+Null models provide a baseline for the evaluation of a model's performance.
+They generate predictions for an outcome of interest without using any predictor variables.
+Instead, they simply average the values of the outcome of interest to make these "predictions".
+We used both RMSE and R² values to evaluate the performance of our model compared to a "null model".
 
@@
 
-This analysis was conducted under R version 4.3.1 on a MacOS operating
-system. The following R packages were used in the development of these
-analyses: here, skimr, broom, tidyverse, ggplot2, dplyr, corrplot,
-ranger, and vip. All processing and analysis code can be found in the
-Supplementary Material file.
+This analysis was conducted under R version 4.3.1 on a MacOS operating system.
+The following R packages were used in the development of these analyses: here, skimr, broom, tidyverse, ggplot2, dplyr, corrplot, ranger, and vip.
+All processing and analysis code can be found in the Supplementary Material file.
 
@@
 
-The 'Quality of Sleep' variable has been selected as our outcome of
-interest. It is important to note that this variable is a subjective
-measure of a participant's self-reported sleep quality. Figure 2 shows
-that most subjects reported a sleep quality score of 8 (out of 10).
+The 'Quality of Sleep' variable has been selected as our outcome of interest.
+It is important to note that this variable is a subjective measure of a participant's self-reported sleep quality.
+Figure 2 shows that most subjects reported a sleep quality score of 8 (out of 10).
 
-![Figure 2: Distribution of Sleep Quality Scores in the
-dataset.](images/unnamed-chunk-17-1.png)
+![Figure 2: Distribution of Sleep Quality Scores in the dataset.](images/unnamed-chunk-17-1.png)
 
-Our Random Forest Model built with 5-fold Cross-Validation explored
-different combinations of predictors to determine which parameters gave
-the most accurate predictions of sleep quality.
+Our Random Forest Model built with 5-fold Cross-Validation explored different combinations of predictors to determine which parameters gave the most accurate predictions of sleep quality.
 
-![Figure 3: The results generated from our RF model. RMSE increases when
-more predictors are added to the model. The R-squared value also
-decreases as more predictors are added to the
-model.](images/RFtuningresults.png)
+![Figure 3: The results generated from our RF model.
+RMSE increases when more predictors are added to the model.
+The R-squared value also decreases as more predictors are added to the model.](images/RFtuningresults.png)
 
-Figure 3 shows that as we include more predictors in our model, RMSE
-increases slightly while the R-squared value decreases. This indicates
-that some variables in our dataset are more important for predicting
-sleep quality than the others. Based off of our variable importance
-graph, we can assume that stress level, sleep duration, and age are the
-factors that carry the most weight when predicting sleep quality.
+Figure 3 shows that as we include more predictors in our model, RMSE increases slightly while the R-squared value decreases.
+This indicates that some variables in our dataset are more important for predicting sleep quality than the others.
+Based off of our variable importance graph, we can assume that stress level, sleep duration, and age are the factors that carry the most weight when predicting sleep quality.
 
-We went on to evaluate how well our RF model predicted values by
-creating an observed vs. predicted plot as seen in figure 4.
+We went on to evaluate how well our RF model predicted values by creating an observed vs. predicted plot as seen in figure 4.
 
-![Figure 4: A comparison of the actual values for sleep quality in our
-dataset (observed) compared to the values predicted by our fitted RF
-model.](images/observed_vs_predicted.png)
+![Figure 4: A comparison of the actual values for sleep quality in our dataset (observed) compared to the values predicted by our fitted RF model.](images/observed_vs_predicted.png)
 
-While the predicted values don't align perfectly with the red "ideal
-value" line, they fall very close overall, indicating that our model can
-sufficiently predict sleep quality using unseen data.
+While the predicted values don't align perfectly with the red "ideal value" line, they fall very close overall, indicating that our model can sufficiently predict sleep quality using unseen data.
 
-![Figure 5: The observed vs. predicted values for our "Null
-model".](images/null_observedvspredicted.png)
+![Figure 5: The observed vs. predicted values for our "Null model".](images/null_observedvspredicted.png)
 
-We can see in figure 5 that the null model only predicts values that are
-the average of our observed values. As expected, these predictions are
-far from our actual observed values.
+We can see in figure 5 that the null model only predicts values that are the average of our observed values. As expected, these predictions are far from our actual observed values.
 
@@
 Figure 6 suggests that regularization may not be the best fit for our data.
-Though the RMSE values are small, we should be careful in analyzing this model
-due to the high level of correlation and collinearity between variables.
+Though the RMSE values are small, we should be careful in analyzing this model due to the high level of correlation and collinearity between variables.
 
-![Figure 7: The observed vs. predicted values for our 
-"LASSO Model.](images/lasso_obsvpred.png)
+![Figure 7: The observed vs. predicted values for our "LASSO Model.](images/lasso_obsvpred.png)
 
-We can see in Figure 7 that the LASSO Model did decently well predicting on 
-the new data. The LASSO model can sufficiently predict sleep quality using
-unseen data.
+We can see in Figure 7 that the LASSO Model did decently well predicting on the new data. The LASSO model can sufficiently predict sleep quality using unseen data.
 
-![Table 1: The results of our GLMs. The "term" column identifies the
-category/variable being analyzed. The "model" column identifies the
-predictors used to fit each model. The "statistic" column is a relative
-measure of the strength of the relationship between each predictor and
-sleep quality. The "estimate" column shows the predicted change in sleep
-quality score as a result of each associated predictor. Each reference
-category that is used for categorical variables is included next to the
-term in parentheses (e.g. Reference: Female).](images/table-01.png)
+![Table 1: The results of our GLMs.
+The "term" column identifies the category/variable being analyzed.
+The "model" column identifies the predictors used to fit each model.
+The "statistic" column is a relative measure of the strength of the relationship between each predictor and sleep quality.
+The "estimate" column shows the predicted change in sleep quality score as a result of each associated predictor.
+Each reference category that is used for categorical variables is included next to the term in parentheses (e.g.
+Reference: Female).](images/table-01.png)
 
-We believed that stress and sleep quality would be negatively
-correlated, sleep duration and quality would be positively correlated,
-and that jobs with longer or more variable hours would negatively affect
-sleep quality. As seen in table 1, we found that all of our assumptions
-were supported. The "intercept" values under the term column represents
-what the expected sleep quality score would be if a continuous predictor
-was at a value of 0. For categorical variables, it represents the
-reference category that the others are being compared to.
+We believed that stress and sleep quality would be negatively correlated, sleep duration and quality would be positively correlated, and that jobs with longer or more variable hours would negatively affect sleep quality.
+As seen in table 1, we found that all of our assumptions were supported.
+The "intercept" values under the term column represents what the expected sleep quality score would be if a continuous predictor was at a value of 0.
+For categorical variables, it represents the reference category that the others are being compared to.
 
@@
 
-A variable importance plot (VIP) (figure 6) was constructed using the
-"best" set of parameters (as guided by our RF model) to determine which
-variables had the strongest impact on sleep quality.
+A variable importance plot (VIP) (figure 6) was constructed using the "best" set of parameters (as guided by our RF model) to determine which variables had the strongest impact on sleep quality.
 
-![Figure 8: The ranking of each variable's importance in determining
-sleep quality according to our Random Forest
-model.](images/variableimportance.png)
+![Figure 8: The ranking of each variable's importance in determining sleep quality according to our Random Forest model.](images/variableimportance.png)
 
-As seen in Figure 8, Sleep Duration, Stress Level, and Age had the
-highest levels of importance in our model. Physical Activity Level was
-excluded entirely, which was to be expected given our exploratory data
-analysis and simple linear regression models.
+As seen in Figure 8, Sleep Duration, Stress Level, and Age had the highest levels of importance in our model.
+Physical Activity Level was excluded entirely, which was to be expected given our exploratory data analysis and simple linear regression models.
 
-![Table 2: A comparison of the RMSE and R-squared values generated by
-the null and RF models. The null model has no R-squared value due to its
-method of generating predictions.](images/metric_table.png){width="510"}
+![Table 2: A comparison of the RMSE and R-squared values generated by the null and RF models.
+The null model has no R-squared value due to its method of generating predictions.](images/metric_table.png){width="510"}
 
-As seen in table 2, our null model was found to have an RMSE of 1.193;
-compared to our best parameter RF model's RMSE of 0.059, this is
-relatively high. The R-squared value doesn't exist for the null model
-since there is technically no variation between the observed and
-predicted values.
+As seen in table 2, our null model was found to have an RMSE of 1.193; compared to our best parameter RF model's RMSE of 0.059, this is relatively high.
+The R-squared value doesn't exist for the null model since there is technically no variation between the observed and predicted values.
 
@@
 
-The differences between the null model and RF model's metrics show that
-our RF model is better suited for predicting sleep quality on unseen
-data than a model using no predictors. The R-squared value of 0.997 is
-extremely high; this shows that a large amount of variation is explained
-using our RF model. The variable importance plot we created answers our
-questions about which variables are most important for predicting sleep
-quality. As we can see, sleep duration, stress level, and age were the
-top three predictors for sleep quality in our data. This reflects the
-relationships we saw in our collinearity plot as well as the clear
-correlations we found between the variables in our EDA. We did not
-expect to see age so high on the list; this could be due to us grouping
-ages together in our EDA and therefore misleading us in our preliminary
-data exploration. We did not use the groups when creating our models as
-we learned that grouping continuous variables is a poor practice as we
-progressed in our MADA course. Our RF model performed better than the LASSO 
-model, which was to be expected. We struggled to create a variable importance
-plot for the LASSO model, which could be due to sensitive parameters and 
-tuning penalties.
+The differences between the null model and RF model's metrics show that our RF model is better suited for predicting sleep quality on unseen data than a model using no predictors.
+The R-squared value of 0.997 is extremely high; this shows that a large amount of variation is explained using our RF model.
+The variable importance plot we created answers our questions about which variables are most important for predicting sleep quality.
+As we can see, sleep duration, stress level, and age were the top three predictors for sleep quality in our data.
+This reflects the relationships we saw in our collinearity plot as well as the clear correlations we found between the variables in our EDA.
+We did not expect to see age so high on the list; this could be due to us grouping ages together in our EDA and therefore misleading us in our preliminary data exploration.
+We did not use the groups when creating our models as we learned that grouping continuous variables is a poor practice as we progressed in our MADA course.
+Our RF model performed better than the LASSO model, which was to be expected.
+We struggled to create a variable importance plot for the LASSO model, which could be due to sensitive parameters and tuning penalties.
 
-Through our linear models, we determined that our hypotheses regarding
-each variable's effect on quality of sleep were supported. The variables
-determined to be the most important by our RF model had the following
-effects on sleep quality: sleep duration was positively correlated, age
-was positively correlated, and stress level was negatively correlated.
+Through our linear models, we determined that our hypotheses regarding each variable's effect on quality of sleep were supported.
+The variables determined to be the most important by our RF model had the following effects on sleep quality: sleep duration was positively correlated, age was positively correlated, and stress level was negatively correlated.
 
@@
 
-The main limitations of our models stemmed from the data itself. As
-stated in our introduction section, we have been unable to identify
-exactly where the data were collected and how it was gathered. The
-values for stress level and sleep duration were also based on a
-subjective rating given by each participant in the study; this could
-introduce individual bias to the data and prevent an objective
-evaluation of the effect of these variables. An inherent limitation in
-RF models is difficulty of interpretation; we attempted to minimize this
-by creating easy to read tables and graphs for the results. We chose not
-to split our data into a train/test split due to the small number of
-observations and presence of rare values in our data. We believe that
-fitting using CV minimized the downsides of this. Future analyses would
-benefit from more detailed meta data and a better understanding of the
-experimental design that generated our observations.
+The main limitations of our models stemmed from the data itself.
+As stated in our introduction section, we have been unable to identify exactly where the data were collected and how it was gathered.
+The values for stress level and sleep duration were also based on a subjective rating given by each participant in the study; this could introduce individual bias to the data and prevent an objective evaluation of the effect of these variables.
+An inherent limitation in RF models is difficulty of interpretation; we attempted to minimize this by creating easy to read tables and graphs for the results.
+We chose not to split our data into a train/test split due to the small number of observations and presence of rare values in our data.
+We believe that fitting using CV minimized the downsides of this.
+Future analyses would benefit from more detailed meta data and a better understanding of the experimental design that generated our observations.
 
-Additionally, while we conducted a LASSO regression model and received favorable 
-results, regularization methods are difficult to conduct and interpret when 
-predictors are highly correlated. The penalty paramaters appeared to be too 
-sensitive and were affected by the collinearity between predictors.
+Additionally, while we conducted a LASSO regression model and received favorable results, regularization methods are difficult to conduct and interpret when predictors are highly correlated.
+The penalty paramaters appeared to be too sensitive and were affected by the collinearity between predictors.
 
-The strengths of our analysis lie in the easily interpretation of our
-GLMs and the predictive power of an RF model. We were able to clearly
-display and quantify the relationships between our predictors and sleep
-quality, answering our question about each variable's effect on our
-outcome of interest. We were also able to create a powerful RF model
-that explained nearly all of the variation in our data as well as a
-ranking of each variable's impact on sleep quality. Fitting our RF model
-with CV allowed us to train the model to create predictions for "unseen"
-data as well.
+The strengths of our analysis lie in the easily interpretation of our GLMs and the predictive power of an RF model.
+We were able to clearly display and quantify the relationships between our predictors and sleep quality, answering our question about each variable's effect on our outcome of interest.
+We were also able to create a powerful RF model that explained nearly all of the variation in our data as well as a ranking of each variable's impact on sleep quality.
+Fitting our RF model with CV allowed us to train the model to create predictions for "unseen" data as well.
 
@@
 
-Overall, while our study limitations prevent us from generalizing the
-findings of this analysis broadly, we were able to clearly answer our
-questions about the data. Our findings supported our hypotheses about
-the effects of each predictor on sleep quality. However, our hypothesis
-about which variables would be the most important was not supported by
-the analysis. Instead, we determined that stress level, sleep duration,
-and age were the most important predictors for our outcome of interest.
-Ultimately, the results indicate that the sleep quality of this study's
-participants is impacted disproportionately by an assortment of
-variables dealing with physical, mental, and environmental factors.
+Overall, while our study limitations prevent us from generalizing the findings of this analysis broadly, we were able to clearly answer our questions about the data.
+Our findings supported our hypotheses about the effects of each predictor on sleep quality.
+However, our hypothesis about which variables would be the most important was not supported by the analysis.
+Instead, we determined that stress level, sleep duration, and age were the most important predictors for our outcome of interest.
+Ultimately, the results indicate that the sleep quality of this study's participants is impacted disproportionately by an assortment of variables dealing with physical, mental, and environmental factors.
 
